### 2015-04-14 intro.tex
first commit of files and text

\section*{Introduction}
Although maize is naturally an out-crossing organism, modern breeding develops highly inbred lines that are then used in controlled crosses to produce hybrids. Hybrid maize has been so successful that it quickly replaced long-standing mass-selected open pollinated varieties (Crabb, 1947). 
Maize inbred lines are now partitioned into separate inbreeding ‘heterotic groups’ that maximize performance and hybrid vigor (heterosis) when an inbred line of one group is crossed with the other group (Tracy and Chandler, 2006). 
The shift towards selection of inbred lines based on their ability to generate good hybrids – referred to as ‘combining ability’ – constituted an abrupt change from the open-pollinated mass selection that breeders practiced for millennia (Anderson, 1944; Troyer, 1999).
Multiple studies with molecular markers have indicated that the modern era of single-cross hybrid maize breeding has led to a dramatic restructuring of population genetic variation (Duvick et al., 2004; Ho et al., 2005; Feng et al., 2006). 
Different heterotic groups have diverged genetically over time to become highly structured and isolated populations. 
Advances in high throughput genotyping and the development of a maize reference genome now enable the observation of maize population structure at high marker density across the whole genome \citep{ganal2011a-large,chia2012maize}. 
So far, these high-density studies have examined a broad spectrum of germplasm at various points in the history of maize to search for the signals of population structure and artificial selection \citep{Hufford2012b, van2012historical}. 
Although selective sweeps remaining from domestication are clearly visible, the impact of selection during modern breeding appears comparatively small in terms of its impact on genomic diversity despite steady, heritable improvement in phenotype (Duvick, 2005). 
The lack of distinct selection signals from modern breeding may be due to specific selective events occurring in different populations, necessitating a more focused look within heterotic groups or even single breeding programs. 
In this study, we apply a genomic approach to study the dynamics of genetic variation over time within an individual selection experiment. 
The Iowa Stiff Stalk Synthetic (BSSS) and the Iowa Corn Borer Synthetic No. 1 (BSCB1) Reciprocal Recurrent Selection Program of the USDA-ARS at Ames, Iowa (hereafter referred to as the Iowa RRS) represents one of the best-documented public experiments on selection for combining ability and hybrid performance. 
BSSS and BSCB1 have been recurrently selected for improved cross-population hybrids (Penny and Eberhart, 1971). 
This model of selection, named reciprocal recurrent selection, provides the generalized model for strategies used in commercial maize breeding (Comstock et al., 1949; Duvick et al., 2004). 
The Iowa RRS experiment proves especially relevant because lines derived from the BSSS population have had a major impact upon the development of commercial hybrids (Duvick et al., 2004; Darrah and Zuber, 1986), the formation of modern heterotic groups (Troyer, 1999; Senior et al., 1998), and the choice of a maize reference genome \citep{schnable2009the-b73-maize}.

### 2015-04-14 main.tex
first commit of files and text

\documentclass[twocolumn,twoside,letterpaper]{article}

\usepackage{color}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{geneticsT2}
\usepackage{times}
\usepackage{hyperref}

\addtolength{\oddsidemargin}{-.2cm}
\addtolength{\evensidemargin}{-1.2cm}
\addtolength{\textwidth}{1.5cm}
\addtolength{\topmargin}{-2cm}
\addtolength{\textheight}{3.5cm}

\renewcommand{\textfraction}{0.01}
\renewcommand{\topfraction}{0.99}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.90}
\renewcommand{\dbltopfraction}{0.95}
\renewcommand{\dblfloatpagefraction}{0.80}
\renewcommand{\sfdefault}{phv}

% %%%
% % for latexml
% \usepackage{latexml}
% 
% % only do this stuff if latexml is not parsing the document
% \iflatexml
%   \renewcommand\sfbf[1]{\sf{\bfseries#1}}
% \else 
% % do this stuff:
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}

\fancyfoot[LE,RO]{{\sfbf \thepage}}
\renewcommand{\headrulewidth}{0pt}
\fancypagestyle{plain}{
	\fancyhf{}
}

%editing commands (please leave in for now)
\newcommand{\jri}[1]{\textcolor{blue}{ \emph{\scriptsize  #1}} }
\definecolor{gerkegreen} {rgb} {0,0.6,0}
\newcommand{\jpg}[1]{\textcolor{mattgreen}{ \em{\scriptsize  #1}} }

\usepackage[normalem]{ulem}
\def\dt{\bgroup
 \markoverwith{\lower-0.2ex\hbox
 {\kern-.03em\vbox{\hrule width.2em\kern0.45ex\hrule}\kern-.03em}}%
 \ULon}
\MakeRobust\dt
\usepackage[normalem]{ulem}
\def\dt{\bgroup
 \markoverwith{\lower-0.2ex\hbox
 {\kern-.03em\vbox{\hrule width.2em\kern0.45ex\hrule}\kern-.03em}}%
 \ULon}
\MakeRobust\dt
% ok latexml pay attention again
\fi

%%%%

\setcounter{footnote}{1}
%
%\title{The genomic impacts of drift and selection for hybrid performance in maize}
%Corresponding author:  Department of Plant Sciences, University of California, Davis, California 95616, USA. 
%    E-mail: \mbox{rossibarra@ucdavis.edu}}\\[0.3cm]
%   \small\sf $^{\ast}$Division of Plant Sciences, University of Missouri, Columbia MO 65211,\\
%   \small\sf $^\dag$Corn Insects and Crop Genetics Research Unit, USDA-Agricultural Research Service, Ames, IA, 50011\\
%   \small\sf $^\ddag$Plant Genetics Research Unit, USDA-Agricultural Research Service, Columbia MO 65211,\\
%   \small\sf $^\S$Department of Plant Sciences, The Center for Population Biology and the Genome Center, University of California, Davis, California 95616 , USA,\\
%   \small\sf $^1$ Present address:  DuPont Pioneer, Johnston, Iowa
%}
   %\small\sf  Present address: Graduate University for Advanced Studies, Hayama, Kanagawa 240-0193, Japan\\
%}

\title{Independent molecular basis of convergent highland adaptation in maize}
\author{
 \small\sfbf{Justin P. Gerke$^{\ast,1}$\thanks{
Corresponding author:  DuPont Pioneer, 8305 NW 62ND Ave, P.O. Box 7060
Johnston, IA, 50131   E-mail: \mbox{justin.gerke@gmail.com}}, Jode W. Edwards$^{\dag}$, Katherine E. Guill$^{\ddag}$, Jeffrey Ross-Ibarra$^{\S}$}\thanks{
Corresponding author:  Department of Plant Sciences, University of California, Davis, California 95616. 
    E-mail: \mbox{rossibarra@ucdavis.edu}},\\ 
\small\sfbf{and Michael D. McMullen$^{\ast,\ddag}$}\\[0.3cm]
   \small\sf $^{\ast}$Division of Plant Sciences, University of Missouri, Columbia MO 65211\\
   \small\sf $^\dag$Corn Insects and Crop Genetics Research Unit, USDA-Agricultural Research Service, Ames, IA, 50011\\
   \small\sf $^\ddag$Plant Genetics Research Unit, USDA-Agricultural Research Service, Columbia MO 65211\\
   \small\sf $^\S$Department of Plant Sciences, Center for Population Biology, and Genome Center, University of California, Davis, CA 95616  
}

%\date{Revised manuscript for \emph{Genetics}, \today}

\abstract{
 Modern maize breeding relies upon selection in inbreeding populations to improve the performance of cross-population hybrids. 
The United States Department of Agriculture - Agricultural Research Service (USDA-ARS) reciprocal recurrent selection experiment between the Iowa Stiff Stalk Synthetic (BSSS) and the Iowa Corn Borer Synthetic No. 1 (BSCB1) populations represents one of the longest standing models of selection for hybrid performance. 
To investigate the genomic impact of this selection program, we used the Illumina MaizeSNP50 high-density SNP array to determine genotypes of progenitor lines and over 600 individuals across multiple cycles of selection. 
Consistent with previous research (Messmer et al., 1991; Labate et al., 1997; Hagdorn et al., 2003; Hinze et al., 2005), we found that genetic diversity within each population steadily decreases, with a corresponding increase in population structure. 
High marker density also enabled the first view of haplotype ancestry, fixation and recombination within this historic maize experiment. 
Extensive regions of haplotype fixation within each population are visible in the pericentromeric regions, where large blocks trace back to single founder inbreds. 
Simulation attributes most of the observed reduction in genetic diversity to genetic drift. 
Signatures of selection were difficult to observe in the background of this strong genetic drift, but heterozygosity in each population has fallen more than expected. 
Regions of haplotype fixation represent the most likely targets of selection, but as observed in other germplasm selected for hybrid performance (Feng et al., 2006), there is no overlap between the most likely targets of selection in the two populations. 
We discuss how this pattern is likely to occur during selection for hybrid performance, and how it poses challenges for dissecting the impacts of modern breeding and selection on the maize genome. }

\usepackage{natbib}
\bibpunct{(}{)}{;}{a}{}{,}

\usepackage{amsmath}

\usepackage{graphicx}

\begin{document}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRO
\input{intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRO

\input{methods}

\input{results}

\input{discussion}

\begin{acknowledgments}
We thank O. S. Smith and members of the Ross-Ibarra lab for comments on earlier versions of the manuscript. J.P.G received support for this research as a Merck Fellow of the Life Sciences Research Foundation. This research was supported by the National Science Foundation (IOS-0820619) and funds provided to USDA-ARS (MDM). Names of products are necessary to report factually on available data: however, neither the USDA, nor any other participating institution guarantees or warrants the standard of the product and the use of the name does not imply approval of the product to the exclusion of others that may also be suitable.
\end{acknowledgments}

\bibliography{references/references.bib}
\bibliographystyle{geneticsT2}

%\suppl

%%%%%%%
% put supplemental figures and tables here:
% e.g.
% \begin{figure}
% ...

%%%%%
% appendices to follow
%\input{HighLowSup1}

\end{document}

### 2015-04-14 results.tex
first commit of files and text

\section*{Results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
\begin{figure}[tb]   
  \begin{center}
   \vspace{-0mm}
   \includegraphics[width=0.4\textwidth]{fig1}
   \renewcommand{\baselinestretch}{0.9}
   \vspace{-3mm}
   \caption{BLAH} 
\vspace{-6mm}
    \label{fig:pca}
  \end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
\begin{figure}[tb]   
  \begin{center}
   \vspace{-0mm}
   \includegraphics[width=0.4\textwidth]{fig2}
   \renewcommand{\baselinestretch}{0.9}
   \vspace{-3mm}
   \caption{BLAH} 
\vspace{-6mm}
    \label{fig:decline}
  \end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
\begin{figure}[tb]   
  \begin{center}
   \vspace{-0mm}
   \includegraphics[width=0.4\textwidth]{fig3}
   \renewcommand{\baselinestretch}{0.9}
   \vspace{-3mm}
   \caption{BLAH} 
\vspace{-6mm}
    \label{fig:heterotic}
  \end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
\begin{figure}[tb]   
  \begin{center}
   \vspace{-0mm}
   \includegraphics[width=0.4\textwidth]{fig4}
   \renewcommand{\baselinestretch}{0.9}
   \vspace{-3mm}
   \caption{BLAH} 
\vspace{-6mm}
    \label{fig:genphys}
  \end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
\begin{figure}[tb]   
  \begin{center}
   \vspace{-0mm}
   \includegraphics[width=0.4\textwidth]{fig5}
   \renewcommand{\baselinestretch}{0.9}
   \vspace{-3mm}
   \caption{BLAH} 
\vspace{-6mm}
    \label{fig:genphys2}
  \end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE

### 2015-04-15 discussion.tex
everything up, now on to formatting and editing

\section*{Discussion}
Reciprocal recurrent selection serves as a model for the method of hybrid maize improvement employed broadly in North America (Duvick et al., 2004). In the Iowa RRS experiment, the BSSS and BSCB1 populations steadily lost genetic diversity as they become more diverged from one another. Principal component analysis shows that as the effective population size and the rates of inbreeding were altered, the rates of change in population structure were altered as well. The population structure and divergence between BSSS and BSCB1 can be largely reproduced by simulation without any selection. The trends in these populations support the hypothesis that the majority of the genetic structure observed can be attributed to genetic drift alone, despite effective selection for phenotypic improvement. This drift was driven by inbreeding through self-pollination and the small effective population sizes used to select a limited number of high-performing, potentially related individuals at each cycle.  The trends observed in the Iowa RRS are also a common theme in similar recurrent selection experiments (Romay et al., 2012; C. Lamkey and A. Lorenz, personal communication). Increased population structure due to increased inbreeding and lowered effective population size are visible in the modern ‘stiff-stalk’ and ‘non stiff-stalk’ heterotic groups as well (van Heerwaarden et al., 2012). Given the lack of strong selection signals in each case, genetic drift has most likely played a role in the current genetic structure of modern maize. 
	Several key inbreds in the ‘stiff-stalk’ heterotic group, B73, B37 and B14, were derived from the BSSS population (Darrah and Zuber, 1986; Troyer, 1999). B37 and B14 were derived from cycle 0, and B73 was derived from a half-sib recurrent selection program also started with the BSSS population. We examined these three inbreds at the pericentromeric regions listed in Table 1, and found that in most cases they carry different haplotypes from those that rose to high frequency in the RRS experiment.  Near isogenic lines that substitute these centromeric haplotypes from key inbreds could be used to determine if the haplotypes at high frequency in cycle 16 of BSSS provide a phenotypic advantage when crossed with BSCB1. 
	Both our study and previous work (Labate et al., 1997) identified genetic differentiation between cycle 0 of BSSS and the founder inbreds. The loss of diversity in the cycle 0 population likely led to the loss of some rare alleles and haplotypes. BSSS was maintained for several years before RRS was initiated, and the differentiation could have occurred due to low effective population size during maintenance. If so, then this drift will have impacted the trajectory of the RRS program and of the key inbreds derived from BSSS. However, cycle 0 has also undergone maintenance since the beginning of the RRS experiment, so some genetic drift may have occurred during this time as well.
	Although drift can explain most of the genetic structure genome-wide, phenotypic data provide clear evidence that selection has altered the frequencies of favorable alleles in the BSSS and BSCB1 populations. Numerous experiments have shown that the selected populations and the hybrids formed from them are superior to their cycle 0 ancestors (Smith, 1983; Keeratinijakal and Lamkey, 1993; Schnicker and Lamkey, 1993; Holthaus and Lamkey, 1995). Improvement occurred not only for hybrid yield, but concurrently for plant architecture and tolerance to high-density planting in both hybrids and inbreds (Brekke et al., 2011a; Brekke et al., 2011b; Edwards, 2011; Lauer et al., 2012). Since these same trends are observed across wider North American germplasm (Tollenaar, 1989; Duvick, 2005), identification and characterization of the loci conferring phenotypic improvement in BSSS and BSCB1 could reveal some of the genetic mechanisms by which maize yield has steadily improved over the decades. We find that the genomic regions of extremely low diversity evident at cycle 16 are unlikely to be produced by simulation, and heterozygosity falls more than expected across the genome as a whole.  However, simulated and observed values are often close, so overall the signature of selection is difficult to detect at any given locus. This is due in large part to a limitation of statistical power imposed by inbreeding and low effective population sizes, though uncertainty about the fine-scale recombination map also plays a role. We show that an identity-by-descent, haplotype based approach provides additional power as it can distinguish between the fixation of rare and common haplotypes. However, the results of single locus simulations are sensitive to sampling error and drift caused by population maintenance. In this case, it is difficult to assess significance with only a single population to test. This type of analysis would be more effective across several replicated populations, which can control for genetic drift due to the independence between the selections in each replicate (C. Lamkey and A. Lorenz, personal communication). 
	van Heerwaarden et al., 2012 assessed variation and tested for selection across a wider range of North American maize germplasm using the same SNP array. Although selected loci were identified in that study, the overall effects on genome diversity and haplotype patterns were relatively small compared with the impact of domestication. That study was a wide survey across many germplasm sources. The limited impact of selection could arise because different haplotypes and loci are selected in different breeding programs. In support of this idea, we find that the most likely targets of selection between the BSSS and BSCB1 populations occur at non-overlapping loci. This non-overlap between putatively selected loci in complementary populations has also been observed in commercial breeding programs (Feng et al., 2006).
The observation that the same targets of selection are not observed in the opposing heterotic populations bears implications for the genetic mechanisms responsible for heterosis and the success of maize hybrids. Classic overdominance models of heterosis predict that at a single locus, two distinct alleles confer heterozygote advantage when combined, while the dominance model predicts that heterosis is driven by dominance effects and the complementation of linked alleles in low-recombination regions (dominance or pseudo-overdominance). In the case of true overdominance, we expect selection should lead to decreased heterozygosity at a locus in both populations as complementary haplotypes are fixed in each group. We find no evidence of this genetic phenomenon. The observed pattern instead favors a dominance model, where fixation of a haplotype in one population simply selects against that same haplotype in the other population. Because most deleterious alleles are rare in both heterotic groups (S. Mezmouk and J. Ross-Ibarra, unpublished results), most haplotypes in the second population will have a different suite of deleterious variants and will complement the fixed haplotype reasonably well such that selection will have little impact on diversity in the second population. Our data are consistent with such a process having been important for hybrid improvement in the Iowa RSS experiment, and the lack of extreme differentiation seen over time across US Corn Belt germplasm is consistent with a role for such selection in maize hybrid improvement in general (van Heerwaarden et al., 2012).
	 Our results are consistent with the dominance and pseudo-overdominance models, but we caution that the exact outcome of particular models will depend strongly on the effects of selection, drift, and the frequencies of beneficial alleles. Disentangling these factors remains a major challenge that will require careful simulation of maize breeding using a wider range of parameters. This is especially true because in a model of hybrid complementation, genetic drift in one population can alter the selective value of alleles in the other population. When drift and selection interact in such a manner, tests for selection that attempt to independently partition the effects of each force may not provide the full picture. Furthermore, uncertainties in the fine-scale relationship between genetic and physical distance make it difficult to assign significance in forward-simulation approaches. In the end, the best test for selection of specific genomic regions will ultimately be conducted by phenotypic observation in the field in balanced tests of different haplotypes. In the case of individual breeding programs, genome-wide genotyping such as we conducted here can identify the lines carrying the recombinant haplotypes and introgressions necessary to conduct such an experiment.

### 2015-04-15 methods.tex
everything up, now on to formatting and editing

\section*{Methods}
\subsection*{The BSSS and BSCB1 Recurrent Selection Program}
The Iowa RRS experiment began with founder inbred lines (Table S1) that were randomly mated to create the BSSS and BSCB1 ‘cycle 0’ base populations. 
Each group was then recurrently selected for improved combining ability with the other group (Penny and Eberhart, 1971; Keeratinijakal and Lamkey, 1993). 
Initially approximately 200-250 starting ‘S0’ plants within each population were simultaneously self-fertilized to generate ‘S1’ lines and crossed to a random sample (4-6) of plants from the other population to generate testcross seed. 
The material was evaluated for highly heritable traits including standability, disease resistance and ear rot resistance in the nursery at the time the testcross seed was made, and 100 testcrosses were then grown in replicated field trials and evaluated for grain yield, grain moisture, and standability.  
S1 seed of 10 selected families within each population were randomly mated to create a new ‘cycle n+1’ population. 
After cycle 4, the testcrosses were carried out on the S1 lines themselves rather than at S0, which leads to another round of selfing prior to the creation of the next cycle. Ten lines (out of 100) were selected and advanced to form the next cycle between cycles 1 and 8, and twenty lines were selected between cycles 9 and 16 (Keeratinijakal and Lamkey, 1993). 
Founder inbreds and samples from cycles 0, 4, 8, 12, and 16 were genotyped at 39,258 SNPs that passed a set of quality filters and could be assigned collinear genetic and physical map positions.

\subsection*{Plants and inbred lines used}
The plants and inbred lines used in this experiment are listed in Table S1. 
We genotyped 36 plants from each of the BSSS and BSCB1 populations at selection cycles 0, 4, 8, 12, and 16. These plants represent descendants of the original populations which have been randomly mated to maintain seed. 
We also genotyped the founder inbreds for each population, however, seed for the founder lines F1B1, CI.617, WD456, and K230 were not available. 
The data for founder line CI.540 was not used because the genotyped material was heterozygous. 
A number of derived lines were also genotyped for calibrating phasing and imputation procedures (see below).

\subsection*{Genotype data}
 Plants from the cycles of selection, founders, and derived lines were grown in a greenhouse and tissue was collected at the 3 leaf stage. Tissue was lyophilized, ground, and DNA extracted by a CTAB procedure (Saghai-Maroof et al., 1984). 
Samples were genotyped using the 24-sample Illumina MaizeSNP50 array (Ganal et al., 2011) according to the Illumina Infinium protocol, and imaged on an Illumina BeadStation at the University of Missouri DNA core facility. Genotypes were determined with the GenomeStudio v2010.2 software using the manufacturer's MaizeSNP50\_B.egt cluster file. \jri{this is an edit because url listed in pages doc no longer valid. correct that this is not a custom .egt file?}
The design of the maize SNP50 Chip included a relatively small ascertainment panel of inbred lines, introducing a bias in the frequencies of SNPs included on the chip (Ganal et al., 2011). 
However, because our simulations are based not on theoretical expectations but instead on sampling from the observed data at cycle 0, we expect ascertainment bias to have a minimal impact on our results. The effect of this ascertainment bias was shown to be minimal in another study of North American germplasm (van Heerwaarden et al., 2012).

48,919 SNPs were called on the Illumina platform from the MaizeSNP50_B.egt cluster file. Genotypes with quality scores of 50 or less were recoded as missing data. Three plants were removed from the data due to an excess of missing data (the derived line B10, a cycle 0 plant from BSSS, and a cycle 8 plant from BSSS). 
In addition, BSCB1 plant 31 from cycle 4 appeared switched with plant 31 from cycle 8 based on our principal component analysis (PCA), so we switched the labels for these two genotypes to correct the mistake. 
To avoid structure among the missing data, we removed any SNP that was coded as missing in more than 3 plants in either group of founders or any group of plants from a particular cycle and population. Preliminary analysis by PCA and heatmap plots of distance matrices revealed two additional likely mix-ups. 
Plant 23 from BSSS cycle 8 was a clear outlier from the BSSS population as a whole and plant 2 from BSSS cycle 0 is likely a mislabeled plant from cycle 16. Since there was no evidence suggesting when mis-labelings occurred, each of these plants was removed from the analysis.

Integrating the genetic and physical map:  The interpretation of our results depends upon a genetic and physical map that is as accurate as possible. We therefore took steps to improve the positions of the SNP markers on the genetic and physical maps relative to version 5A.59 of the maize genome assembly (maizesequence.org). 
The probable physical position of each SNP based was obtained by comparing SNP context sequences to the genome sequence. For this purpose, SNP context sequences were defined as the sequence 25 bp upstream of the SNP, the bp representing the SNP itself, followed by 25 bp downstream of the SNP, making a total sequence length of 51bp. 
When a single genomic location was queried by two separate probes on the array, we chose the probe with higher quality calls and dropped the other marker from the dataset. 
To assign a genetic position for each SNP, we used a map derived from the B73xMo17 (IBM) mapping population similar to the IBM framework map in Ganal et al. (Ganal et al., 2011). 
This genetic map contains 4,217 framework SNP markers, which provides a much higher density than the map used to order the 5A.59 release of the maize genome sequence. As a result, we identified several places in the genome where the physical positions were incorrect according to our genetic map. 
These cases included both simple reversals of the physical map relative to the genetic map, and also the assignment of blocks of markers to the wrong linkage group, which we refer to as mis-mapped blocks. SNPs at these loci were reordered to match our genetic map. To maintain collinearity between the genetic and physical map, the physical positions of these SNPs were reassigned as follows. Individual mis-mapped markers were simply removed from the data. 
We also removed small reversals and mis-mapped blocks (<10 kb). 
Small rearrangements of this sort are more likely to represent mis-mapped paralagous sequence than true errors in the physical map. 
When larger reversals were identified, we transposed the physical positions of the SNPs from one end of the segment to the other. 
Mis-mapped blocks were often larger than the physical gap into which they were moved. 
We therefore assigned the first SNP of the block to a position 10 kb downstream from the previous SNP on the correct linkage group. We then recalculated genomic coordinates for the rest of the chromosome based on the marker distances within the translocated segment. The last SNP of the block was also given a 10 kb cushion between itself and the next SNP on the correct linkage group.

Only a portion of the SNPs on the array had genetically mapped positions. ‘Unmapped’ SNPs (those with a physical position but no genetic position) bounded by two mapped markers were moved along with their mapped neighbors if both anchoring mapped markers were also moved. However, it is unclear whether unmapped SNPs just outside of these anchors should be kept in place or moved along with the adjacent SNPs. Since most inversions were small relative to the genetic map (and would therefore still fall in the same window of a sliding window analysis), these SNPs were left in place. However, markers bordering translocations were removed to ensure there were no markers mapped to the incorrect linkage group. Unmapped SNPs were then assigned a genetic position by linear interpolation of genetic vs. physical distance using the approx() function in R (www.R-project.org). 
The IBM genetic map distances were then converted to single-meiosis map distances using the formulae of Winkler et al. (Winkler et al., 2003). 
Finally, SNPs located at physical positions outside of those bounded by the genetic map (such as the telomeres) were assigned the genetic position of their nearest mapped neighbor. Since moved segments were arbitrarily joined 10 kb from their nearest genetic neighbor, we acknowledge that the physical positions of these markers are only estimates. However, the estimated junctions are small relative to the genetic windows used for our analysis. 
The final map used is provided as supplemental material (Supplementary File S1).

\subsection*{Haplotype phasing}
 Although the genotypes of the plants from each population are unphased, the homozygous genotypes of the founders and derived inbreds provide excellent prior information for a probabilistic estimation of genotype phase in the populations. We therefore used fastPHASE (Scheet and Stephens, 2006) to estimate the genotype phase of each plant. To estimate the error in phasing, we created test cases by combining the genotypes of two derived inbreds into a hypothetical “F1 hybrid” of unknown phase. This F1 was presented to fastPHASE with the rest of the data, except that its parent inbreds were removed. Analyses of several hypothetical F1’s from different cycles of selection revealed very low phasing error rates (Table S3). Therefore the phased genotypes of cycle 0 plants were used as the starting data for simulations (see below).

\subsection*{Diversity and principal component analysis}
 Heterozygosity (H) was measured as:

\begin{align*}
H = 2p(1-p)
\end{align*}

where p and (1-p) are the frequencies of the two SNP alleles. FST (Hudson et al., 1992) was calculated using the HBKpermute program in the analysis package of the software library libsequence (Thornton, 2003). 
All results were plotted using the R package ggplot2 (Wickham, 2009). 
We conducted PCA by singular value decomposition, as described in (McVean, 2009).

\subsection*{Simulations}  
Our simulation sought to model the effects of genetic drift in the Iowa RRS experiment independent of any selection, and our model thus closely followed the published methods of the Iowa RRS (Penny and Eberhart, 1971; Keeratinijakal and Lamkey, 1993). 
Starting individuals in each population were constructed by randomly sampling two distinct haplotypes with replacement from the phased haplotypes of cycle 0. 
In the actual random mating scheme used in the Iowa RRS experiment, a single pairing could only contribute four gametes to the next generation (two kernels each from two ears), and our simulation reflects this. 
Advanced cycles were simulated by randomly mating gametes from self-fertilized plants of the previous cycle until 10 new individuals were created. 
The first cycle involved two rounds of random mating, whereas all subsequent cycles used one round. 
After cycle 5, the process employed two rounds of selfing instead of one. 
After cycle 7, the population size was increased from 10 to 20. 
At cycles 4, 8, 12, and 16, the plants were randomly mated to match the sample size of the observed data. 
The genotypes of these simulated random matings are the final results of each simulation that were analyzed in the same way as the observed data.

All recombination events in the simulations were carried out in R with the hypred software package (cran.r-project.org/web/packages/hypred/). The simulations were executed in parallel on a computing cluster, with unique random number seeds drawn for each simulation. Statistics were calculated for each simulation using the same formula as with the experimental data. We used non-overlapping sliding windows of equal genetic distance to account for the non-independence of markers in low-recombination regions when calculating measures of significance. For the haplotype-based, single-locus simulations, recombination was simply replaced with binomial sampling of two alleles.

### 2015-04-15 results.tex
everything up, now on to formatting and editing

@@
 \section*{Results}
+Increase in population structure and loss of genetic diversity:  Founder inbreds and samples from cycles 0, 4, 8, 12, and 16 were genotyped at 39,258 SNPs that passed a set of quality filters and could be assigned collinear genetic and physical map positions (See Materials and Methods for details).
+Change in population structure throughout the Iowa RRS experiment can be observed visually by a principal component analysis (PCA).
+A joint analysis of individuals from all the selection cycles (Figure \ref{fig:pca}) clearly separates the BSSS and BSCB1 populations along the first axis of variation, with increasing separation as the experiment progressed.
+The second axis of variation primarily separates the cycles from one another within each population.
+There is no separation between the founders of the two populations, but at cycle 0, the BSSS population shows more divergence from the founders than does BSCB1.
+This could be due to drift during either the population’s construction or its subsequent maintenance.
+Structure continued to develop within each population over the course of the experiment.
+There is an especially wide gap between cycles 4 and 8, which correlates with the addition of an extra generation of self-pollination prior to selection at each cycle.
+The distance between cycles then decreases dramatically after cycle 8, and this corresponds to doubling in effective population size.
 
@@
 
+No new genetic material was intentionally introduced into either population after the experiments’ inception, so the substantial increase in genetic distance could only arise from the loss of genetic diversity within each population.
+Consistent with previous studies of the Iowa RRS, genome-wide genetic diversity decreased steadily across cycles of selection in both populations (Figure \ref{fig:decline}).
+The decrease in heterozygosity (H) was much smaller when the two populations are combined, indicating the loss of different alleles within BSCB1 and BSSS.
+This leads to strong genetic differentiation, as FST rose by a factor of ten between the founder lines and the populations at cycle 16.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
+
+We noticed an irregular increase in the number of polymorphic markers between BSSS cycles 4 and 8 (Table S2).
+All of these newly polymorphic markers were present at extremely low frequency and were spread among various individuals.
+This may represent a series of minor alleles that were not captured in our sample of cycle 4 individuals and thus appeared to ‘resurface’ at cycle 8.
+Alternatively, the pattern may be the result of minor contamination at some point in the population’s history.
+It was observed that an allele of the sugary gene associated with sweet corn appeared in the population at this time (O.
+S.
+Smith, personal communication), suggesting contamination may be the cause.
+However, the low frequency of the ‘new’ alleles (typically only one or two alleles out of 72 possible in 36 diploid samples) means their effect on the population diversity is minimal.
+We did not attempt to incorporate this contamination into our simulation approaches, as it only makes our tests for low heterozygosity slightly more conservative.
+	Fixation of large genomic regions:  Figure 3 shows heterozygosity varying along the genome across cycle 16 of each population.
+Of particular note in this figure are extremely large regions of zero or near-zero heterozygosity spanning tens of megabases that lie in the pericentromeric regions.
+These regions experience low rates of meiotic recombination, which creates an expanded physical map relative to their genetic length (Ganal et al., 2011).
+In general, the majority of fixed haplotype segments are small in genetic space (most are less than 2 cM) regardless of their physical size.
+One exception is a region of fixation that spans 7 cM on chromosome 1 in the BSCB1 population.
+The sheer physical size of the pericentromeric regions yields extremely high marker density on the genetic map, allowing for clear resolution of haplotype phasing and recombination breakpoints.
+To further examine the fixation in these regions, we computationally imputed haplotype phase in the BSSS and BSCB1 populations and used the phased data to track haplotype frequencies and founder of origin.
+In most cases, these fixed haplotypes can be traced back to single founder inbreds.
+For example, in BSSS a 60 Mb, 1.2 cM region of chromosome 9 became fixed by cycle 12 (Figure 4) and traces back to the founder Os420.
+In BSCB1, a 60 Mb, 0.7 cM region became fixed on chromosome 4 and traces back to A340 (Figure 5).
+Table 1 gives a summary of the large genomic regions that have become fixed or nearly fixed by cycle 16.
+These regions represent large blocks of linked alleles that show no evidence of recombination since at least the development of the founding inbred lines in the 1920’s and 1930’s.
+	The role of genetic drift:  There is clear evidence of phenotypic improvement in response to selection in the Iowa RRS populations (Smith, 1983; Keeratinijakal and Lamkey, 1993; Schnicker and Lamkey, 1993; Holthaus and Lamkey, 1995; Brekke et al., 2011a; Brekke et al., 2011b; Edwards, 2011) and large changes in genetic structure indicated by molecular markers.
+A central issue for these maize populations and others like them is whether the changes observed at the molecular level are caused directly by selection on phenotype, or indirectly due to the genetic drift that selection imposes through inbreeding and small effective population sizes.
+To gauge the roles of selection and drift, we conducted simulations of the crossing and selection schemes used in the RRS experiment, with genetic recombination modeled on a framework map of a cross between B73 (a BSSS derived line used to construct the reference genome) with Mo17 (an inbred with some co-ancestry to both the BSSS and BSCB1 founder germplasm)(Gerdes et al., 1993; Lee et al., 2002).
+Selections were executed at random in each simulation, so the patterns observed across simulations represent the expected distribution of effects caused by recombination and genetic drift.
+We conducted 10,000 simulations, sampled individuals from each cycle equal in number to the observed samples, measured heterozygosity, and compared the results to the observed data.
+	Averaged across the genome, the vast majority of the reduction in diversity observed in both populations can be attributed to genetic drift.
+However we do observe deviations from the simulated values (Figure S1).
+The observed data show higher than expected heterozygosity at cycle 4, suggesting selection for heterozygosity in early cycles, a deviation from simulated and actual breeding practices, or an under-sampling of the diversity present in the original cycle 0 populations.
+Nevertheless, as the cycles progress heterozygosity falls more than expected in both populations and the observed values at cycle 16 are significantly lower than the simulated data.
+Simulations across a number of different marker densities were consistent with this result (data not shown).
+ To examine the behavior of specific regions, we also calculated the observed and simulated results for each 2 cM segment of the genome.
+The dynamics observed across most of the genome are largely insensitive to changes in window size (we tested from 2-4 cM, data not shown), and are consistent with strong genetic drift imposed by selection practices.
+A subset of loci were flagged as significant, and these loci almost always overlapped regions of fixation or near-zero heterozygosity in one population.
+However, in these regions the values obtained by simulation are often quite low, such that the extreme simulated values sometimes lie only a few percentage points away from the observed value (Supplementary File S2), indicating that drift alone can explain most of the drop in diversity.
+Since the population size of the Iowa RRS is small (10-20), many biallelic SNPs should fix by chance regardless of their starting minor allele frequencies.
+The deviation from simulation arises not from changes in allele frequencies per se, but rather from the fixation of linked markers across larger than expected genetic distances.
+The validity of significance cutoffs therefore depend on the accuracy of our genetic map, and the maize genetic map is known to vary among genetic backgrounds across short genetic distances (< 5cM) (McMullen et al., 2009).
+Deviations of the observed and simulated results could be due to selection, variation / inaccuracy in the genetic map, or a combination of these factors.
+To explore the roles of selection and drift independent of the genetic map, we returned to the large regions of fixation in the centromeres which showed no recombination across the full RRS experiment.
+Given the lack of recombination, each of these regions can be analyzed by the simulation of a single locus, and the high density of markers allows the clear resolution of the individual haplotypes.
+We used the computationally phased data to measure the frequency of the fixed haplotype at each cycle, and assessed the probability of observing the fixation event given the initial frequency.
+The BSSS chromosome 9 haplotype fixed at cycle 16 was at low frequency at cycle 0 (7 out of 68 haplotypes), but increased rapidly in frequency by cycle 8 (66 out of 70 haplotypes).
+Simulation of the haplotype as a single locus in the RRS experiment produces this increase in frequency in only 3.9\% of 1000 independent simulations, whereas the haplotype was lost in over 80\% of the simulations.
+In BSCB1, a 30 Mb, 1.6 cM region of chromosome 2 became nearly fixed by cycle 8 (67/70) despite a prevalence of 4/72 at cycle 0, which occurred 1.5\% of the time by simulation.
+Although these results suggest that selection may have pushed these haplotypes to fixation, the fact that fixation of such a rare haplotype still occurred in some simulations speaks to the strong genetic drift imposed upon the BSSS and BSCB1 populations.
+Interestingly, each of these two genomic regions harbored a different cycle 0 haplotype at higher frequency, but these higher-frequency haplotypes were subsequently lost within the RRS population.
+In other cases, the haplotypes which eventually fixed were at moderate frequency in the cycle 0 populations and drift to fixation in the majority of simulations.
+
+%Table 1.  Ancestry of haplotypes fixed in the cycle 16 population. 
+%
+%
+%Population	Chromosome	"Genetic Interval
+%(cM)"	"Physical Interval
+%(Mb)"	Founder Inbred 	Derived Lines with Haplotype
+%BSSS	3	40.2-41.4	67.7-123	CI187-2	B94
+%BSSS	3	43-48.5	129.2-157.1	NDa	B89, B94
+%BSSS	4	39.7-42.1	39.9-82.7	CI187-2	B89, B94, B67, B72, B39, B43
+%BSSS	9	30.6-33.5	20.8-26.6	Oh7b	B89, B94, B43, B17, B72, B84, B67
+%BSSS	9	34.3-35.6c	30.8-90.4c	Os420	B89, B94
+%BSCB1	2	50.2-51.8	80.6-114.5	CC5	B90, B91, B95, B97, B99
+%BSCB1	4	42.1-42.8	82.7-140	NDa	B90, B95, B97
+%BSCB1	8	46.1-50.7	125.1-145.6	P8	B90, B97, B91, B99, B54
+%
+%a ND = Not Determined (Either a recombinant haplotype or originates from an un-genotyped founder)
+%
+%b Although Oh7 is a BSCB1 founder, it is a descendant of CI.540, an un-genotyped BSSS founder. BSSS segments matching Oh7 presumably derive from CI.540
+%
+%c  Founders Ind_B2 (BSSS), CI187-2 (BSSS), R4 (BSCB1), and I205 (BSCB1) are all IBD at this region of chromosome 9
 

### 2015-04-19 methods.tex
text, some refs, messed up biblio

@@
 \section*{Methods}
-\subsection*{The BSSS and BSCB1 Recurrent Selection Program}
+\subsection*{The BSSS and BSCB1 recurrent selection program}
 The Iowa RRS experiment began with founder inbred lines (Table S1) that were randomly mated to create the BSSS and BSCB1 ‘cycle 0’ base populations. 
-Each group was then recurrently selected for improved combining ability with the other group (Penny and Eberhart, 1971; Keeratinijakal and Lamkey, 1993). 
+Each group was then recurrently selected for improved combining ability with the other group \citep{penny1971twenty,  keeratinijakal1993responses}. 
 Initially approximately 200-250 starting ‘S0’ plants within each population were simultaneously self-fertilized to generate ‘S1’ lines and crossed to a random sample (4-6) of plants from the other population to generate testcross seed.
@@
 After cycle 4, the testcrosses were carried out on the S1 lines themselves rather than at S0, which leads to another round of selfing prior to the creation of the next cycle.
-Ten lines (out of 100) were selected and advanced to form the next cycle between cycles 1 and 8, and twenty lines were selected between cycles 9 and 16 (Keeratinijakal and Lamkey, 1993).
+Ten lines (out of 100) were selected and advanced to form the next cycle between cycles 1 and 8, and twenty lines were selected between cycles 9 and 16 \citep{keeratinijakal1993responses}.
 Founder inbreds and samples from cycles 0, 4, 8, 12, and 16 were genotyped at 39,258 SNPs that passed a set of quality filters and could be assigned collinear genetic and physical map positions.
@@
 These plants represent descendants of the original populations which have been randomly mated to maintain seed. 
-We also genotyped the founder inbreds for each population, however, seed for the founder lines F1B1, CI.617, WD456, and K230 were not available. 
+We also genotyped the founder inbreds for each population, with the exception of lines F1B1, CI.617, WD456, and K230 for which seed was not available. 
 The data for founder line CI.540 was not used because the genotyped material was heterozygous. 
@@
 Plants from the cycles of selection, founders, and derived lines were grown in a greenhouse and tissue was collected at the 3 leaf stage. 
-Tissue was lyophilized, ground, and DNA extracted by a CTAB procedure (Saghai-Maroof et al., 1984).
-Samples were genotyped using the 24-sample Illumina MaizeSNP50 array (Ganal et al., 2011) according to the Illumina Infinium protocol, and imaged on an Illumina BeadStation at the University of Missouri DNA core facility.
+Tissue was lyophilized, ground, and DNA extracted by a CTAB procedure \citep{saghai-maroof1984ribosomal}. 
+Samples were genotyped using the 24-sample Illumina MaizeSNP50 array \citep{ganal2011a-large} according to the Illumina Infinium protocol, and imaged on an Illumina BeadStation at the University of Missouri DNA core facility.
 Genotypes were determined with the GenomeStudio v2010.2 software using the manufacturer's MaizeSNP50\_B.egt cluster file.
@@
 correct that this is not a custom .egt file?}
-The design of the maize SNP50 Chip included a relatively small ascertainment panel of inbred lines, introducing a bias in the frequencies of SNPs included on the chip (Ganal et al., 2011). 
+The design of the maize SNP50 Chip included a relatively small ascertainment panel of inbred lines, introducing a bias in the frequencies of SNPs included on the chip \citep{ganal2011a-large}. 
 However, because our simulations are based not on theoretical expectations but instead on sampling from the observed data at cycle 0, we expect ascertainment bias to have a minimal impact on our results.
-The effect of this ascertainment bias was shown to be minimal in another study of North American germplasm (van Heerwaarden et al., 2012).
+The effect of this ascertainment bias was shown to be minimal in another study of North American germplasm \citep{van2012historical}.
 
-48,919 SNPs were called on the Illumina platform from the MaizeSNP50_B.egt cluster file.
-Genotypes with quality scores of 50 or less were recoded as missing data.
+We called 48,919 SNPs on the Illumina platform from the MaizeSNP50_B.egt cluster file. Genotypes with quality scores of 50 or less were recoded as missing data. 
 Three plants were removed from the data due to an excess of missing data (the derived line B10, a cycle 0 plant from BSSS, and a cycle 8 plant from BSSS). 
@@
 
-Integrating the genetic and physical map:  The interpretation of our results depends upon a genetic and physical map that is as accurate as possible.
-We therefore took steps to improve the positions of the SNP markers on the genetic and physical maps relative to version 5A.59 of the maize genome assembly (maizesequence.org).
+\subsection*{Integrating the genetic and physical map}
+The interpretation of our results depends upon a genetic and physical map that is as accurate as possible. 
+We therefore took steps to improve the positions of the SNP markers on the genetic and physical maps relative to version 5A.59 of the maize genome assembly. 
 The probable physical position of each SNP based was obtained by comparing SNP context sequences to the genome sequence. 
@@
 When a single genomic location was queried by two separate probes on the array, we chose the probe with higher quality calls and dropped the other marker from the dataset. 
-To assign a genetic position for each SNP, we used a map derived from the B73xMo17 (IBM) mapping population similar to the IBM framework map in Ganal et al. (Ganal et al., 2011). 
+To assign a genetic position for each SNP, we used a map derived from the B73xMo17 (IBM) mapping population similar to the IBM framework map in \citet{ganal2011a-large}. 
 This genetic map contains 4,217 framework SNP markers, which provides a much higher density than the map used to order the 5A.59 release of the maize genome sequence. 
@@
 These cases included both simple reversals of the physical map relative to the genetic map, and also the assignment of blocks of markers to the wrong linkage group, which we refer to as mis-mapped blocks.
-SNPs at these loci were reordered to match our genetic map.
-To maintain collinearity between the genetic and physical map, the physical positions of these SNPs were reassigned as follows.
-Individual mis-mapped markers were simply removed from the data.
-We also removed small reversals and mis-mapped blocks (<10 kb). 
+SNPs at these loci were reordered to match our genetic map. To maintain collinearity between the genetic and physical map, the physical positions of these SNPs were reassigned as follows. 
+Individual mis-mapped markers small reversals and mis-mapped blocks (<10 kb) were removed from the data. 
 Small rearrangements of this sort are more likely to represent mis-mapped paralagous sequence than true errors in the physical map. 
@@
 However, markers bordering translocations were removed to ensure there were no markers mapped to the incorrect linkage group. 
-Unmapped SNPs were then assigned a genetic position by linear interpolation of genetic vs.
-physical distance using the approx() function in R (www.R-project.org).
-The IBM genetic map distances were then converted to single-meiosis map distances using the formulae of Winkler et al. (Winkler et al., 2003). 
+Unmapped SNPs were then assigned a genetic position by linear interpolation of genetic vs. physical distance using the approx() function in R \citep{rteam}. 
+The IBM genetic map distances were then converted to single-meiosis map distances using the formulae of \citet{winkler2003determination}. 
 Finally, SNPs located at physical positions outside of those bounded by the genetic map (such as the telomeres) were assigned the genetic position of their nearest mapped neighbor.
+\jri{so everything beyond 0 gets a 0 position on the map instead of negative?}
 Since moved segments were arbitrarily joined 10 kb from their nearest genetic neighbor, we acknowledge that the physical positions of these markers are only estimates. 
@@
 Although the genotypes of the plants from each population are unphased, the homozygous genotypes of the founders and derived inbreds provide excellent prior information for a probabilistic estimation of genotype phase in the populations.
-We therefore used fastPHASE (Scheet and Stephens, 2006) to estimate the genotype phase of each plant.
+We therefore used fastPHASE \citep{scheet2006fast} to estimate the genotype phase of each plant.
 To estimate the error in phasing, we created test cases by combining the genotypes of two derived inbreds into a hypothetical “F1 hybrid” of unknown phase.
@@
 
-where p and (1-p) are the frequencies of the two SNP alleles.
-FST (Hudson et al., 1992) was calculated using the HBKpermute program in the analysis package of the software library libsequence (Thornton, 2003).
-All results were plotted using the R package ggplot2 (Wickham, 2009). 
-We conducted PCA by singular value decomposition, as described in (McVean, 2009).
+where $p$ and $(1-p)$ are the frequencies of the two SNP alleles. 
+$F_{ST}$ \citep{hudson1992statistical} was calculated using the HBKpermute program in the analysis package (\url{https://github.com/molpopgen/analysis}) of the software library libsequence \citep{thornton2003libsequence}.
+All results were plotted using the R package ggplot2 \citep{wickham2009ggplot2}. 
+We conducted PCA by singular value decomposition, as described in \cite{mcvean2009genealogical}.
 
 \subsection*{Simulations}  
-Our simulation sought to model the effects of genetic drift in the Iowa RRS experiment independent of any selection, and our model thus closely followed the published methods of the Iowa RRS (Penny and Eberhart, 1971; Keeratinijakal and Lamkey, 1993).
+Our simulation sought to model the effects of genetic drift in the Iowa RRS experiment independent of any selection, and our model thus closely followed the published methods of the Iowa RRS \citep{penny1971twenty,  keeratinijakal1993responses}.
 Starting individuals in each population were constructed by randomly sampling two distinct haplotypes with replacement from the phased haplotypes of cycle 0. 
@@
 At cycles 4, 8, 12, and 16, the plants were randomly mated to match the sample size of the observed data. 
-The genotypes of these simulated random matings are the final results of each simulation that were analyzed in the same way as the observed data.
+The genotypes of these simulated random matings are the final results of each simulation and were analyzed in the same way as the observed data.
 
-All recombination events in the simulations were carried out in R with the hypred software package (cran.r-project.org/web/packages/hypred/).
-The simulations were executed in parallel on a computing cluster, with unique random number seeds drawn for each simulation.
-Statistics were calculated for each simulation using the same formula as with the experimental data.
+All recombination events in the simulations were carried out in R with the hypred software package (\url{cran.r-project.org/web/packages/hypred/}). 
+Simulations were executed in parallel on a computing cluster, with unique random number seeds drawn for each simulation. 
+Statistics were calculated for each simulation using the same formula as the experimental data. 
 We used non-overlapping sliding windows of equal genetic distance to account for the non-independence of markers in low-recombination regions when calculating measures of significance. 

### 2015-04-19 results.tex
text, some refs, messed up biblio

@@
 \section*{Results}
-Increase in population structure and loss of genetic diversity:  Founder inbreds and samples from cycles 0, 4, 8, 12, and 16 were genotyped at 39,258 SNPs that passed a set of quality filters and could be assigned collinear genetic and physical map positions (See Materials and Methods for details).
+
+\subsection*{Population structure and genetic diversity}
+\jri{shortened section title, this ok?}
+Founder inbreds and samples from cycles 0, 4, 8, 12, and 16 were genotyped at 39,258 SNPs that passed a set of quality filters and could be assigned collinear genetic and physical map positions (See Materials and Methods for details).
 Change in population structure throughout the Iowa RRS experiment can be observed visually by a principal component analysis (PCA). 
@@
 There is no separation between the founders of the two populations, but at cycle 0, the BSSS population shows more divergence from the founders than does BSCB1. 
-This could be due to drift during either the population’s construction or its subsequent maintenance.
+This could be due to drift during either the population’s construction or subsequent maintenance. 
 Structure continued to develop within each population over the course of the experiment. 
 There is an especially wide gap between cycles 4 and 8, which correlates with the addition of an extra generation of self-pollination prior to selection at each cycle. 
-The distance between cycles then decreases dramatically after cycle 8, and this corresponds to doubling in effective population size.
+The distance between cycles then decreases dramatically after cycle 8, and corresponding to the increased effective population size (See Materials and Methods).
 
@@
 
-No new genetic material was intentionally introduced into either population after the experiments’ inception, so the substantial increase in genetic distance could only arise from the loss of genetic diversity within each population.
-Consistent with previous studies of the Iowa RRS, genome-wide genetic diversity decreased steadily across cycles of selection in both populations (Figure \ref{fig:decline}).
-The decrease in heterozygosity (H) was much smaller when the two populations are combined, indicating the loss of different alleles within BSCB1 and BSSS.
-This leads to strong genetic differentiation, as FST rose by a factor of ten between the founder lines and the populations at cycle 16.
+No new genetic material was intentionally introduced into either population after the experiment's inception, so the substantial increase in genetic distance could only arise from the loss of genetic diversity within each population.
+Consistent with previous studies of the Iowa RRS, genome-wide genetic diversity (expected heterozygosity, H) decreases steadily across cycles of selection in both populations (Figure \ref{fig:decline}).
+\jri{need a citation here of "previous studies"?}
+The loss of heterozygosity is smaller when the two populations are considered together, indicating the loss of different alleles within BSCB1 and BSSS. 
+This genetic differentiation is reflected by the tenfold increase in $F_{ST}$ between the founder lines and the populations at cycle 16.
 
@@
 Alternatively, the pattern may be the result of minor contamination at some point in the population’s history. 
-It was observed that an allele of the sugary gene associated with sweet corn appeared in the population at this time (O.
-S.
-Smith, personal communication), suggesting contamination may be the cause.
-However, the low frequency of the ‘new’ alleles (typically only one or two alleles out of 72 possible in 36 diploid samples) means their effect on the population diversity is minimal.
+It was observed that an allele of the sugary gene associated with sweet corn appeared in the population at this time (O.S. Smith, personal communication), suggesting contamination may be the cause. 
+However, the low frequency of the ‘new’ alleles (typically only one or two alleles out of 72 possible in 36 diploid samples) means their effect on population diversity is minimal.  
 We did not attempt to incorporate this contamination into our simulation approaches, as it only makes our tests for low heterozygosity slightly more conservative.
-	Fixation of large genomic regions:  Figure 3 shows heterozygosity varying along the genome across cycle 16 of each population.
-Of particular note in this figure are extremely large regions of zero or near-zero heterozygosity spanning tens of megabases that lie in the pericentromeric regions.
-These regions experience low rates of meiotic recombination, which creates an expanded physical map relative to their genetic length (Ganal et al., 2011).
-In general, the majority of fixed haplotype segments are small in genetic space (most are less than 2 cM) regardless of their physical size.
-One exception is a region of fixation that spans 7 cM on chromosome 1 in the BSCB1 population.
-The sheer physical size of the pericentromeric regions yields extremely high marker density on the genetic map, allowing for clear resolution of haplotype phasing and recombination breakpoints.
-To further examine the fixation in these regions, we computationally imputed haplotype phase in the BSSS and BSCB1 populations and used the phased data to track haplotype frequencies and founder of origin.
-In most cases, these fixed haplotypes can be traced back to single founder inbreds.
-For example, in BSSS a 60 Mb, 1.2 cM region of chromosome 9 became fixed by cycle 12 (Figure 4) and traces back to the founder Os420.
-In BSCB1, a 60 Mb, 0.7 cM region became fixed on chromosome 4 and traces back to A340 (Figure 5).
-Table 1 gives a summary of the large genomic regions that have become fixed or nearly fixed by cycle 16.
-These regions represent large blocks of linked alleles that show no evidence of recombination since at least the development of the founding inbred lines in the 1920’s and 1930’s.
-	The role of genetic drift:  There is clear evidence of phenotypic improvement in response to selection in the Iowa RRS populations (Smith, 1983; Keeratinijakal and Lamkey, 1993; Schnicker and Lamkey, 1993; Holthaus and Lamkey, 1995; Brekke et al., 2011a; Brekke et al., 2011b; Edwards, 2011) and large changes in genetic structure indicated by molecular markers.
-A central issue for these maize populations and others like them is whether the changes observed at the molecular level are caused directly by selection on phenotype, or indirectly due to the genetic drift that selection imposes through inbreeding and small effective population sizes.
-To gauge the roles of selection and drift, we conducted simulations of the crossing and selection schemes used in the RRS experiment, with genetic recombination modeled on a framework map of a cross between B73 (a BSSS derived line used to construct the reference genome) with Mo17 (an inbred with some co-ancestry to both the BSSS and BSCB1 founder germplasm)(Gerdes et al., 1993; Lee et al., 2002).
-Selections were executed at random in each simulation, so the patterns observed across simulations represent the expected distribution of effects caused by recombination and genetic drift.
-We conducted 10,000 simulations, sampled individuals from each cycle equal in number to the observed samples, measured heterozygosity, and compared the results to the observed data.
-	Averaged across the genome, the vast majority of the reduction in diversity observed in both populations can be attributed to genetic drift.
-However we do observe deviations from the simulated values (Figure S1).
-The observed data show higher than expected heterozygosity at cycle 4, suggesting selection for heterozygosity in early cycles, a deviation from simulated and actual breeding practices, or an under-sampling of the diversity present in the original cycle 0 populations.
-Nevertheless, as the cycles progress heterozygosity falls more than expected in both populations and the observed values at cycle 16 are significantly lower than the simulated data.
-Simulations across a number of different marker densities were consistent with this result (data not shown).
- To examine the behavior of specific regions, we also calculated the observed and simulated results for each 2 cM segment of the genome.
-The dynamics observed across most of the genome are largely insensitive to changes in window size (we tested from 2-4 cM, data not shown), and are consistent with strong genetic drift imposed by selection practices.
-A subset of loci were flagged as significant, and these loci almost always overlapped regions of fixation or near-zero heterozygosity in one population.
-However, in these regions the values obtained by simulation are often quite low, such that the extreme simulated values sometimes lie only a few percentage points away from the observed value (Supplementary File S2), indicating that drift alone can explain most of the drop in diversity.
-Since the population size of the Iowa RRS is small (10-20), many biallelic SNPs should fix by chance regardless of their starting minor allele frequencies.
-The deviation from simulation arises not from changes in allele frequencies per se, but rather from the fixation of linked markers across larger than expected genetic distances.
-The validity of significance cutoffs therefore depend on the accuracy of our genetic map, and the maize genetic map is known to vary among genetic backgrounds across short genetic distances (< 5cM) (McMullen et al., 2009).
-Deviations of the observed and simulated results could be due to selection, variation / inaccuracy in the genetic map, or a combination of these factors.
-To explore the roles of selection and drift independent of the genetic map, we returned to the large regions of fixation in the centromeres which showed no recombination across the full RRS experiment.
-Given the lack of recombination, each of these regions can be analyzed by the simulation of a single locus, and the high density of markers allows the clear resolution of the individual haplotypes.
-We used the computationally phased data to measure the frequency of the fixed haplotype at each cycle, and assessed the probability of observing the fixation event given the initial frequency.
-The BSSS chromosome 9 haplotype fixed at cycle 16 was at low frequency at cycle 0 (7 out of 68 haplotypes), but increased rapidly in frequency by cycle 8 (66 out of 70 haplotypes).
-Simulation of the haplotype as a single locus in the RRS experiment produces this increase in frequency in only 3.9\% of 1000 independent simulations, whereas the haplotype was lost in over 80\% of the simulations.
-In BSCB1, a 30 Mb, 1.6 cM region of chromosome 2 became nearly fixed by cycle 8 (67/70) despite a prevalence of 4/72 at cycle 0, which occurred 1.5\% of the time by simulation.
-Although these results suggest that selection may have pushed these haplotypes to fixation, the fact that fixation of such a rare haplotype still occurred in some simulations speaks to the strong genetic drift imposed upon the BSSS and BSCB1 populations.
-Interestingly, each of these two genomic regions harbored a different cycle 0 haplotype at higher frequency, but these higher-frequency haplotypes were subsequently lost within the RRS population.
-In other cases, the haplotypes which eventually fixed were at moderate frequency in the cycle 0 populations and drift to fixation in the majority of simulations.
 
-%Table 1.  Ancestry of haplotypes fixed in the cycle 16 population. 
-%
-%
-%Population	Chromosome	"Genetic Interval
-%(cM)"	"Physical Interval
-%(Mb)"	Founder Inbred 	Derived Lines with Haplotype
-%BSSS	3	40.2-41.4	67.7-123	CI187-2	B94
-%BSSS	3	43-48.5	129.2-157.1	NDa	B89, B94
-%BSSS	4	39.7-42.1	39.9-82.7	CI187-2	B89, B94, B67, B72, B39, B43
-%BSSS	9	30.6-33.5	20.8-26.6	Oh7b	B89, B94, B43, B17, B72, B84, B67
-%BSSS	9	34.3-35.6c	30.8-90.4c	Os420	B89, B94
-%BSCB1	2	50.2-51.8	80.6-114.5	CC5	B90, B91, B95, B97, B99
-%BSCB1	4	42.1-42.8	82.7-140	NDa	B90, B95, B97
-%BSCB1	8	46.1-50.7	125.1-145.6	P8	B90, B97, B91, B99, B54
-%
-%a ND = Not Determined (Either a recombinant haplotype or originates from an un-genotyped founder)
-%
-%b Although Oh7 is a BSCB1 founder, it is a descendant of CI.540, an un-genotyped BSSS founder. BSSS segments matching Oh7 presumably derive from CI.540
-%
-%c  Founders Ind_B2 (BSSS), CI187-2 (BSSS), R4 (BSCB1), and I205 (BSCB1) are all IBD at this region of chromosome 9
+\subsection*{Fixation of large genomic regions}
+Figure \ref{fig:heterotic} shows heterozygosity varying along the genome across cycle 16 of each population. 
+Of particular note are extremely large pericentromeric regions of zero or near-zero heterozygosity spanning tens of megabases. 
+These regions experience low rates of meiotic recombination, which creates an expanded physical map relative to their genetic length \citep{ganal2011a-large}. 
+In general, the majority of fixed haplotype segments are small (<2 cM) in genetic space regardless of their physical size; one exception is 7 cM region on chromosome 1 in the BSCB1 population.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
+
+The sheer physical size of the pericentromeric regions yields extremely high marker density on the genetic map, allowing for clear resolution of haplotype phasing and recombination breakpoints. 
+To further examine the fixation in these regions, we computationally imputed haplotype phase in the BSSS and BSCB1 populations and used the phased data to track haplotype frequencies and founder of origin.
+In most cases, these fixed haplotypes can be traced back to single founder inbreds. 
+For example, in BSSS, a 60 Mb (1.2 cM) region of chromosome 9 became fixed by cycle 12 (Figure \ref{fig:genphys}) and traces back to the founder Os420. 
+In BSCB1, a 60 Mb (0.7 cM) region became fixed on chromosome 4 and traces back to A340 (Figure \ref{fig:genphys2}). 
+Table \ref{tab:fix} gives a summary of the large genomic regions that have become fixed or nearly fixed by cycle 16. 
+These regions represent blocks of linked loci that show no evidence of recombination since at least the development of the founding inbred lines in the 1920’s and 1930’s.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
+
+\subsection*{The role of genetic drift}
+There is clear evidence of phenotypic improvement in response to selection in the Iowa RRS populations (Smith, 1983; Keeratinijakal and Lamkey, 1993; Schnicker and Lamkey, 1993; Holthaus and Lamkey, 1995; Brekke et al., 2011a; Brekke et al., 2011b; Edwards, 2011) and large changes in genetic structure indicated by molecular markers.
+A central issue for these maize populations and others like them is whether the changes observed at the molecular level are caused directly by selection on phenotype, or indirectly due to the genetic drift that selection imposes through inbreeding and small effective population sizes.
+To gauge the roles of selection and drift, we conducted simulations of the crossing and selection schemes used in the RRS experiment, with genetic recombination modeled on a framework map of a cross between B73 (a BSSS derived line used to construct the reference genome) with Mo17 (an inbred with some co-ancestry to both the BSSS and BSCB1 founder germplasm)\citep{gerdes1993compilation}\citex{Lee et al., 2002}.
+Selections were executed at random in each simulation, so the patterns observed across simulations represent the expected distribution of effects caused only by recombination and genetic drift. 
+We conducted 10,000 simulations, sampled individuals from each cycle equal in number to the observed samples, measured heterozygosity, and compared the results to the observed data.
+
+Averaged across the genome, the vast majority of the reduction in diversity observed in both populations can be attributed to genetic drift. 
+Nonetheless, do observe deviations from the simulated values (Figure S1). 
+The observed data show higher than expected heterozygosity at cycle 4, suggesting selection for heterozygosity in early cycles, a deviation from simulated and actual breeding practices, or an under-sampling of the diversity present in the original cycle 0 populations.
+\jri{where is this excess of diversity?
+if it's all pericentromeric can we argue for the first hypothesis?}
+As the cycles progress, heterozygosity falls more rapidly than expected in both populations, and the observed values at cycle 16 are significantly lower than the simulated data.  
+Simulations across a number of different marker densities were consistent with this result (data not shown).
+
+To examine the behavior of specific regions, we also calculated the observed and simulated results for each 2 cM segment of the genome. 
+The dynamics observed across most of the genome are largely insensitive to changes in window size (we tested from 2-4 cM, data not shown), and are consistent with strong genetic drift imposed by selection practices.
+\jri{did we test smaller regions?
+not enough markers?}
+A subset of loci were flagged as significant, and these loci almost always overlapped regions of fixation or near-zero heterozygosity in one population. 
+However, in these regions the values obtained by simulation are often quite low, such that the extreme simulated values sometimes lie only a few percentage points away from the observed value (Supplementary File S2), indicating that drift alone can explain most of the drop in diversity.
+
+Since the population size of the Iowa RRS is small (10-20), many biallelic SNPs should fix by chance regardless of their starting minor allele frequencies.
+The deviation from simulation arises not from changes in allele frequencies per se, but rather from the fixation of linked markers across larger than expected genetic distances.
+The validity of significance cutoffs therefore depend on the accuracy of our genetic map, and the maize genetic map is known to vary among genetic backgrounds across short genetic distances (< 5cM) (McMullen et al., 2009).
+\jri{need to cite either \citep{bauer2013intraspecific} who shows there's lots of variation or maybe also \citep{rodgers2015recombination} who argues it's stable (but stable vs.
+variable is somewhat subjective}
+Deviations of the observed and simulated results could be due to selection, variation / inaccuracy in the genetic map, or a combination of these factors.
+
+To explore the roles of selection and drift independent of the genetic map, we returned to the large regions of fixation in the centromeres which showed no recombination across the full RRS experiment.
+Given the lack of recombination, each of these regions can be analyzed by the simulation of a single locus, and the high density of markers allows the clear resolution of the individual haplotypes. 
+We used the computationally phased data to measure the frequency of the fixed haplotype at each cycle, and assessed the probability of observing the fixation event given the initial frequency. 
+The BSSS chromosome 9 haplotype fixed at cycle 16 was at low frequency at cycle 0 (7 out of 68 haplotypes), but increased rapidly in frequency by cycle 8 (66 out of 70 haplotypes). 
+Simulation of the haplotype as a single locus in the RRS experiment produces this increase in frequency in only 3.9\% of 1000 independent simulations, whereas the haplotype was lost in over 80\% of the simulations.
+In BSCB1, a 30 Mb, 1.6 cM region of chromosome 2 became nearly fixed by cycle 8 (67/70) despite a prevalence of 4/72 at cycle 0, which occurred 1.5\% of the time by simulation. 
+Although these results suggest that selection may have pushed these haplotypes to fixation, the fact that fixation of such a rare haplotype still occurred in some simulations speaks to the strong genetic drift imposed upon the BSSS and BSCB1 populations.
+\jri{do you think we need to do more to explain here or above?
+in the previous paragraph we argue biallelic loci should fix by chance regardless of starting allele freq.}
+Interestingly, each of these two genomic regions harbored a different cycle 0 haplotype at higher frequency, but these higher-frequency haplotypes were subsequently lost within the RRS population. 
+In other cases, the haplotypes which eventually fixed were at moderate frequency in the cycle 0 populations and drift to fixation in the majority of simulations.
+
+\renewcommand{\arraystretch}{1.1}
+\begin{table}[tb]
+
+\begin{center}
+\caption{Ancestry of haplotypes fixed in the cycle 16 population.}
+  \textbf{}\\[-2mm]
+{\fontsize{6}{9}\sf
+\begin{tabular}{llllll}
+
+Population &	 Chr.& Interval \%(cM) & Interval \%(Mb) &	Founder  & Derived Lines \\
+BSSS &3	&40.2-41.4	&67.7-123	&CI187-2&	B94  \\
+    \hline
+ &  &  &  &  &   \\
+
+    \hline
+    \end{tabular}
+    \label{tabfix}
+}
+\end{center}
+\end{table}
+\renewcommand{\arraystretch}{1}
+
+%BSSS	
+%BSSS	3	43-48.5	129.2-157.1	NDa	B89, B94
+%BSSS	4	39.7-42.1	39.9-82.7	CI187-2	B89, B94, B67, B72, B39, B43
+%BSSS	9	30.6-33.5	20.8-26.6	Oh7b	B89, B94, B43, B17, B72, B84, B67
+%BSSS	9	34.3-35.6c	30.8-90.4c	Os420	B89, B94
+%BSCB1	2	50.2-51.8	80.6-114.5	CC5	B90, B91, B95, B97, B99
+%BSCB1	4	42.1-42.8	82.7-140	NDa	B90, B95, B97
+%BSCB1	8	46.1-50.7	125.1-145.6	P8	B90, B97, B91, B99, B54
+%
+%a ND = Not Determined (Either a recombinant haplotype or originates from an un-genotyped founder)
+%
+%b Although Oh7 is a BSCB1 founder, it is a descendant of CI.540, an un-genotyped BSSS founder. BSSS segments matching Oh7 presumably derive from CI.540
+%
+%c  Founders Ind_B2 (BSSS), CI187-2 (BSSS), R4 (BSCB1), and I205 (BSCB1) are all IBD at this region of chromosome 9

### 2015-04-19 results.tex
supp figures, tables

@@
    \vspace{-0mm}
-   \includegraphics[width=0.4\textwidth]{fig1}
+   \includegraphics[width=0.5\textwidth]{fig1}
    \renewcommand{\baselinestretch}{0.9}
    \vspace{-3mm}
-   \caption{BLAH} 
+   \caption{Principal component analysis of the SNP data from Iowa RRS.
+The axes represent the first two eigenvectors from an analysis of cycles 0-16, with projection of the founder lines onto the vector space.
+The variation explained by each eigenvector is given in parentheses on the axes.
+The populations steadily diverge at increasing cycles, with less distinction visible between the founder groups.
+The comparatively large distance between cycles 4 and 8 corresponds to a switch from one to two generations of selfing at each cycle.
+The smaller separation between cycles 8-16 corresponds to an increase in effective population size from ten to twenty.
+The BSSS cycle 0 population has drifted away from the BSSS founders, despite the absence of intentional selection during the creation and maintenance of cycle 0.
+} 
 \vspace{-6mm}
@@
    \vspace{-0mm}
-   \includegraphics[width=0.4\textwidth]{fig2}
+   \includegraphics[width=0.5\textwidth]{fig2}
    \renewcommand{\baselinestretch}{0.9}
    \vspace{-3mm}
-   \caption{BLAH} 
+   \caption{Heterozygosity and $F_{ST}$ across cycles. $H$ (left panel) and $F_{ST}$ (right panel) plotted as a function of selection cycle in each population.
+} 
 \vspace{-6mm}
@@
    \vspace{-0mm}
-   \includegraphics[width=0.4\textwidth]{fig3}
+   \includegraphics[width=0.5\textwidth]{fig3}
    \renewcommand{\baselinestretch}{0.9}
    \vspace{-3mm}
-   \caption{BLAH} 
+   \caption{ Heterozygosity ($H$) at cycle16 across all ten chromosomes in each population.
+$H$ is calculated on 15-marker sliding windows with 5 marker steps.
+Each point is plotted at the midpoint of the 15-marker window.
+} 
 \vspace{-6mm}
@@
 For example, in BSSS, a 60 Mb (1.2 cM) region of chromosome 9 became fixed by cycle 12 (Figure \ref{fig:genphys}) and traces back to the founder Os420. 
-In BSCB1, a 60 Mb (0.7 cM) region became fixed on chromosome 4 and traces back to A340 (Figure \ref{fig:genphys2}). 
+In BSCB1, a 60 Mb (0.7 cM) region became fixed on chromosome 4 and traces back to A340 (Figure \ref{fig:genphys2}). \jri{should figures 4 and 5 be made into a single 4-panel figure?}
 Table \ref{tab:fix} gives a summary of the large genomic regions that have become fixed or nearly fixed by cycle 16. 
@@
    \vspace{-0mm}
-   \includegraphics[width=0.4\textwidth]{fig4}
+   \includegraphics[width=0.5\textwidth]{fig4}
    \renewcommand{\baselinestretch}{0.9}
    \vspace{-3mm}
-   \caption{BLAH} 
+   \caption{Heterozygosity ($H$) across chromosome 9 in each cycle of BSSS plotted on the physical (A) or genetic (B) map.
+H is calculated on 15-marker sliding windows, with 5 marker steps between each calculation.
+Each data point is color-coded based on a linear transformation of recombination rate (red = low recombination rate).
+Shaded regions represent 2 cM windows found to have heterozygosity values significantly lower than expected by simulation at a given cycle.
+Light shading denotes significance at P < 0.0025, whereas the darker shading represents significance at P < 0.001.
+Comparing the physical and genetic maps reveals that the regions of fixation are extremely small genetically, but some contain much of the physical content of the chromosome.
+} 
 \vspace{-6mm}
@@
    \vspace{-0mm}
-   \includegraphics[width=0.4\textwidth]{fig5}
+   \includegraphics[width=0.5\textwidth]{fig5}
    \renewcommand{\baselinestretch}{0.9}
    \vspace{-3mm}
-   \caption{BLAH} 
+   \caption{Heterozygosity ($H$) across chromosome 4 in each cycle of BSCB1 plotted on the physical (A) or genetic (B) map.
+$H$ is calculated on 15 marker-sliding windows, with 5 marker steps between each calculation.
+Each data point is color-coded based on a linear transformation of recombination rate (red = low recombination rate).
+Shaded regions represent 2 cM windows found to have heterozygosity values significantly lower than expected by simulation at a given cycle.
+Light shading denotes significance at P < 0.0025, whereas the darker shading represents significance at P < 0.001.
+} 
 \vspace{-6mm}
@@
 
-Population &	 Chr.& Interval \%(cM) & Interval \%(Mb) &	Founder  & Derived Lines \\
+Population &	 Chr.& Interval (cM) & Interval (Mb) &	Founder & Derived Lines \\
+ \hline
 BSSS &3	&40.2-41.4	&67.7-123	&CI187-2&	B94  \\
-    \hline
+BSSS&	3	&43-48.5&	129.2-157.1	&NDa	&B89, B94 \\
+BSSS	&4	&39.7-42.1	&39.9-82.7&	CI187-2	&B89, B94, B67, B72, B39, B43\\
+
  &  &  &  &  &   \\ 
-
     \hline
@@
 %BSSS	
-%BSSS	3	43-48.5	129.2-157.1	NDa	B89, B94
-%BSSS	4	39.7-42.1	39.9-82.7	CI187-2	B89, B94, B67, B72, B39, B43
+%
+%
 %BSSS	9	30.6-33.5	20.8-26.6	Oh7b	B89, B94, B43, B17, B72, B84, B67

### 2015-04-19 sup1.tex
supp figures, tables

\begin{figure}[tb]   
  \begin{center}
   \vspace{-0mm}
   \includegraphics[width=0.4\textwidth]{FigS1}
   \renewcommand{\baselinestretch}{0.9}
   \vspace{-3mm}
   \caption{Heterozygosity in each population, observed vs. simulated data. Heterozygosity was calculated as the average across all markers genome-wide  in the BSSS (A)  and BSCB1 (B) populations. The observed data is marked by the red line, the simulations by gray dots, and the median of the simulations by a green dot. } 
\vspace{-6mm}
    \label{fig:s1}
  \end{center}
\end{figure}

\begin{figure}[tb]   
  \begin{center}
   \vspace{-0mm}
   \includegraphics[width=0.4\textwidth]{S2}
\includegraphics[width=0.4\textwidth]{S2_genetic}
   \renewcommand{\baselinestretch}{0.9}
   \vspace{-3mm}
   \caption{Heterozygosity ($H$) at cycle16 across all ten chromosomes in each population, calculated on 15-marker sliding windows with 5 marker steps as in Figure 3. Heterozygosity values in BSSS (blue dots) and BSCB1 (red dots) are superimposed in one panel. 2 cM windows of low heterozygosity that were reproduced in less than 10 of 10,000 simulations (P<0.001) are shaded in light blue (BSSS) or pink (BSCB). At this cutoff, there is no overlap between highlighted windows in the two populations. (A) Physical map. (B) Genetic map. } 
\vspace{-6mm}
    \label{fig:s1}
  \end{center}
\end{figure}

\newpage

\begin{table}
\caption{Plants and Lines Genotyped in this study.
Backgrounds of the founder inbreds are from \citet{hagdorn2003molecular}}
\begin{tabular}{ | l | l | l | l | }

	{\bf Plants from the selection cycles} &  &  &  \\ 
	Population  & Cycle & \# plants &   \\ 
	BSSS & 0 & 34 &  \\ 
	BSSS & 4 & 36 &  \\ 
	BSSS & 8 & 35 &  \\ 
	BSSS & 12 & 36 &  \\ 
	BSSS & 16 & 36 &  \\ 
	BSCB1 & 0 & 36 &  \\ 
	BSCB1 & 4 & 36 &  \\ 
	BSCB1 & 8 & 35 &  \\ 
	BSCB1 & 12 & 36 &  \\ 
	BSCB1 & 16 & 36 &  \\ 
	 &  &  &  \\ 
	BSSS Founders Genotyped &  &  &  \\ 
	Inbred & Background / Pedigree &  &  \\ 
	Ind\_Tr9\_1\_1\_6 & Reid Early Dent (Troyer Strain) &  &  \\ 
	Oh3167B & Echelberger Clarage &  &  \\ 
	I224 & Iodent &  &  \\ 
	Ind467(744) & Reid Medium &  &  \\ 
	CI.187-2 & Krug-Nebraska Reid x IA Gold Mine &  &  \\ 
	Os420 & Osterland Yellow Dent &  &  \\ 
	I159 & Iodent &  &  \\ 
	A3G-3-1-3 & BL345BxIAI129 &  &  \\ 
	Ind\_Fe2\_1073* & Troyer Reid (Early) &  &  \\ 
	Ill\_Hy & IL High Yield &  &  \\ 
	Ill\_12E & unknown &  &  \\ 
	Ind\_AH83 & Funk 176A &  &  \\ 
	Ind\_B2* & Troyer Reid (Late Butler), parent of a founder &  &  \\ 
	LE23 & IL Low Ear &  &  \\ 
	*Parents of unavailable line F1B1 &  &  &  \\ 
	 &  &  &  \\ 
	BSCB1 Founders Genotyped &  &  &  \\ 
	Inbred & Background / Pedigree &  &  \\ 
	I205 & Iodent &  &  \\ 
	Oh51A & [(OH56xWf9)Oh56] (Wooster Clarage x ?) &  &  \\ 
	A340 & 4-29 x 64 (Silver King x Northwestern Dent) &  &  \\ 
	Ill\_Hy & IL High Yield &  &  \\ 
	Oh33 & Clarage &  &  \\ 
	Oh07 & C.I.540xIII.L &  &  \\ 
	R4 & Funk Yellow Dent &  &  \\ 
	Oh40B & eight line LSC composite &  &  \\ 
	P8 & Palin Reid &  &  \\ 
	L317 & LSC &  &  \\ 
	CC5 & Golden Glow (W23) &  &  \\ 
	 &  &  &  \\ 
	Founders not analyzed &  &  &  \\ 
	Inbred & Group & Reason &  \\ 
	CI.540 & BSSS & heterozygous genotype &  \\ 
	F1B1 & BSSS & unavailable &  \\ 
	CI.617 & BSSS & unavailable &  \\ 
	WD456 & BSSS & unavailable &  \\ 
	K230 & BSCB1 & source segregates phenotypically &  \\ 
	 &  &  &  \\ 
	Derived Lines Used &  &  &  \\ 
	Inbred & Group & Cycle & Notes \\ 
	B10 & BSSS & 0 & thrown out; poor data quality \\ 
	B42 & BSCB1 & 0 &  \\ 
	B14A & BSSS & 0 & Cuzco x B14 \\ 
	B43 & BSSS & 0 &  \\ 
	B10 & BSSS & 0 &  \\ 
	B37 & BSSS & 0 &  \\ 
	B44 & BSSS & 0 &  \\ 
	B17 & BSSS & 0 &  \\ 
	B69 & BSSS & 0 &  \\ 
	B39 & BSSS & 0 &  \\ 
	B90 & BSCB1 & 7 &  \\ 
	B40 & BSSS & 0 &  \\ 
	B54 & BSCB1 & 0 &  \\ 
	B78 & BSSS & 8 & from half-sib recurrent selection program \\ 
	B72 & BSSS & 3 & from half-sib recurrent selection program \\ 
	B84 & BSSS & 7 & from half-sib recurrent selection program \\ 
	B94 & BSSS & 8 &  \\ 
	B99 & BSCB1 & 10 &  \\ 
	B11 & BSSS & 0 &  \\ 
	B89 & BSSS & 7 &  \\ 
	B95 & BSCB1 & 7 &  \\ 
	B91 & BSCB1 & 8 &  \\ 
	B67 & BSSS & 0 &  \\ 
	B73 & BSSS & 5 & from half-sib recurrent selection program \\ 
	B97 & BSCB1 & 9 &  \\ 
	    \label{tab:s1}  % caption is needed to make this work
\end{tabular}
\end{table}
\newpage

\begin{table}
\caption{Evidence of minor contamination between cycles 4 and 8 in BSSS.}
\begin{tabular}{ | l | l | l | l | }
\hline
	\  & \  & \  & \  \\ \hline
	SNPS polymorphic in cycle: & But not cycle: & in BSSS & in BSCB1 \\ \hline
	4 & 0 & 290 & 225 \\ \hline
	8 & 0 & 2242 & 319 \\ \hline
	12 & 0 & 1227 & 181 \\ \hline
	16 & 0 & 344 & 128 \\ \hline
	8 & 4 & 4499 & 1081 \\ \hline
	12 & 4 & 2328 & 669 \\ \hline
	16 & 4 & 551 & 477 \\ \hline
	12 & 8 & 1821 & 1405 \\ \hline
	16 & 8 & 178 & 997 \\ \hline
	16 & 12 & 542 & 666 \\ \hline
	0 & Founders & 1202 & 1707 \\ \hline
	4 & Founders & 822 & 885 \\ \hline
	8 & Founders & 1201 & 798 \\ \hline
	12 & Founders & 816 & 550 \\ \hline
	16 & Founders & 445 & 460 \\ \hline
	 &  &  &  \\ \hline
	Heterozygosity (H) of SNPS present in cycle 8 but not cycle 4 &  &  &  \\ \hline
	Statistic & BSSS & BSCB &  \\ \hline
	Mean & 0.052 & 0.052 &  \\ \hline
	Median & 0.028 & 0.028 &  \\ \hline
	    \label{tab:s2}  % caption is needed to make this work
\end{tabular}
\end{table}
\newpage

\begin{table}
\caption{Switch error rates from computationally phasing ‘hybrids’ simulated from derived inbred lines.}
\begin{tabular}{ | l | l | l | l | l | l | }

\hline
	\  & \  & \  & \  & \  & \  \\ \hline
	Simulated 'Hybrid' & Derived Lines Used as Priors & Population & Possible Switches & Switch errors & Rate \\ \hline
	B11xB67 & all & BSSS & 12195 & 41 & 0.003 \\ \hline
	B17xB44 & all & BSSS & 12230 & 58 & 0.005 \\ \hline
	B39xB37 & all & BSSS & 11485 & 129 & 0.011 \\ \hline
	B43xB69 & all & BSSS & 12266 & 26 & 0.002 \\ \hline
	B73xB72 & all & BSSS & 12043 & 47 & 0.004 \\ \hline
	B78xB94 & all & BSSS & 11658 & 51 & 0.004 \\ \hline
	B89xB84 & all & BSSS & 11517 & 53 & 0.005 \\ \hline
	B11xB67 & cycle 0 & BSSS & 12195 & 45 & 0.004 \\ \hline
	B17xB44 & cycle 0 & BSSS & 12230 & 54 & 0.004 \\ \hline
	B39xB37 & cycle 0 & BSSS & 11485 & 126 & 0.011 \\ \hline
	BB43xB69 & cycle 0 & BSSS & 12266 & 32 & 0.003 \\ \hline
	B42xB54 & all & BSCB1 & 11260 & 38 & 0.003 \\ \hline
	B90xB97 & all & BSCB1 & 6499 & 52 & 0.008 \\ \hline
	B91xB95 & all & BSCB1 & 7830 & 56 & 0.007 \\ \hline
	B99xB97 & all & BSCB1 & 6779 & 52 & 0.008 \\ \hline
	BB42xB54 & cycle 0 & BSCB1 & 11260 & 57 & 0.005 \\ \hline
	    \label{tab:s3}  % caption is needed to make this work
\end{tabular}
\end{table}

### 2015-04-27 discussion.tex
submodules are a pain.

@@
 Given the lack of strong selection signals in each case, genetic drift has most likely played a role in the current genetic structure of modern maize.
+
 	Several key inbreds in the ‘stiff-stalk’ heterotic group, B73, B37 and B14, were derived from the BSSS population (Darrah and Zuber, 1986; Troyer, 1999).
@@
 However, cycle 0 has also undergone maintenance since the beginning of the RRS experiment, so some genetic drift may have occurred during this time as well.
+
 	Although drift can explain most of the genetic structure genome-wide, phenotypic data provide clear evidence that selection has altered the frequencies of favorable alleles in the BSSS and BSCB1 populations.
@@
 This non-overlap between putatively selected loci in complementary populations has also been observed in commercial breeding programs (Feng et al., 2006).
+
 The observation that the same targets of selection are not observed in the opposing heterotic populations bears implications for the genetic mechanisms responsible for heterosis and the success of maize hybrids.
@@
 Our data are consistent with such a process having been important for hybrid improvement in the Iowa RSS experiment, and the lack of extreme differentiation seen over time across US Corn Belt germplasm is consistent with a role for such selection in maize hybrid improvement in general (van Heerwaarden et al., 2012).
+
 Our results are consistent with the dominance and pseudo-overdominance models, but we caution that the exact outcome of particular models will depend strongly on the effects of selection, drift, and the frequencies of beneficial alleles.

### 2015-04-27 discussion.tex
work on discussion

--- 2015-04-15 discussion.tex
@@
 \section*{Discussion}
-Reciprocal recurrent selection serves as a model for the method of hybrid maize improvement employed broadly in North America (Duvick et al., 2004).
-In the Iowa RRS experiment, the BSSS and BSCB1 populations steadily lost genetic diversity as they become more diverged from one another.
-Principal component analysis shows that as the effective population size and the rates of inbreeding were altered, the rates of change in population structure were altered as well.
-The population structure and divergence between BSSS and BSCB1 can be largely reproduced by simulation without any selection.
-The trends in these populations support the hypothesis that the majority of the genetic structure observed can be attributed to genetic drift alone, despite effective selection for phenotypic improvement.
+
+In the Iowa RRS experiment, the BSSS and BSCB1 populations steadily lost genetic diversity as they became more differentiated from one another, and principal component analysis shows that as the effective population size and the rates of inbreeding were altered, the rates of change in population structure were altered as well.
+These patterns of population structure, diversity, and differentiation between BSSS and BSCB1 can be largely reproduced by simulation without any selection, supporting the hypothesis that the majority of the genetic structure observed can be attributed to genetic drift alone, despite effective selection for phenotypic improvement.
 This drift was driven by inbreeding through self-pollination and the small effective population sizes used to select a limited number of high-performing, potentially related individuals at each cycle.
-The trends observed in the Iowa RRS are also a common theme in similar recurrent selection experiments (Romay et al., 2012; C.
-Lamkey and A.
-Lorenz, personal communication).
-Increased population structure due to increased inbreeding and lowered effective population size are visible in the modern ‘stiff-stalk’ and ‘non stiff-stalk’ heterotic groups as well (van Heerwaarden et al., 2012).
-Given the lack of strong selection signals in each case, genetic drift has most likely played a role in the current genetic structure of modern maize.
-	Several key inbreds in the ‘stiff-stalk’ heterotic group, B73, B37 and B14, were derived from the BSSS population (Darrah and Zuber, 1986; Troyer, 1999).
+
+Reciprocal recurrent selection serves as a model for the method of hybrid maize improvement employed broadly in North America \citep{duvick2004long}.
+The trends observed in the Iowa RRS are also a common theme in similar recurrent selection experiments \citep{Romay2013,lamkey2014relative}. 
+Increased population structure due to increased inbreeding and lowered effective population size are visible in the modern ‘stiff-stalk’ and ‘non stiff-stalk’ heterotic groups as well \citep{van2012historical}.
+Given the lack of strong selection signals in each case, genetic drift has most likely played a large role in the current genetic structure of modern maize.
+
+Several key inbreds in the ‘stiff-stalk’ heterotic group, B73, B37 and B14, were derived from the BSSS population \citep{darrah19861985, 1986; troyer1999background}. 
 B37 and B14 were derived from cycle 0, and B73 was derived from a half-sib recurrent selection program also started with the BSSS population. 
-We examined these three inbreds at the pericentromeric regions listed in Table 1, and found that in most cases they carry different haplotypes from those that rose to high frequency in the RRS experiment.
+We examined these three inbreds at the pericentromeric regions listed in Table \ref{tabfix} \jri{correct?}, and found that in most cases they carry different haplotypes from those that rose to high frequency in the RRS experiment.
 Near isogenic lines that substitute these centromeric haplotypes from key inbreds could be used to determine if the haplotypes at high frequency in cycle 16 of BSSS provide a phenotypic advantage when crossed with BSCB1.
-	Both our study and previous work (Labate et al., 1997) identified genetic differentiation between cycle 0 of BSSS and the founder inbreds.
+Both our study and previous work \citep{labate1997molecular} identified genetic differentiation between cycle 0 of BSSS and the founder inbreds. 
 The loss of diversity in the cycle 0 population likely led to the loss of some rare alleles and haplotypes. 
-BSSS was maintained for several years before RRS was initiated, and the differentiation could have occurred due to low effective population size during maintenance.
+The BSSS was maintained for several years before RRS was initiated, and the differentiation could have occurred due to low effective population size during maintenance. 
 If so, then this drift will have impacted the trajectory of the RRS program and of the key inbreds derived from BSSS. 
 However, cycle 0 has also undergone maintenance since the beginning of the RRS experiment, so some genetic drift may have occurred during this time as well.
+
 Although drift can explain most of the genetic structure genome-wide, phenotypic data provide clear evidence that selection has altered the frequencies of favorable alleles in the BSSS and BSCB1 populations.
-Numerous experiments have shown that the selected populations and the hybrids formed from them are superior to their cycle 0 ancestors (Smith, 1983; Keeratinijakal and Lamkey, 1993; Schnicker and Lamkey, 1993; Holthaus and Lamkey, 1995).
+Numerous experiments have shown that the selected populations and the hybrids formed from them are superior to their cycle 0 ancestors (Smith, 1983; keeratinijakal1993responses, Schnicker and Lamkey, 1993; Holthaus and Lamkey, 1995).
 Improvement occurred not only for hybrid yield, but concurrently for plant architecture and tolerance to high-density planting in both hybrids and inbreds (Brekke et al., 2011a; Brekke et al., 2011b; Edwards, 2011; Lauer et al., 2012).

### 2015-04-27 discussion.tex
all refs added

@@
 Although drift can explain most of the genetic structure genome-wide, phenotypic data provide clear evidence that selection has altered the frequencies of favorable alleles in the BSSS and BSCB1 populations.
-Numerous experiments have shown that the selected populations and the hybrids formed from them are superior to their cycle 0 ancestors (Smith, 1983; keeratinijakal1993responses, Schnicker and Lamkey, 1993; Holthaus and Lamkey, 1995).
-Improvement occurred not only for hybrid yield, but concurrently for plant architecture and tolerance to high-density planting in both hybrids and inbreds (Brekke et al., 2011a; Brekke et al., 2011b; Edwards, 2011; Lauer et al., 2012).
-Since these same trends are observed across wider North American germplasm (Tollenaar, 1989; Duvick, 2005), identification and characterization of the loci conferring phenotypic improvement in BSSS and BSCB1 could reveal some of the genetic mechanisms by which maize yield has steadily improved over the decades.
+Numerous experiments have shown that the selected populations and the hybrids formed from them are superior to their cycle 0 ancestors \citep{smith1983evaluation, keeratinijakal1993responses, schnicker1993interpopulation, holthaus1995population}.
+Improvement occurred not only for hybrid yield, but concurrently for plant architecture and tolerance to high-density planting in both hybrids and inbreds \citep{brekke2011selection, brekke2011selectionb, edwards2011changes, lauer2012morphological}.
+Since these same trends are observed across wider North American germplasm \citep{tollenaar1989genetic}, identification and characterization of the loci conferring phenotypic improvement in BSSS and BSCB1 could reveal some of the genetic mechanisms by which maize yield has steadily improved over the decades.
 We find that the genomic regions of extremely low diversity evident at cycle 16 are unlikely to be produced by simulation, and heterozygosity falls more than expected across the genome as a whole.  
@@
 In this case, it is difficult to assess significance with only a single population to test.
-This type of analysis would be more effective across several replicated populations, which can control for genetic drift due to the independence between the selections in each replicate (C.
-Lamkey and A.
-Lorenz, personal communication).
-	van Heerwaarden et al., 2012 assessed variation and tested for selection across a wider range of North American maize germplasm using the same SNP array.
-Although selected loci were identified in that study, the overall effects on genome diversity and haplotype patterns were relatively small compared with the impact of domestication.
+This type of analysis is more effective across several replicated populations, which can control for genetic drift due to the independence between the selections in each replicate \citep{lamkey2014relative}.
+
+\citet{van2012historical} assessed variation in and tested for selection across a wider range of North American maize germplasm using the same SNP array. 
+Although selected loci were identified in that study, the overall effects on genomic diversity and haplotype patterns were relatively small compared with the impact of domestication. 
 That study was a wide survey across many germplasm sources. 
@@
 In support of this idea, we find that the most likely targets of selection between the BSSS and BSCB1 populations occur at non-overlapping loci. 
-This non-overlap between putatively selected loci in complementary populations has also been observed in commercial breeding programs (Feng et al., 2006).
+This non-overlap between putatively selected loci in complementary populations has also been observed in commercial breeding programs \citep{feng2006temporal}.
 
@@
 The observed pattern instead favors a dominance model, where fixation of a haplotype in one population simply selects against that same haplotype in the other population. 
-Because most deleterious alleles are rare in both heterotic groups (S.
-Mezmouk and J.
-Ross-Ibarra, unpublished results), most haplotypes in the second population will have a different suite of deleterious variants and will complement the fixed haplotype reasonably well such that selection will have little impact on diversity in the second population.
-Our data are consistent with such a process having been important for hybrid improvement in the Iowa RSS experiment, and the lack of extreme differentiation seen over time across US Corn Belt germplasm is consistent with a role for such selection in maize hybrid improvement in general (van Heerwaarden et al., 2012).
+Because most deleterious alleles are rare in both heterotic groups \citep{Mezmouk2014}, most haplotypes in the second population will have a different suite of deleterious variants and will complement the fixed haplotype reasonably well, such that selection will have little impact on diversity in the second population.
+Our data are consistent with such a process having been important for hybrid improvement in the Iowa RSS experiment, and the lack of extreme differentiation seen over time across US Corn Belt germplasm is consistent with a role for such selection in maize hybrid improvement in general \citep{van2012historical}.
 

### 2015-04-27 intro.tex
all refs added

@@
 Although maize is naturally an out-crossing organism, modern breeding develops highly inbred lines that are then used in controlled crosses to produce hybrids. 
-Hybrid maize has been so successful that it quickly replaced long-standing mass-selected open pollinated varieties (Crabb, 1947).
-Maize inbred lines are now partitioned into separate inbreeding ‘heterotic groups’ that maximize performance and hybrid vigor (heterosis) when an inbred line of one group is crossed with the other group (Tracy and Chandler, 2006).
-The shift towards selection of inbred lines based on their ability to generate good hybrids – referred to as ‘combining ability’ – constituted an abrupt change from the open-pollinated mass selection that breeders practiced for millennia (Anderson, 1944; Troyer, 1999).
-Multiple studies with molecular markers have indicated that the modern era of single-cross hybrid maize breeding has led to a dramatic restructuring of population genetic variation (Duvick et al., 2004; Ho et al., 2005; Feng et al., 2006).
+Hybrid maize has been so successful that it quickly replaced long-standing mass-selected open pollinated varieties \citep{crabb1947hybrid}. 
+Maize inbred lines are now partitioned into separate inbreeding ‘heterotic groups’ that maximize performance and hybrid vigor (heterosis) when an inbred line of one group is crossed with the other group \citep{tracy2006historical}.
+The shift towards selection of inbred lines based on their ability to generate good hybrids – referred to as ‘combining ability’ – constituted an abrupt change from the open-pollinated mass selection that breeders practiced for millennia \citep{Anderson, 1944; troyer1999background}.
+Multiple studies with molecular markers have indicated that the modern era of single-cross hybrid maize breeding has led to a dramatic restructuring of population genetic variation \citep{duvick2004long, ho2005extent, feng2006temporal}.
 Different heterotic groups have diverged genetically over time to become highly structured and isolated populations. 
@@
 So far, these high-density studies have examined a broad spectrum of germplasm at various points in the history of maize to search for the signals of population structure and artificial selection \citep{Hufford2012b, van2012historical}.
-Although selective sweeps remaining from domestication are clearly visible, the impact of selection during modern breeding appears comparatively small in terms of its impact on genomic diversity despite steady, heritable improvement in phenotype (Duvick, 2005).
+\jri{add more}
+Although selective sweeps remaining from domestication are clearly visible, the impact of selection during modern breeding appears comparatively small in terms of its impact on genomic diversity \citep{Hufford2012b,van2012historical}, despite steady, heritable improvement in phenotype \citep{duvick2005contribution}.
 The lack of distinct selection signals from modern breeding may be due to specific selective events occurring in different populations, necessitating a more focused look within heterotic groups or even single breeding programs.
@@
 1 (BSCB1) Reciprocal Recurrent Selection Program of the USDA-ARS at Ames, Iowa (hereafter referred to as the Iowa RRS) represents one of the best-documented public experiments on selection for combining ability and hybrid performance.
-BSSS and BSCB1 have been recurrently selected for improved cross-population hybrids (Penny and Eberhart, 1971). 
-This model of selection, named reciprocal recurrent selection, provides the generalized model for strategies used in commercial maize breeding (Comstock et al., 1949; Duvick et al., 2004). 
-The Iowa RRS experiment proves especially relevant because lines derived from the BSSS population have had a major impact upon the development of commercial hybrids (Duvick et al., 2004; Darrah and Zuber, 1986), the formation of modern heterotic groups (Troyer, 1999; Senior et al., 1998), and the choice of a maize reference genome \citep{schnable2009the-b73-maize}.
+BSSS and BSCB1 have been recurrently selected for improved cross-population hybrids \citep{penny1971twenty}. 
+\jri{add some text explaining recurrent selection}
+This model of selection, named reciprocal recurrent selection, provides the generalized model for strategies commonly used in commercial maize breeding \citep{comstock1949breeding, duvick2004long}. 
+The Iowa RRS experiment proves especially relevant because lines derived from the BSSS population have had a major impact upon the development of commercial hybrids \citep{duvick2004long, darrah19861985}, the formation of modern heterotic groups \citep{troyer1999background; senior1998utility}, and the choice of a maize reference genome \citep{schnable2009the-b73-maize}.

### 2015-05-08 Fig3/Fig3plotting.R
updated code for figure 3

@@ -1,0 +1,25 @@
+#!/usr/bin/Rscript
+rm(list=ls())
+library(ggplot2)
+library(dplyr)
+alldata =c()
+for(i in 1:10){
+filename=paste("./data/H_cycle16_LG", i, ".txt", sep="")
+Hdata <- read.table(filename, header=T)
+alldata <- rbind(alldata, Hdata)
+}
+
+BSSSHdata = alldata[,-2]
+BSCBHdata = alldata[,-1]
+colnames(BSSSHdata) = c("H", "LG", "Cycle", "Start", "Stop", "physpos", "Genpos", "Rate")
+colnames(BSCBHdata) = c("H", "LG", "Cycle", "Start", "Stop", "physpos", "Genpos", "Rate")
+BSSSHdata$Population = "BSSS"
+BSCBHdata$Population = "BSCB1"
+both = rbind(BSSSHdata, BSCBHdata)
+
+png("../fig3.png", height=900, width=700)
+ggplot(data=both) + aes(x=physpos/1000000, y=H, colour=Population) + facet_grid(LG~Population) + geom_point(size=1) + scale_colour_manual(values=c("blue", "red2")) + theme_bw() + xlab("Physical Position (MB)")
+dev.off()
+
+
+

### 2015-05-08 Fig3/overlappingH.sh

@@ -1,0 +1,62 @@
+#!/bin/bash
+
+for cycle in 0 4 8 12 16
+do
+
+for chrom in {1..10}
+do
+
+R --slave --no-save <<HEREFILE
+
+source("../rcode/building_functions.R")
+source("../rcode/analysis_functions.R")
+source("../rcode/formatting_functions.R")
+source("../rcode/writing_functions.R")
+load("../data/new_full_set.RData")
+
+
+
+	currentseq=1:sum(new_full_set\$LG==${chrom})
+        windows <- define.overlapping.windows(currentseq, window=15, step=5)
+
+	chrommarkers <- new_full_set\$calls[new_full_set\$LG==${chrom},]
+	chromgen <- new_full_set\$F2map[new_full_set\$LG==${chrom}]
+	chromphys <- new_full_set\$position[new_full_set\$LG==${chrom}]
+
+windowtable <- t(apply(windows, 1, function(x){
+  midmarker <- ceiling(median(x[1]:x[2]))
+  physpos <- chromphys[midmarker]
+  phystart <- chromphys[x[1]]
+  phystop <- chromphys[x[2]]
+  genstart <- chromgen[x[1]]
+  genstop <- chromgen[x[2]]
+  genpos <- chromgen[midmarker]
+  rate <-(genstop-genstart)/(phystop/1000000 - phystart/1000000)
+  return(c(x[1], x[2], physpos, genpos, rate))  
+  }))
+
+  windowtable <- data.frame(windowtable)
+
+  colnames(windowtable) <- c("start", "stop", "physpos", "genpos", "rate")
+
+  outbreds <- split.status(new_full_set, "Outbred")
+
+outbreds <- split.all(outbreds)
+
+realresults <- t(apply(windows, 1, function(x){
+  BSSScalls <- outbreds\$BSSS_${cycle}\$calls[outbreds\$BSSS_${cycle}\$LG==${chrom},]  
+  BSCBcalls <- outbreds\$BSCB_${cycle}\$calls[outbreds\$BSCB_${cycle}\$LG==${chrom},]
+  hetSS <- averageH(BSSScalls[x[1]:x[2],])
+  hetNSS <- averageH(BSCBcalls[x[1]:x[2],])
+  return(c(hetSS, hetNSS, ${chrom}, ${cycle}))
+}))
+
+colnames(realresults) <- c("HetSS", "HetNSS", "LG", "Cycle")
[…]

### 2015-05-08 rcode/analysis_functions.R
added old R functions

@@ -1,0 +1,238 @@
+#function to get frequency of hets at each population_cycle
+get.hetfreq <- function(dataset){
+	hets <- is.het(dataset$calls[1:length(dataset$calls)])
+	hetfreq <- (sum(hets, na.rm=T))/(sum(hets, na.rm=T) + sum(!hets, na.rm=T))
+	return(hetfreq)
+}
+
+
+#get HWE significance of each marker in a set
+get.HWE <- function(marker){
+	if (length(levels(as.factor(marker))) > 1) {
+	marker <- genotype(marker, sep="")
+	teststat <- HWE.exact(marker)$p.value
+	}
+	else{teststat <- NA}
+	return(teststat)
+}
+
+#get the HWE probabilities for a dataset
+dataset.HWE <- function(dataset){
+	pvals <- apply(dataset$calls, 1, get.HWE)
+	return(pvals)	
+}
+
+
+#count number of switches in phased data
+countphase <- function(prephased, phased) {
+	prephased[prephased=="?"] <- NA
+	hetcode <- apply(prephased, 2, function(x){
+			hets <- (x[1] != x[2])
+			hets_na <- hets==TRUE & !is.na(hets)
+			return(hets_na)
+			})	
+	prephasedhets <- prephased[,hetcode]
+	phasedhets <- phased[,hetcode]
+
+	phasecount <- c()
+	for (i in 1:ncol(phasedhets)){
+		phasecount[i] <- which(phasedhets[,i]==prephasedhets[1,i])
+	}
+	#print(paste("# of sites:", ncol(phasedhets)))
+	possible <- ncol(phasedhets)-1
+	switches <- sum(abs(diff(phasecount)))
+	#print(paste("# of switches", switches ))
+	switchrate <- switches/(ncol(phasedhets)-1)
+	#ratename<-"rate="
+	#switchrateout <- sprintf("%s\t%.7f\n", ratename,switchrate) 
+	#cat(switchrateout)
+	results <- list(possible, switches)
+	return(results)	
+
+}		
+
+calcH <- function(genotypes){
+		alleles <- unlist(strsplit(genotypes, split=""))
+		allelecounts <- table(alleles)
+		total <- sum(allelecounts)
+		if(length(allelecounts)==2){
+		freq1 <- allelecounts[1]/total
+		freq2 <- allelecounts[2]/total
+		H <- 1-(freq1^2 + freq2^2)
+		}
+		else if (length(allelecounts)==1){
+			H<-0
+			}
[…]

### 2015-05-08 rcode/building_functions.R

@@ -1,0 +1,111 @@
+
+#function to get genotypes from a table of lines
+
+get.genos <- function(linetable, sample_codes, genotypes){
+	markernames <- genotypes[,1]
+	justgenotypes <- genotypes[,-1]
+	linenames <- linetable$Line
+	calls <- matrix(ncol=length(linetable$SampleID), nrow=dim(genotypes)[[1]])
+	for(i in 1:length(linetable$SampleID)){
+		calls[,i] <- as.character(justgenotypes[,which(sample_codes == linetable$SampleID[i])])
+	}
+	dimnames(calls) <- list(markernames,linenames)
+	return(calls)
+}
+
+
+get.scores <- function(linetable, sample_codes, genotypes){
+	markernames <- genotypes[,1]
+	justgenotypes <- genotypes[,-1]
+	linenames <- linetable$Line
+	calls <- matrix(ncol=length(linetable$SampleID), nrow=dim(genotypes)[[1]])
+	for(i in 1:length(linetable$SampleID)){
+		calls[,i] <- justgenotypes[,which(sample_codes == linetable$SampleID[i])]
+	}
+	dimnames(calls) <- list(markernames,linenames)
+	return(calls)
+}
+
+
+#function to change hets to missing data
+is.het <- function(call){
+	return(substr(call,1,1) != substr(call,2,2))
+	}
+
+remove.hets <- function(calls){
+	hets <- is.het(calls)
+	calls[which(hets=="TRUE")] <- NA
+	return(calls)
+}
+
+#cut out markers with a certain missing data threshold
+na.cut.markers <- function(dataset, threshold) {
+	nacounts <- apply(dataset$calls, 1, function(x){sum(is.na(x))})
+	torun <- dataset
+	keep <- nacounts <= threshold
+	newset <- row.prune(dataset=dataset, keep=keep)
+	return(newset)
+	}
+
+#remove a particular line from the dataset	
+remove.line <- function(line, dataset){
+	column <- dimnames(dataset$calls)[[2]]!=line
+	newset <- column.prune(dataset=dataset, keep=column)
+	return(newset)
+	}	
+
+#remove monomorphic calls	
+remove.monos <- function(dataset){
+	alleles <- apply(dataset$calls, 1, function(x){
+		counts <- rle(sort(as.vector(x)))
+		counts <- length(counts$values)
[…]

### 2015-05-08 rcode/formatting_functions.R

@@ -1,0 +1,125 @@
+
+####get rid of het calls, for phasing calibration using derived inbreds
+is.het <- function(call){
+	return(substr(call,1,1) != substr(call,2,2))
+	}
+
+remove.hets <- function(calls){
+	hets <- is.het(calls)
+	calls[which(hets=="TRUE")] <- NA
+	return(calls)
+}
+
+
+
+
+
+
+#split dataset by group	
+split.lines <- function(dataset, groupname){
+	keep <- which(dataset$group==groupname)
+	newname <- dataset
+	newname$calls <- dataset$calls[,keep]
+    newname$scores <- dataset$scores[,keep]
+   	newname$status <- dataset$status[keep]
+   	newname$group <- dataset$group[keep]
+   	newname$lines <- dataset$lines[keep]
+	newname$cycle <- dataset$cycle[keep]
+	return(newname)
+	}	
+
+#split based on derived/founder/outbred
+split.status <- function(dataset, statusname){
+	keep <- which(dataset$status==as.character(statusname))
+	newname <- dataset
+	newname$calls <- dataset$calls[,keep]
+    newname$scores <- dataset$scores[,keep]
+   	newname$status <- dataset$status[keep]
+   	newname$group <- dataset$group[keep]
+   	newname$lines <- dataset$lines[keep]
+	newname$cycle <- dataset$cycle[keep]
+	return(newname)
+	}
+	
+	
+split.chromosomes <- function(dataset){
+	finalset <- lapply(levels(as.factor(dataset$LG)), function(x){
+		keep <- which(dataset$LG==x)
+		newname <- row.prune(dataset, keep)
+		return(newname)
+		})
+	return(finalset)
+}	
+	
+	
+	
+	
+	
+#prune out rows(markers) given a logical or vector of indexes
+row.prune <- function(dataset, keep){
+	newname <- dataset
+	newname$calls <- dataset$calls[keep,]
+    newname$scores <- dataset$scores[keep,]
+    newname$LG <- dataset$LG[keep]
+    newname$position <- dataset$position[keep]
+    newname$markernames <- dataset$markernames[keep]
+	newname$markercodes <- dataset$markercodes[keep]
+	newname$alleles <- dataset$alleles[keep,]
+	newname$genmap <- dataset$genmap[keep]
+	newname$F2map <- dataset$F2map[keep]
+	return(newname)
+	}
+
+
+
[…]

### 2015-05-08 rcode/writing_functions.R

@@ -1,0 +1,212 @@
+write.flopfile <- function(dataset, outfile){
+	#transpose the matrix and replace NA's
+	calls <- dataset$calls
+	calls <- t(calls)
+	missing <- is.na(calls)
+	calls[which(missing==TRUE)] <- "--"
+	#add the sequence position
+	calls <- rbind(as.character(dataset$position), calls)
+	#add chromosome
+	calls <- rbind(as.character(dataset$LG), calls)
+	#add line names
+	first_column <- c("Chromosome", "Position", dimnames(dataset$calls)[[2]])
+	calls <- cbind(first_column, calls)
+	write.table(calls, file=outfile, quote=F, row.names=F, col.names=F, sep=" ")
+}
+
+#write a dataset to PLINK outfiles	
+write.plink <- function(dataset, pedfile, mapfile, delimiter, assn="FALSE"){
+	pheno <- 1
+	linetag <- "0 0 0"
+	#write the pedfile
+	calls <- dataset$calls
+	calls[which(is.na(calls))] <- "00"
+	if(delimiter=="group"){
+		popname <- as.character(dataset$group)
+		}
+	else if(delimiter=="status"){
+		popname <- as.character(dataset$status)
+		}
+	else if (delimiter=="cycle"){
+		popname <- paste(as.character(dataset$cycle), as.character(dataset$status), sep="_")
+		}		
+	if(assn=="TRUE"){
+		pheno <- as.numeric(as.factor(dataset$group))
+		linetag <- ""
+		}	
+	
+	lines <- apply(rbind(popname, dimnames(calls)[[2]], pheno, calls), 2, function(line) {
+		var <- unlist(strsplit(line[-(1:2)], split=""))
+		paste(c(line[1:2], linetag, var), collapse=" ")
+		})
+	cat(lines, file=pedfile, sep="\n")
+			
+	#write the mapfile
+	mapdata <- cbind(as.character(dataset$LG), as.character(dataset$markernames), rep(0,length(dataset$LG)), as.character(dataset$position))
+	write.table(mapdata, file=mapfile, col.names=F, row.names=F, quote=F, sep="\t")
+	}
+
+#write out a fasta file
+write.fasta <- function(calls, outfile) {
+	missing <- is.na(calls)
+	calls[which(missing==TRUE)] <- "NN"
+	
[…]

### 2015-05-08 FigS1/BSCB.R
updated Figure S1 with new sim data

@@ -1,0 +1,35 @@
+#!/usr/bin/Rscript
+rm(list=ls())
+library(ggplot2)
+library(reshape)
+
+simresults <- read.table("../data/BSCB_fullsim_HH.txt")
+colnames(simresults) = c("Cycle4", "Cycle8", "Cycle12", "Cycle16")
+melted_CB_all = melt(simresults)
+colnames(melted_CB_all) = c("Cycle", "H")
+melted_CB_all$Pruning = "all"
+
+colnames(melted_CB_all) = c("Cycle",  "H", "Pruning")
+medians = plyr::ddply(.data=melted_CB_all, .variables=c("Cycle", "Pruning"), .fun=function(x){median(x[,2])})
+colnames(medians) = c("Cycle", "Pruning",  "Med")
+realdata = read.table("BSCB_realresults.txt", header=T)
+getbounds = function(x){
+  quantile(x[,2], probs=c(0.01, 0.99))
+}
+bounds = plyr::ddply(.data=melted_CB_all, .variables=c("Cycle", "Pruning"), .fun=getbounds)
+colnames(bounds) = c("Cycle", "Pruning", "one", "ninenine")
+realdata = realdata[realdata$Pruning=="all",]
+png("../FigS1B.png", width=400,height=300)
+myplot = ggplot()
+myplot = myplot + geom_point(data=melted_CB_all, aes(x=Cycle, y=H), alpha=0.1) + geom_line(data=bounds, aes(x=Cycle, y=one, group=Pruning), color="black") + geom_line(data=bounds, aes(x=Cycle, y=ninenine, group=Pruning), color="black")
+myplot = myplot + geom_point(data=medians, aes(x=Cycle, y=Med), color="green") 
+myplot = myplot + geom_line(data=realdata, aes(x=Cycle, y=H, group=Pruning), color="red") + geom_point(data=realdata, aes(x=Cycle, y=H), color="red") + theme_bw() + ggtitle("BSCB1 Genome-wide Simulation")
+myplot + theme(axis.title.x=element_text(size=0))
+dev.off()
+
+# png("BSCB_sims_gray.png", width=1000,height=600)
+# test = ggplot()
[…]

### 2015-05-08 FigS1/BSSS.R

@@ -1,0 +1,37 @@
+#!/usr/bin/Rscript
+rm(list=ls())
+library(ggplot2)
+library(reshape)
+
+
+
+simresults <- read.table("../data/BSSS_fullsim_HH.txt")
+colnames(simresults) = c("Cycle4", "Cycle8", "Cycle12", "Cycle16")
+melted_CB_all = melt(simresults)
+colnames(melted_CB_all) = c("Cycle", "H")
+melted_CB_all$Pruning = "all"
+
+colnames(melted_CB_all) = c("Cycle",  "H", "Pruning")
+medians = plyr::ddply(.data=melted_CB_all, .variables=c("Cycle", "Pruning"), .fun=function(x){median(x[,2])})
+colnames(medians) = c("Cycle", "Pruning",  "Med")
+realdata = read.table("BSSS_realresults.txt", header=T)
+getbounds = function(x){
+  quantile(x[,2], probs=c(0.01, 0.99))
+}
+bounds = plyr::ddply(.data=melted_CB_all, .variables=c("Cycle", "Pruning"), .fun=getbounds)
+colnames(bounds) = c("Cycle", "Pruning", "one", "ninenine")
+realdata = realdata[realdata$Pruning=="all",]
+png("../FigS1A.png", width=400,height=300)
+myplot = ggplot()
+myplot = myplot + geom_point(data=melted_CB_all, aes(x=Cycle, y=H), alpha=0.1) + geom_line(data=bounds, aes(x=Cycle, y=one, group=Pruning), color="black") + geom_line(data=bounds, aes(x=Cycle, y=ninenine, group=Pruning), color="black")
+myplot = myplot + geom_point(data=medians, aes(x=Cycle, y=Med), color="green") 
+myplot =myplot + geom_line(data=realdata, aes(x=Cycle, y=H, group=Pruning), color="red") + geom_point(data=realdata, aes(x=Cycle, y=H), color="red") + theme_bw() + ggtitle("BSSS Genome-wide Simulation")
+myplot + theme(axis.title.x=element_text(size=0))
+dev.off()
+
+# png("BSSS_sims_gray.png", width=1000,height=600)
+# test = ggplot()
[…]

### 2015-05-08 FigS2/FigS2plotting.R
updated Figure S2 with new sim data

@@ -1,0 +1,34 @@
+#!/usr/bin/Rscript
+rm(list=ls())
+library(ggplot2)
+alldata =c()
+for(i in 1:10){
+filename=paste("../Fig3/data/H_cycle16_LG", i, ".txt", sep="")
+Hdata <- read.table(filename, header=T)
+alldata <- rbind(alldata, Hdata)
+}
+
+filename="../data/2cm_sim_collated_results.txt"
+simdata <- read.table(filename, header=T)
+
+
+sigs=simdata[!is.na(simdata$Sigat001),]
+
+sigs$Sigat001 = factor(sigs$Sigat001, ordered=T, levels=c("BSCB", "BSSS", "Both"))
+levels(sigs$Sigat001) = c("BSCB1", "BSSS", "Both")
+blocks=geom_rect(data=sigs, aes(fill=Sigat001, xmin=phystart/1000000, xmax=phystop/1000000, ymin=0, ymax=0.5)) 
+BSSS_H <- geom_point(data=alldata, aes(x=physpos/1000000, y=HetSS), colour="blue", size=1)
+BSCB1_H <- geom_point(data=alldata, aes(x=physpos/1000000, y=HetNSS), colour="red2", size=1)
+png('../FigS2A.png')
+ggplot(data=alldata) + blocks + BSSS_H + BSCB1_H + facet_grid(LG~.) + ylab("Heterozygosity") + xlab("Position (Mb)") + theme_bw() +
+scale_fill_manual(values=c("pink1", "skyblue2", "green"), name="P < 0.001") + theme(panel.border=element_rect(size=0.1))  + theme(axis.text.y=element_text(size=8))
+dev.off()
+
+blocks=geom_rect(data=sigs, aes(fill=Sigat001, xmin=genstart, xmax=genstop, ymin=0, ymax=0.5)) 
+BSSS_H <- geom_point(data=alldata, aes(x=genpos, y=HetSS), colour="blue", size=1)
+BSCB1_H <- geom_point(data=alldata, aes(x=genpos, y=HetNSS), colour="red2", size=1)
+png('../FigS2B.png')
+ggplot(data=alldata) + blocks + BSSS_H + BSCB1_H + facet_grid(LG~.) + ylab("Heterozygosity") + xlab("Position (cM)") + theme_bw() +
+  scale_fill_manual(values=c("pink1", "skyblue2", "green"), name="P < 0.001") + theme(panel.border=element_rect(size=0.1))  + theme(axis.text.y=element_text(size=8))
+dev.off()
+

### 2015-05-10 results.tex
joined figures

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
-\begin{figure}[tb]   
+\begin{figure*}[tb]   
   \begin{center}
-   \vspace{-0mm}
-   \includegraphics[width=0.5\textwidth]{fig4}
-   \renewcommand{\baselinestretch}{0.9}
-   \vspace{-3mm}
-   \caption{Heterozygosity ($H$) across chromosome 9 in each cycle of BSSS plotted on the physical (A) or genetic (B) map.
+  % \vspace{-0mm}
+   \includegraphics[width=\linewidth]{fig45}
+ %  \renewcommand{\baselinestretch}{0.9}
+ %  \vspace{-3mm}
+   \caption{Heterozygosity ($H$) across chromosomes 9 (left) and 4 (right) in each cycle of BSSS plotted on the physical (top) or genetic (bottom) map.
 H is calculated on 15-marker sliding windows, with 5 marker steps between each calculation.
@@
   \end{center}
-\end{figure}
+\end{figure*}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
-\begin{figure}[tb]   
-  \begin{center}
-   \vspace{-0mm}
-   \includegraphics[width=0.5\textwidth]{fig5}
-   \renewcommand{\baselinestretch}{0.9}
-   \vspace{-3mm}
-   \caption{Heterozygosity ($H$) across chromosome 4 in each cycle of BSCB1 plotted on the physical (A) or genetic (B) map.
+%\begin{figure}[tb]   
+%  \begin{center}
+%   \vspace{-0mm}
+%   \includegraphics[width=0.5\textwidth]{fig5}
+%   \renewcommand{\baselinestretch}{0.9}
+%   \vspace{-3mm}
+%   \caption{Heterozygosity ($H$) across chromosome 4 in each cycle of BSCB1 plotted on the physical (A) or genetic (B) map.
 $H$ is calculated on 15 marker-sliding windows, with 5 marker steps between each calculation.
@@
 Light shading denotes significance at P < 0.0025, whereas the darker shading represents significance at P < 0.001.
-} 
-\vspace{-6mm}
-    \label{fig:genphys2}
-  \end{center}
-\end{figure}
+%} 
+%\vspace{-6mm}
+%    \label{fig:genphys2}
+%  \end{center}
+%\end{figure}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE

### 2015-05-10 results.tex
notes in intro, pasted supp. figs, cleaned up supp. tables s2-s4, fixed title

@@
 In most cases, these fixed haplotypes can be traced back to single founder inbreds. 
-For example, in BSSS, a 60 Mb (1.2 cM) region of chromosome 9 became fixed by cycle 12 (Figure \ref{fig:genphys}) and traces back to the founder Os420. 
-In BSCB1, a 60 Mb (0.7 cM) region became fixed on chromosome 4 and traces back to A340 (Figure \ref{fig:genphys2}). \jri{should figures 4 and 5 be made into a single 4-panel figure?}
+For example, in BSSS, a 60 Mb (1.2 cM) region of chromosome 9 became fixed by cycle 12 and traces back to the founder Os420, and in BSCB1, a 60 Mb (0.7 cM) region became fixed on chromosome 4 and traces back to A340 (Figure \ref{fig:genphys}).
 Table \ref{tab:fix} gives a summary of the large genomic regions that have become fixed or nearly fixed by cycle 16. 
@@
  %  \vspace{-3mm}
-   \caption{Heterozygosity ($H$) across chromosomes 9 (left) and 4 (right) in each cycle of BSSS plotted on the physical (top) or genetic (bottom) map.
-H is calculated on 15-marker sliding windows, with 5 marker steps between each calculation.
+   \caption{Heterozygosity ($H$) in each cycle across chromosome 9 of the BSSS (left) and chromosome 4 of the BSCB1 (right) plotted on the physical (top) and genetic (bottom) map.
+$H$ is calculated on 15-marker sliding windows, with 5 marker steps between each calculation.
 Each data point is color-coded based on a linear transformation of recombination rate (red = low recombination rate).
 Shaded regions represent 2 cM windows found to have heterozygosity values significantly lower than expected by simulation at a given cycle.
-Light shading denotes significance at P < 0.0025, whereas the darker shading represents significance at P < 0.001.
+Light shading denotes significance at $P<0.0025$, whereas the darker shading represents significance at $P<0.001$.
 Comparing the physical and genetic maps reveals that the regions of fixation are extremely small genetically, but some contain much of the physical content of the chromosome.
@@
 A central issue for these maize populations and others like them is whether the changes observed at the molecular level are caused directly by selection on phenotype, or indirectly due to the genetic drift that selection imposes through inbreeding and small effective population sizes.
-To gauge the roles of selection and drift, we conducted simulations of the crossing and selection schemes used in the RRS experiment, with genetic recombination modeled on a framework map of a cross between B73 (a BSSS derived line used to construct the reference genome) with Mo17 (an inbred with some co-ancestry to both the BSSS and BSCB1 founder germplasm)\citep{gerdes1993compilation}\citex{Lee et al., 2002}.
-Selections were executed at random in each simulation, so the patterns observed across simulations represent the expected distribution of effects caused only by recombination and genetic drift. 
-We conducted 10,000 simulations, sampled individuals from each cycle equal in number to the observed samples, measured heterozygosity, and compared the results to the observed data.
+To gauge the roles of selection and drift, we conducted simulations of the crossing and selection schemes used in the RRS experiment.
+Selection was executed at random in each simulation, so the patterns observed across simulations represent the expected distribution of effects caused only by recombination and genetic drift. 
+We conducted 10,000 simulations, with genetic recombination modeled on a framework map derived from the IBM population \citep{lee2002expanding}, a cross between inbred lines B73 (a BSSS derived line used to construct the reference genome) and Mo17 (an inbred with some co-ancestry to both the BSSS and BSCB1 founder germplasm).
+We sampled individuals from each cycle equal in number to the observed samples, measured heterozygosity, and compared the results to the observed data.
 
 Averaged across the genome, the vast majority of the reduction in diversity observed in both populations can be attributed to genetic drift. 
-Nonetheless, do observe deviations from the simulated values (Figure S1). 
+Nonetheless, we do observe deviations from the simulated values (Figure \ref{fig:s1}). 
 The observed data show higher than expected heterozygosity at cycle 4, suggesting selection for heterozygosity in early cycles, a deviation from simulated and actual breeding practices, or an under-sampling of the diversity present in the original cycle 0 populations.

### 2015-05-10 sup1.tex
notes in intro, pasted supp. figs, cleaned up supp. tables s2-s4, fixed title

@@
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
+
 \begin{figure}[tb]   
   \begin{center}
-   \vspace{-0mm}
-   \includegraphics[width=0.4\textwidth]{FigS1}
-   \renewcommand{\baselinestretch}{0.9}
-   \vspace{-3mm}
+   \includegraphics[width=\linewidth]{figs1_jri}
    \caption{Heterozygosity in each population, observed vs.
@@
 }
-\vspace{-6mm}
     \label{fig:s1}
@@
 
-\begin{figure}[tb]   
+\pagebreak
+
+\begin{figure*}[tb]   
   \begin{center}
-   \vspace{-0mm}
-   \includegraphics[width=0.4\textwidth]{S2}
-\includegraphics[width=0.4\textwidth]{S2_genetic}
-   \renewcommand{\baselinestretch}{0.9}
-   \vspace{-3mm}
-   \caption{Heterozygosity ($H$) at cycle16 across all ten chromosomes in each population, calculated on 15-marker sliding windows with 5 marker steps as in Figure 3.
+\includegraphics[width=\linewidth]{figs2_jri}
+   \caption{Heterozygosity ($H$) at cycle 16 across all ten chromosomes in each population, calculated on 15-marker sliding windows with 5 marker steps as in Figure 3.
 Heterozygosity values in BSSS (blue dots) and BSCB1 (red dots) are superimposed in one panel.
@@
 }
-\vspace{-6mm}
-    \label{fig:s1}
+    \label{fig:s2}
   \end{center}
-\end{figure}
+\end{figure*}
 
@@
 \hline
-	\  & \  & \  & \  \\ \hline
-	SNPS polymorphic in cycle: & But not cycle: & in BSSS & in BSCB1 \\ \hline
+	\textbf{SNPs polymorphic in cycle} & \textbf{But not cycle} & \textbf{BSSS} & \textbf{BSCB1} \\ \hline
 	4 & 0 & 290 & 225 \\ \hline
@@
 	16 & Founders & 445 & 460 \\ \hline
-	 &  &  &  \\ \hline
-	Heterozygosity (H) of SNPS present in cycle 8 but not cycle 4 &  &  &  \\ \hline
-	Statistic & BSSS & BSCB &  \\ \hline
-	Mean & 0.052 & 0.052 &  \\ \hline
-	Median & 0.028 & 0.028 &  \\ \hline
+\end{tabular}
 		    \label{tab:s2}  % caption is needed to make this work
+\end{table}
+\newpage
+
+\begin{table}
+\caption{Heterozygosity of SNPs in cycle 8 but not cycle 4 of the BSSS. \jri{are these numbers supposed to be the same? I checked the Pages doc and they are the same there too.}}
+\begin{tabular}{  l | l | l | }
+\multicolumn{1}{l}{} & \multicolumn{1}{l}{BSSS} & \multicolumn{1}{l}{BSSS}   \\ \cline{2-3}
+	Mean & 0.052 & 0.052   \\ \cline{2-3}
+	Median & 0.028 & 0.028   \\ \cline{2-3}
 \end{tabular}
 \end{table}
-\newpage
 
@@
 \hline
-	\  & \  & \  & \  & \  & \  \\ \hline
-	Simulated 'Hybrid' & Derived Lines Used as Priors & Population & Possible Switches & Switch errors & Rate \\ \hline
+	\textbf{Simulated 'Hybrid'} & \textbf{Derived Lines Used as Priors} & \textbf{Population} & \textbf{Possible Switches} & \textbf{Switch errors} & \textbf{Rate} \\ \hline
 	B11xB67 & all & BSSS & 12195 & 41 & 0.003 \\ \hline
@@
 	B99xB97 & all & BSCB1 & 6779 & 52 & 0.008 \\ \hline
-	BB42xB54 & cycle 0 & BSCB1 & 11260 & 57 & 0.005 \\ \hline
+	BB42xB54 & cycle 0 & BSCB1 & 11260 & 57 & 0.005 \\\hline
+\end{tabular}
 	\label{tab:s3}  % caption is needed to make this work
-\end{tabular}
 \end{table}

### 2015-05-14 Fig4/fig4.R
all figures with simulation data updated

@@ -1,0 +1,100 @@
+#!/usr/bin/Rscript
+
+rm(list=ls())
+library(ggplot2)
+library(dplyr)
+
+loadcycle = function(cycle, mychrom){
+  infile = paste("../Fig3/data/H_cycle", cycle, "_LG", mychrom, ".txt", sep='')
+  myfile = read.table(infile, header=T)
+  return(myfile)
+}
+  
+loadchrom = function(chrom){
+  cycles = c(0,4,8,12,16)
+  tables = lapply(cycles, loadcycle, mychrom=chrom) 
+  results = do.call(rbind, tables)
+  return(results)
+}
+
+chromdata_H = lapply(1:10, loadchrom)
+
+simresults = read.table("../data/2cm_sim_collated_results.txt", header=T)
+
+
+#The code to make plots is ugly because ggplot has some unpredictable (at least to me) behavior when wrapped in functions due to scope issues.
+####code to make plot for chromosome 9 BSSS.  ####
+
+  Chrom=9 
+  Hetgrp="BSSS"
+  Hdata = chromdata_H[[Chrom]]
+  mytitle = paste("Chromosome ", Chrom, ", ", Hetgrp, sep="" )
+  
+  sigregions001 = filter(simresults, Sigat001 == "BSSS" | Sigat001 == "Both") %>% filter(LG==Chrom)
+  sigregions0025 = filter(simresults, Sigat0025 == "BSSS" | Sigat0025 == "Both") %>% filter(LG==Chrom)
+	sigat001rect_phys <-  geom_rect(data=sigregions001, aes(xmin=phystart/1000000, xmax=phystop/1000000, ymin=0, ymax=0.5), fill="lightskyblue1") 
+	sigat0025rect_phys <- geom_rect(data=sigregions0025, aes(xmin=phystart/1000000, xmax=phystop/1000000, ymin=0, ymax=0.5), fill="lightskyblue") 
+	Hplot_phys <- geom_point(size = 1, data=Hdata, aes(x=physpos/1000000, y=HetSS, colour=sqrt(rate^(1/4)))) 
[…]

### 2015-05-14 intro.tex

@@
 \section*{Introduction}
-Although maize is naturally an out-crossing organism, modern breeding develops highly inbred lines that are then used in controlled crosses to produce hybrids. 
-Hybrid maize has been so successful that it quickly replaced long-standing mass-selected open pollinated varieties \citep{crabb1947hybrid}. 
-Maize inbred lines are now partitioned into separate inbreeding ‘heterotic groups’ that maximize performance and hybrid vigor (heterosis) when an inbred line of one group is crossed with the other group \citep{tracy2006historical}.
-The shift towards selection of inbred lines based on their ability to generate good hybrids – referred to as ‘combining ability’ – constituted an abrupt change from the open-pollinated mass selection that breeders practiced for millennia \citep{Anderson, 1944; troyer1999background}.
-Multiple studies with molecular markers have indicated that the modern era of single-cross hybrid maize breeding has led to a dramatic restructuring of population genetic variation \citep{duvick2004long, ho2005extent, feng2006temporal}.
-Different heterotic groups have diverged genetically over time to become highly structured and isolated populations. 
+Although maize is naturally an outcrossing organism, modern breeding takes advantage of highly inbred lines that are used in controlled crosses to produce hybrids. 
+Hybrid maize, first developed in the early 20th century by \jri{East and Shull?}, was so successful that within a few decades it had completely replaced mass-selected open pollinated varieties in the United States \citep{crabb1947hybrid}.
+The shift towards selection of inbred lines based on their ability to generate good hybrids – referred to as ‘combining ability’ – constituted an abrupt change from the open-pollinated mass selection that breeders practiced for millennia \citep{anderson1944sources, troyer1999background}.
+Maize inbred lines are now partitioned into separate ‘heterotic groups’ that maximize performance and hybrid vigor (heterosis) when inbreds from different heterotic groups are crossed with each other \citep{tracy2006historical}.
+
+Multiple studies with molecular markers have indicated that these heterotic groups have diverged genetically over time to become highly structured and isolated populations, resulting in a dramatic restructuring of population genetic variation \citep{duvick2004long, ho2005extent, feng2006temporal}.
 Advances in high throughput genotyping and the development of a maize reference genome now enable the observation of maize population structure at high marker density across the whole genome \citep{ganal2011a-large,chia2012maize}.
-So far, these high-density studies have examined a broad spectrum of germplasm at various points in the history of maize to search for the signals of population structure and artificial selection \citep{Hufford2012b, van2012historical}.
+These studies have examined a broad spectrum of germplasm at various points in the history of maize to search for the signals of population structure and artificial selection \citep{Hufford2012b, van2012historical}.
 \jri{add more}
@@
 The lack of distinct selection signals from modern breeding may be due to specific selective events occurring in different populations, necessitating a more focused look within heterotic groups or even single breeding programs.
-In this study, we apply a genomic approach to study the dynamics of genetic variation over time within an individual selection experiment. 
+\jri{add text about golden glow, krug, and lamkey}
+
+Here we take a detailed look at the dynamics of genome-wide patterns of diversity over time within an individual breeding program. 
 The Iowa Stiff Stalk Synthetic (BSSS) and the Iowa Corn Borer Synthetic No.
@@
 The Iowa RRS experiment proves especially relevant because lines derived from the BSSS population have had a major impact upon the development of commercial hybrids \citep{duvick2004long, darrah19861985}, the formation of modern heterotic groups \citep{troyer1999background; senior1998utility}, and the choice of a maize reference genome \citep{schnable2009the-b73-maize}.
+\jri{something about heterosis}

### 2015-06-10 main.tex
text, some refs, messed up biblio
supp figures, tables
submodules are a pain.
ref tweaks

@@
 \usepackage{hyperref}
+\usepackage{multirow}
 
@@
 \newcommand{\jpg}[1]{\textcolor{mattgreen}{ \em{\scriptsize  #1}} }
+\newcommand{\citex}[1]{\textcolor{red}{\bf #1}}
 
@@
 \setcounter{footnote}{1}
-%
-%\title{The genomic impacts of drift and selection for hybrid performance in maize}
-%Corresponding author:  Department of Plant Sciences, University of California, Davis, California 95616, USA. 
-%    E-mail: \mbox{rossibarra@ucdavis.edu}}\\[0.3cm]
-%   \small\sf $^{\ast}$Division of Plant Sciences, University of Missouri, Columbia MO 65211,\\
-%   \small\sf $^\dag$Corn Insects and Crop Genetics Research Unit, USDA-Agricultural Research Service, Ames, IA, 50011\\
-%   \small\sf $^\ddag$Plant Genetics Research Unit, USDA-Agricultural Research Service, Columbia MO 65211,\\
-%   \small\sf $^\S$Department of Plant Sciences, The Center for Population Biology and the Genome Center, University of California, Davis, California 95616 , USA,\\
-%   \small\sf $^1$ Present address:  DuPont Pioneer, Johnston, Iowa
-%}
-   %\small\sf  Present address: Graduate University for Advanced Studies, Hayama, Kanagawa 240-0193, Japan\\
-%}
 
-\title{Independent molecular basis of convergent highland adaptation in maize}
+\title{The genomic impacts of drift and selection for hybrid performance in maize}
 \author{
@@
 To investigate the genomic impact of this selection program, we used the Illumina MaizeSNP50 high-density SNP array to determine genotypes of progenitor lines and over 600 individuals across multiple cycles of selection.
-Consistent with previous research (Messmer et al., 1991; Labate et al., 1997; Hagdorn et al., 2003; Hinze et al., 2005), we found that genetic diversity within each population steadily decreases, with a corresponding increase in population structure.
+Consistent with previous research \citep{messmer1991genetic, labate1997molecular,  hagdorn2003molecular, hinze2005population}, we found that genetic diversity within each population steadily decreases, with a corresponding increase in population structure.
 High marker density also enabled the first view of haplotype ancestry, fixation and recombination within this historic maize experiment. 
@@
 Signatures of selection were difficult to observe in the background of this strong genetic drift, but heterozygosity in each population has fallen more than expected. 
-Regions of haplotype fixation represent the most likely targets of selection, but as observed in other germplasm selected for hybrid performance (Feng et al., 2006), there is no overlap between the most likely targets of selection in the two populations.
+Regions of haplotype fixation represent the most likely targets of selection, but as observed in other germplasm selected for hybrid performance \citep{feng2006temporal}, there is no overlap between the most likely targets of selection in the two populations.
 We discuss how this pattern is likely to occur during selection for hybrid performance, and how it poses challenges for dissecting the impacts of modern breeding and selection on the maize genome. }
@@
 
-\bibliography{references/references.bib}
+\bibliography{references.bib}
 \bibliographystyle{geneticsT2}
 
-%\suppl
-
-%%%%%%%
-% put supplemental figures and tables here:
-% e.g.
-% \begin{figure}
-% ...
-
-%%%%%
-% appendices to follow
-%\input{HighLowSup1}
+\suppl
+\input{sup1}
 

### 2015-06-10 results.tex
ref tweaks

@@
 No new genetic material was intentionally introduced into either population after the experiment's inception, so the substantial increase in genetic distance could only arise from the loss of genetic diversity within each population.
-Consistent with previous studies of the Iowa RRS, genome-wide genetic diversity (expected heterozygosity, H) decreases steadily across cycles of selection in both populations (Figure \ref{fig:decline}).
-\jri{need a citation here of "previous studies"?}
+Consistent with previous studies of the Iowa RRS \citep{messmer1991genetic, labate1997molecular,  hagdorn2003molecular, hinze2005population}, genome-wide genetic diversity (expected heterozygosity, H) decreases steadily across cycles of selection in both populations (Figure \ref{fig:decline}).
+\jri{do all these citations make sense here?
+I just copied these from the abstract}
 The loss of heterozygosity is smaller when the two populations are considered together, indicating the loss of different alleles within BSCB1 and BSSS. 

### 2015-06-10 intro.tex
intro edits done

@@
 These studies have examined a broad spectrum of germplasm at various points in the history of maize to search for the signals of population structure and artificial selection \citep{Hufford2012b, van2012historical}.
-\jri{add more}
 Although selective sweeps remaining from domestication are clearly visible, the impact of selection during modern breeding appears comparatively small in terms of its impact on genomic diversity \citep{Hufford2012b,van2012historical}, despite steady, heritable improvement in phenotype \citep{duvick2005contribution}.
-The lack of distinct selection signals from modern breeding may be due to specific selective events occurring in different populations, necessitating a more focused look within heterotic groups or even single breeding programs.
-\jri{add text about golden glow, krug, and lamkey}
+The lack of distinct selection signals from modern breeding may be due to population-specific selection; this interpretation is supported by the measured success in identifying targets of selection in individual experimental populations selected for a limited number of phenotypes \citep{hirsch2014insights, beissinger2014genome}.
 

### 2015-06-11 discussion.tex
intro edits done
jri DONE

@@
 
-In the Iowa RRS experiment, the BSSS and BSCB1 populations steadily lost genetic diversity as they became more differentiated from one another, and principal component analysis shows that as the effective population size and the rates of inbreeding were altered, the rates of change in population structure were altered as well.
+Our analysis of diversity in the Iowa RRS experiment reveals a steady loss of diversity in the BSSS and BSCB1 populations as they became increasingly differentiated from one another over time.
+Principal component analysis shows that as the effective population size and the rates of inbreeding were altered, the rates of change in population structure were altered as well. 
 These patterns of population structure, diversity, and differentiation between BSSS and BSCB1 can be largely reproduced by simulation without any selection, supporting the hypothesis that the majority of the genetic structure observed can be attributed to genetic drift alone, despite effective selection for phenotypic improvement.
@@
 Reciprocal recurrent selection serves as a model for the method of hybrid maize improvement employed broadly in North America \citep{duvick2004long}.
-The trends observed in the Iowa RRS are also a common theme in similar recurrent selection experiments \citep{Romay2013,lamkey2014relative}. 
+The trends observed in the Iowa RRS are also a common theme in similar recurrent selection experiments \citep{romay2012effect,lamkey2014relative}.
 Increased population structure due to increased inbreeding and lowered effective population size are visible in the modern ‘stiff-stalk’ and ‘non stiff-stalk’ heterotic groups as well \citep{van2012historical}.
@@
 Near isogenic lines that substitute these centromeric haplotypes from key inbreds could be used to determine if the haplotypes at high frequency in cycle 16 of BSSS provide a phenotypic advantage when crossed with BSCB1.
+
 Both our study and previous work \citep{labate1997molecular} identified genetic differentiation between cycle 0 of BSSS and the founder inbreds. 
@@
 However, simulated and observed values are often close, so overall the signature of selection is difficult to detect at any given locus. 
-This is due in large part to a limitation of statistical power imposed by inbreeding and low effective population sizes, though uncertainty about the fine-scale recombination map also plays a role. 
-We show that an identity-by-descent, haplotype based approach provides additional power as it can distinguish between the fixation of rare and common haplotypes. 
+This is due in large part to a limitation of statistical power imposed by inbreeding and low effective population sizes.
+\citet{lorenz2015selection} analyze a similar selection experiment with comparable effective population size and also find little little statistical evidence for selection on individual loci in spite of genetic gain for the trait of interest.
+Nonetheless, we show that an identity-by-descent, haplotype based approach provides additional power, as it can distinguish between the fixation of rare and common haplotypes. 
 However, the results of single locus simulations are sensitive to sampling error and drift caused by population maintenance. 
@@
 That study was a wide survey across many germplasm sources. 
-The limited impact of selection could arise because different haplotypes and loci are selected in different breeding programs. 
+The limited overall impact of selection could arise because different haplotypes and loci are selected in different breeding programs. 
 In support of this idea, we find that the most likely targets of selection between the BSSS and BSCB1 populations occur at non-overlapping loci. 

### 2015-06-30 Fig4/fig4.R
updated figure 4 into a single panel

@@ -35,7 +35,7 @@ simresults = read.table("../data/2cm_sim_collated_results.txt", header=T)
 	sigat001rect_phys <-  geom_rect(data=sigregions001, aes(xmin=phystart/1000000, xmax=phystop/1000000, ymin=0, ymax=0.5), fill="lightskyblue1") 
 	sigat0025rect_phys <- geom_rect(data=sigregions0025, aes(xmin=phystart/1000000, xmax=phystop/1000000, ymin=0, ymax=0.5), fill="lightskyblue") 
 	Hplot_phys <- geom_point(size = 1, data=Hdata, aes(x=physpos/1000000, y=HetSS, colour=sqrt(rate^(1/4)))) 
-	myphysplot=ggplot(data=Hdata) + facet_grid(Cycle~.) + sigat001rect_phys + sigat0025rect_phys + Hplot_phys + theme_bw() + xlab("Position(Mb)")+ ylab("Heterozygosity") + scale_colour_gradient2(low="red", mid="orange", high="yellow", midpoint=median(Hdata$rate^(1/4))) 
+	myphysplot=ggplot(data=Hdata) + facet_grid(Cycle~.) + sigat001rect_phys + sigat0025rect_phys + Hplot_phys  + xlab("Position(Mb)")+ ylab("Heterozygosity") + scale_colour_gradient2(low="red", mid="orange", high="yellow", midpoint=median(Hdata$rate^(1/4))) + ggtitle("BSSS Chromosome 9")
 
   axesopts = theme(axis.text.y=element_text(size=6), axis.text.x=element_text(size=6), axis.title.y=element_text(size=8, angle=90), axis.title.x=element_text(size=8))
 
@@ -44,7 +44,7 @@ simresults = read.table("../data/2cm_sim_collated_results.txt", header=T)
   simdata1rect_gen <-  geom_rect(data=sigregions001, aes(xmin=genstart, xmax=genstop, ymin=0, ymax=0.5), fill="lightskyblue1") 
   simdata2rect_gen <- geom_rect(data=sigregions0025, aes(xmin=genstart, xmax=genstop, ymin=0, ymax=0.5), fill="lightskyblue") 
   Hplot_gen <- geom_point(size = 1, data=Hdata, aes(x=genpos, y=HetSS, colour=sqrt(rate^(1/4)))) 
[…]

### 2015-06-30 Fig4/fig4.R
minor change to figure appearance

@@ -35,7 +35,7 @@ simresults = read.table("../data/2cm_sim_collated_results.txt", header=T)
 	sigat001rect_phys <-  geom_rect(data=sigregions001, aes(xmin=phystart/1000000, xmax=phystop/1000000, ymin=0, ymax=0.5), fill="lightskyblue1") 
 	sigat0025rect_phys <- geom_rect(data=sigregions0025, aes(xmin=phystart/1000000, xmax=phystop/1000000, ymin=0, ymax=0.5), fill="lightskyblue") 
 	Hplot_phys <- geom_point(size = 1, data=Hdata, aes(x=physpos/1000000, y=HetSS, colour=sqrt(rate^(1/4)))) 
-	myphysplot=ggplot(data=Hdata) + facet_grid(Cycle~.) + sigat001rect_phys + sigat0025rect_phys + Hplot_phys  + xlab("Position(Mb)")+ ylab("Heterozygosity") + scale_colour_gradient2(low="red", mid="orange", high="yellow", midpoint=median(Hdata$rate^(1/4))) + ggtitle("BSSS Chromosome 9")
+	myphysplot=ggplot(data=Hdata) + facet_grid(Cycle~.) + sigat001rect_phys + sigat0025rect_phys + Hplot_phys  + xlab("Position(Mb)")+ ylab("Heterozygosity") + scale_colour_gradient2(low="red", mid="orange", high="yellow", midpoint=median(Hdata$rate^(1/4))) + theme_bw() + ggtitle("BSSS, Chromosome 9")
 
   axesopts = theme(axis.text.y=element_text(size=6), axis.text.x=element_text(size=6), axis.title.y=element_text(size=8, angle=90), axis.title.x=element_text(size=8))
 
@@ -44,7 +44,7 @@ simresults = read.table("../data/2cm_sim_collated_results.txt", header=T)
   simdata1rect_gen <-  geom_rect(data=sigregions001, aes(xmin=genstart, xmax=genstop, ymin=0, ymax=0.5), fill="lightskyblue1") 
   simdata2rect_gen <- geom_rect(data=sigregions0025, aes(xmin=genstart, xmax=genstop, ymin=0, ymax=0.5), fill="lightskyblue") 
   Hplot_gen <- geom_point(size = 1, data=Hdata, aes(x=genpos, y=HetSS, colour=sqrt(rate^(1/4)))) 
[…]

### 2015-06-30 response.tex
response to review

\documentclass[onecolumn,oneside,letterpaper]{article}

\usepackage{response}
%\usepackage{geneticsT2}

\usepackage{times}
\usepackage{color}
\usepackage{hyperref}

% rayout %
\addtolength{\oddsidemargin}{-2.75cm}
\addtolength{\evensidemargin}{-0.75cm}

\addtolength{\textwidth}{5.5cm}
\addtolength{\topmargin}{-3cm}
\addtolength{\textheight}{4.5cm}

%\parindent=4em
%\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex}

\setlength{\parindent}{0pt}
\setlength{\parskip}{10.0pt}
%\renewcommand{\itemindent}{50pt}
\setlength{\leftskip}{0.5cm}
\renewcommand{\labelitemi}{$-$}

\newcommand{\jri}[1]{\textcolor{blue}{ \emph{\scriptsize  #1}} }
\newcommand{\jristop}{\textcolor{red}{\hline \bf JRI STOPPED HERE\\ \hline}}
\newcommand{\st}[1]{\textcolor{red}{ #1}}
\newcommand{\comst}[1]{\textcolor{red}{ \em{\scriptsize  (#1)}} }
\definecolor{mattgreen} {rgb} {0,0.6,0}
\newcommand{\mbh}[1]{\textcolor{mattgreen}{ \em{\scriptsize  #1}} }
\definecolor{peterpurple}{rgb}{.6,0,.6}
\newcommand{\plr}[1]{\textcolor{peterpurple}{ \emph{\scriptsize (#1)}} }

\renewcommand{\textfraction}{0.01}
\renewcommand{\topfraction}{0.99}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.90}
\renewcommand{\dbltopfraction}{0.95}
\renewcommand{\dblfloatpagefraction}{0.80}
\renewcommand{\sfdefault}{phv}

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\fancyfoot[CE,CO]{\thepage}
\renewcommand{\headrulewidth}{0pt}
\fancypagestyle{plain}{
	\fancyhf{}
}

\usepackage{natbib}
\bibpunct{(}{)}{;}{a}{}{,}

% space of double hline in Table
\doublerulesep = 0.4pt

\title{Response to Reviews}

\usepackage{natbib}
\bibpunct{(}{)}{;}{a}{}{,}

\usepackage{amsmath}

\usepackage{graphicx}

\begin{document}

\maketitle

Stephen's comments:

The reviewers make a number of important points that should be considered in revision. Of particular note, Reviewer 1 highlights the importance of considering the possible effects of purging during the experiment, and this intersects with reviewer 2's comments about lack of details of the experiment and what was being selected, and what the outcome of the experiment was. Could purging, for example, have occurred over the course of the experiment, and did yield improve within lines in addition to increased yield in hybrids? Reviewer 2 also notes a lack of details on the simulation approach; I agree with this, and also thought this was perhaps confusing wording on page 14: "Selections were executed at random".

On the statistical side, candidate selected regions are picked out and effectively 1-sided tests are reported for the rare allele that has increased in frequency. While the authors are reasonably cautious about these cases anyway, at least some attempt to report an experiment-wide p value or FDR for these regions seems useful.

Do the results provide sufficient power to reject the overdominance model? How strong would the signal be likely to be?

One additional comment: the authors note that heterozygosity increases early in the experiment, and discuss various possible explanations for this. But couldn't directional selection on rare alleles be the driver of this, i.e. early on there will be a rise in allele frequencies, increasing diversity, and then subsequently a loss due to fixations?

Reviews:
Reviewer 1 Reviewer's Comments to Author...
I really enjoyed reading this manuscript.  I think this is a well done study 
that provides novel insight in to basic processes in breeding of maize.  Many of 
these results have implications for other species as well.  I have a number of 
minor comments below that I think should be addressed but these do not diminish 
my enthusiasm for the manuscript.

The abstract could use some editing.  There are a number of references in this 
abstract and the abstract is not well focused on the specific findings of this 
study.

In the introduction the authors describe the concept of heterotic groups and 
state that these lines maximize performance and heterosis when crossed.  I would 
caution that this should be clarified to heterosis for yield rather than for all 
traits. Research by other groups (ie Flint-Garcia et al 2009) suggests that 
heterosis can vary for different traits and that the definition of heterotic 
groups is primarily based on yield heterosis, not heterosis for other traits.

The introduction references several studies that looked at genome-wide 
population structure in breeding programs from the Ross-Ibarra group.  I 
hesitate to even make this suggestion (do to potential flaws) but it might be 
appropriate to also include Jiao et al nat Genetics 2012 here as this study also 
looked at a large dataset of SNPs in modern maize.

I was quite intrigued by the difference in distance between cycles 4 and then 8-
16.  The authors suggest that this could be due to some differences in genetic 
maintenance of the populations.  I am curious if there were any particular 
environmental influences during that time period that may have led to stronger 
selection or if there were any differences in agronomic practices that have 
occurred during the long-term experiment (ie irrigation, fertilization, plant 
density, etc).  it would be helpful to know if any of these factors may have 
contributed.

It was a bit surprising to me that there was no differentiation among the 
founder inbreds of the two populations.  If PCA was run only on the founders 
(without any other lines) do you get separation of the two groups or not?

Figure 3 is quite useful and helps to show large physical regions that have lost 
heterozygosity.  The authors discuss the physical-genetic difference.  It would 
be useful to have a supplemental figure that shows these same values (cycle 16 
heterozygosity on y-axis) versus genetic position on x-axis.  I am curious if 
there are ever regions of multiple cM on chromosome arms that become fixed.  
This would be a nice complementary figure.  These are provided for some 
chromosomes in 4B and 5B but would be useful to see genome wide.

The large regions of fixation are quite intriguing.  Is there any evidence for 
major flowering time or disease resistance QTL in these regions?

I am somewhat baffled as to why the Fig S1 was put in supplemental materials.  A 
central aspect of this study is this issue of genetic drift relative to 
selection and there are two pages of results text for this section.  However, 
this useful figure is only in supplemental materials.  I would strongly 
encourage moving this figure to the actual manuscript.  If space is an issue 
then only use figu 4 or fig 5 as these are quite redundant with each other in 
terms of conceptual views. 
The legend to fig S1 could use additional detail.  I am guessing that the black 
lines are 95% CI but this is not clear.  Please explain the black lines.

Figure S2 is not ever referenced in the text as far as I could find.  The data 
table of the same results is referenced but I don’t think the figure is.

The last paragraph of the results section assesses the fixation of particular 
haplotypes over time.  I think it could be very useful to visualize this data 
for each of the large blocks.  This could be done by using a graph of allele 
frequency over time for each region.  Right now these results are presented in 
text form for only two regions but it would be very interesting to see these 
results for all eight regions shown in table 1.  What is the fixation at each 
cycle in the observed data and how often was this region fixed in simulations.  
I appreciate the authors point that these fixations do occur at low rates in 
simulations and since they are studying a genome-wide pattern a number of low 
frequency events are expected to occur by chance.

The authors interpret the rate of heterozygosity in terms of selection and 
drift.  I am assuming that they are interpreting selection to mean the selection 
for advantageous alleles in hybrid test combinations.  However, I wonder about 
the inadvertent selection to maintain heterozygosity?  In work from McMullen and 
Buckler it was argued that the gain in vigour due to heterozygosity of 
pericentromeric regions was larger than expected and that this may be due to 
increased ability to cross and maintain these plants.  Was this concept included 
in the modeling of the data?  In other words, while drift will tend to increase 
homozygosity in a population there may be counteracting forces of heterosis 
effects that will maintain heterozygosity in the same population.  Just curious 
about this concept.

The authors spend some time in the discussion regarding models of heterosis and 
extrapolation of these results to discuss selection.  While this manuscript 
describes materials that are quite relevant for maize breeding programs I think 
it is important to note that these types of materials are not universal.  In 
this case the synthetic populations are derived from inbred lines.  This is 
important in terms of the nature of deleterious alleles that remain.  Strongly 
deleterious alleles have likely been purged from the founders as each founder 
was an inbred line that had been selected for some level of vigor.  This means 
that the remaining single-locus deleterious alleles will all be of quite minor 
effect and therefore selection is likely to act on alleles of minor effect 
rather than major effect deleterious alleles.  I think that is would be useful 
to remind readers of this attribute of the population.  My guess is that if this 
experiment was done with a maize (or teosinte) population that had never been 
passed through an inbred “bottleneck” that there was be some alleles of major 
effect that would be purged quickly and that the role of drift and selection 
might be slightly different.

Reviewer 2 Reviewer's Comments to Author...
Girke et al. in "The genomic impacts of drift and selection for hybrid performance in maize" report a 
retrospective examination of genomic patterns of diversity in a maize experimental population. The data 
set consists of genotyping of ~38,000 SNPs genotyped in individuals from various phases of the selection 
experiment. While the manuscript consists primarily of population genetic analysis, the authors attempted 
to better integrate a popular maize genetic map with the current reference genome.

The strength of the manuscript is that the authors design simulations that closely parallel the experimental 
population design to determine exceptations for change in heterozygsosity and allele frequency 
differentiation. The simulations are very useful in that clearly establish a major role of genetic drift, owing 
in large part to small effective population size, in the differentiation of experimental populations. There are 
a couple of points of clarity needed with regard to the simulations. First, it isn't clear how recombination 
was handled within the simulation, other than an indication that "all recombinations events in the 
simulations were carried out in R with the hyped software package…" (p 11). Presumably some number (0 - 
2?) recombination events per arm generation was included in the simulation, with probability of 
recombination at a location determine by genetic map length. Something like this needs to be stated. Also 
related to the simulations, there are multiple references in the text to empirical results as "deviation from 
simulation" (e.g., p 15) when I suspect what the authors wish to say is that the results differ from neutral 
expectations based on simulations.

The other major weakness of the paper is that relatively little background on the experimental populations 
is provided. There is a long list of citations of prior work on this experimental population, but little 
exposition. A better presentation of the specific target of selection is needed (I infer that selection is on 
improved yield through increased hybrid performance). Ultimately in the Discussion, there is additional 
discussion of the phenotypes, but the presentation is minimal (again pointing to multiple citations) and 
provides little insight into the nature of the selection observed. Better discussion of phenotypes potentially 
targeted by selection is essential if the manuscript is to end, as it does, with a call for further dissection of 
targets of selection in this population.

A third major issue that manuscript seems to be not at all informed by multiple experimental evolution 
studies conducted on Drosophila and other model systems. This contributes to a couple of awkward 
citations of personal communication (p 17 and 19), where ideas about loss of diversity and need for 
replication of experimental populations have been reported in the primary literature. See Drosophila papers 
by Tom Turner and Anthony Long and colleagues.

Finally, the Discussion seems a bit overly long and could potentially be more tightly written.

Minor points: 
- Careful copy editing is needed. Heterozygsoity in the figures and figure captions is referred to variously 
as "H" with and without italics and also as heterozygosity spelled out, sometimes within th esame figure or 
figure caption. Simiarily, Fst needs to be presented in a consistent manner. Also edit citations for 
consistent capitalization and in=text citations for consistent spacing. Also:
On page 7, line 10: A sentence should not begin with a number.
Page 10, line 14: Citation format is incorrect. 
Page 11, line 16: Number of SNPs passing filters should be reported in ‘Materials and Methods’
Figure 2 caption: What is “H combined” and how was it calculated?
Figure 4 and 5: Please include the simulated data (mean + variance) so the pattern is more apparent.

- Figures could be better labeled, for example, simply label panels A & B of Figure 4  as "Genetic Map and 
"Physical Map" and the cycles as "Cycle."

- When discussing the plant material that was used in the analysis (page 7, line 13), two individuals that 
appeared to be mislabeled had their labels switched. However, later in the paragraph, another labeling 
issue was discovered, and these individuals were removed from the analyses. What is the rationale behind 
the differential treatment for these two labeling issues?

- For the integration of the physical and genetic maps, there are two concerns:
1) Do the SNPs on the Maize 50K SNP chip already have reference positions? How were the SNP contextual 
sequences mapped back to the reference genome? If it was a BLAST-type tool, was there an e-value 
threshold that was used to remove low-confidence positions?

2) When integrating the genetic and physical maps, how many SNPs occurred in “mis-mapped blocks?”

- When the cycles of selection shift from 10 individuals to 20 individuals, does this really constitute a 
doubling in effective population size?

- A subheading on heterosis would help better organize the Discussion.

- The Hudson et al. 1992 measure of Fst that the authors report is a sequence-based measure based on 
average pairwise diversity within and between loci. How were the "loci" defined here?

- On page 6, what is the difference in a "found inbred" and a "founder line?" I suspect this distinction will 
be lost on many readers.

- On page 15."consistent with strong genetic drift imposed by selection practices," needs to be reworded.

In summary, this is an interesting manuscript that would benefit from improved clarity of presentation.

\bibliography{references}
\bibliographystyle{geneticsT2}

\end{document}

### 2015-06-30 response.tex
starting response

@@
 
-Stephen's comments:
+\section*{Editor's comments}
 
@@
 
+\response{blah}
+
 Do the results provide sufficient power to reject the overdominance model? How strong would the signal be likely to be?
+
+\response{blah}
 
@@
 
-Reviews:
-Reviewer 1 Reviewer's Comments to Author...
+\response{blah}
+
+\section*{Reviews}
+\subsection*{Reviewer 1}
 I really enjoyed reading this manuscript.  I think this is a well done study 
@@
 
+\response{blah}
+
 In the introduction the authors describe the concept of heterotic groups and 
@@
 
+\response{blah}
+
 The introduction references several studies that looked at genome-wide 
@@
 looked at a large dataset of SNPs in modern maize.
+
+\response{While \citet{jiao2012genome} have decided against including citation of the Jiao et al paper, mostly because we are unsure of the veracity of the results (see \citet{lorenz2013phylogenetic}) and concerned by the fact that a wholesale recoding of their genotyping data appeared to have little impact on their conclusions.}
 
@@
 
+\response{blah}
+
 It was a bit surprising to me that there was no differentiation among the 
@@
 (without any other lines) do you get separation of the two groups or not?
+
+\response{blah}
 
@@
 
+\response{blah}
+
 The large regions of fixation are quite intriguing.  Is there any evidence for 
 major flowering time or disease resistance QTL in these regions?
+
+\response{blah}
 
@@
 The legend to fig S1 could use additional detail.  I am guessing that the black 
-lines are 95% CI but this is not clear.  Please explain the black lines.
+lines are 95\% CI but this is not clear.  Please explain the black lines.
+
+\response{blah}
 
@@
 table of the same results is referenced but I don’t think the figure is.
+
+\response{blah}
 
@@
 
+\response{blah}
+
 The authors interpret the rate of heterozygosity in terms of selection and 
@@
 about this concept.
+
+\response{blah}
 
@@
 
-Reviewer 2 Reviewer's Comments to Author...
-Girke et al. in "The genomic impacts of drift and selection for hybrid performance in maize" report a 
-retrospective examination of genomic patterns of diversity in a maize experimental population. The data 
+\response{blah}
+
+\subsection*{Reviewer 2}
+Girke et al.
+in "The genomic impacts of drift and selection for hybrid performance in maize" report a retrospective examination of genomic patterns of diversity in a maize experimental population.
+The data
 set consists of genotyping of ~38,000 SNPs genotyped in individuals from various phases of the selection 
@@
 
+\response{blah}
+
 The other major weakness of the paper is that relatively little background on the experimental populations 
 is provided. There is a long list of citations of prior work on this experimental population, but little 
-exposition. A better presentation of the specific target of selection is needed (I infer that selection is on 
+exposition.
+
+\response{blah}
+
+ A better presentation of the specific target of selection is needed (I infer that selection is on 
 improved yield through increased hybrid performance). Ultimately in the Discussion, there is additional 
@@
 targets of selection in this population.
+
+\response{blah}
 
@@
 
+\response{blah}
+
 Finally, the Discussion seems a bit overly long and could potentially be more tightly written.
+
+\response{blah}
 
@@
 
-- Figures could be better labeled, for example, simply label panels A & B of Figure 4  as "Genetic Map and 
+\response{blah}
+
+Figures could be better labeled, for example, simply label panels A \& B of Figure 4  as "Genetic Map and 
 "Physical Map" and the cycles as "Cycle."
 
-- When discussing the plant material that was used in the analysis (page 7, line 13), two individuals that 
+When discussing the plant material that was used in the analysis (page 7, line 13), two individuals that 
 appeared to be mislabeled had their labels switched. However, later in the paragraph, another labeling 
@@
 
-- For the integration of the physical and genetic maps, there are two concerns:
+\response{blah}
+
+For the integration of the physical and genetic maps, there are two concerns:
 1) Do the SNPs on the Maize 50K SNP chip already have reference positions? How were the SNP contextual 
@@
 
+\response{blah}
+
 2) When integrating the genetic and physical maps, how many SNPs occurred in “mis-mapped blocks?”
 
-- When the cycles of selection shift from 10 individuals to 20 individuals, does this really constitute a 
+\response{blah}
+
+When the cycles of selection shift from 10 individuals to 20 individuals, does this really constitute a 
 doubling in effective population size?
 
-- A subheading on heterosis would help better organize the Discussion.
+\response{blah}
 
-- The Hudson et al. 1992 measure of Fst that the authors report is a sequence-based measure based on 
+A subheading on heterosis would help better organize the Discussion.
+
+\response{blah}
+
+The Hudson et al. 1992 measure of Fst that the authors report is a sequence-based measure based on 
 average pairwise diversity within and between loci. How were the "loci" defined here?
 
-- On page 6, what is the difference in a "found inbred" and a "founder line?" I suspect this distinction will 
+\response{blah}
+
+On page 6, what is the difference in a "found inbred" and a "founder line?" I suspect this distinction will 
 be lost on many readers.
 
-- On page 15."consistent with strong genetic drift imposed by selection practices," needs to be reworded.
+\response{blah}
+
+On page 15."consistent with strong genetic drift imposed by selection practices," needs to be reworded.
+
+\response{blah}
 

### 2015-07-02 FigS1/BSCB.R
combined figure S1

@@ -19,13 +19,14 @@ getbounds = function(x){
 bounds = plyr::ddply(.data=melted_CB_all, .variables=c("Cycle", "Pruning"), .fun=getbounds)
 colnames(bounds) = c("Cycle", "Pruning", "one", "ninenine")
 realdata = realdata[realdata$Pruning=="all",]
-png("../FigS1B.png", width=400,height=300)
+#png("../FigS1B.png", width=400,height=300)
 myplot = ggplot()
 myplot = myplot + geom_point(data=melted_CB_all, aes(x=Cycle, y=H), alpha=0.1) + geom_line(data=bounds, aes(x=Cycle, y=one, group=Pruning), color="black") + geom_line(data=bounds, aes(x=Cycle, y=ninenine, group=Pruning), color="black")
 myplot = myplot + geom_point(data=medians, aes(x=Cycle, y=Med), color="green") 
 myplot = myplot + geom_line(data=realdata, aes(x=Cycle, y=H, group=Pruning), color="red") + geom_point(data=realdata, aes(x=Cycle, y=H), color="red") + theme_bw() + ggtitle("BSCB1 Genome-wide Simulation")
-myplot + theme(axis.title.x=element_text(size=0))
-dev.off()
+plot_BSCB1 = myplot + theme(axis.title.x=element_text(size=0)) + scale_y_continuous(limits=c(0.1,0.3))
+save(plot_BSCB1, file="plot_BSCB1.rds")
+#dev.off()
 
 # png("BSCB_sims_gray.png", width=1000,height=600)
 # test = ggplot()

### 2015-07-02 FigS1/BSSS.R

@@ -7,27 +7,28 @@ library(reshape)
 
 simresults <- read.table("../data/BSSS_fullsim_HH.txt")
 colnames(simresults) = c("Cycle4", "Cycle8", "Cycle12", "Cycle16")
-melted_CB_all = melt(simresults)
-colnames(melted_CB_all) = c("Cycle", "H")
-melted_CB_all$Pruning = "all"
+melted_SS_all = melt(simresults)
+colnames(melted_SS_all) = c("Cycle", "H")
+melted_SS_all$Pruning = "all"
 
-colnames(melted_CB_all) = c("Cycle",  "H", "Pruning")
-medians = plyr::ddply(.data=melted_CB_all, .variables=c("Cycle", "Pruning"), .fun=function(x){median(x[,2])})
+colnames(melted_SS_all) = c("Cycle",  "H", "Pruning")
+medians = plyr::ddply(.data=melted_SS_all, .variables=c("Cycle", "Pruning"), .fun=function(x){median(x[,2])})
 colnames(medians) = c("Cycle", "Pruning",  "Med")
 realdata = read.table("BSSS_realresults.txt", header=T)
 getbounds = function(x){
   quantile(x[,2], probs=c(0.01, 0.99))
 }
-bounds = plyr::ddply(.data=melted_CB_all, .variables=c("Cycle", "Pruning"), .fun=getbounds)
+bounds = plyr::ddply(.data=melted_SS_all, .variables=c("Cycle", "Pruning"), .fun=getbounds)
 colnames(bounds) = c("Cycle", "Pruning", "one", "ninenine")
 realdata = realdata[realdata$Pruning=="all",]
-png("../FigS1A.png", width=400,height=300)
+#png("../FigS1A.png", width=400,height=300)
 myplot = ggplot()
-myplot = myplot + geom_point(data=melted_CB_all, aes(x=Cycle, y=H), alpha=0.1) + geom_line(data=bounds, aes(x=Cycle, y=one, group=Pruning), color="black") + geom_line(data=bounds, aes(x=Cycle, y=ninenine, group=Pruning), color="black")
+myplot = myplot + geom_point(data=melted_SS_all, aes(x=Cycle, y=H), alpha=0.1) + geom_line(data=bounds, aes(x=Cycle, y=one, group=Pruning), color="black") + geom_line(data=bounds, aes(x=Cycle, y=ninenine, group=Pruning), color="black")
[…]

### 2015-07-02 FigS1/combineplot.R

@@ -1,0 +1,14 @@
+rm(list=ls())
+library(ggplot2)
+library(cowplot)
+
+
+load("plot_BSCB1.rds")
+load("plot_BSSS.rds")
+
+combined = plot_grid(plot_BSSS, plot_BSCB1, labels=c("A", "B"))
+save_plot("../fig_S1_combined.pdf", combined,
+          ncol = 2, 
+          nrow = 1, 
+          base_aspect_ratio = 1.1
+)

### 2015-07-02 FigS2/FigS2plotting.R
combined figure S2

@@ -19,16 +19,24 @@ levels(sigs$Sigat001) = c("BSCB1", "BSSS", "Both")
 blocks=geom_rect(data=sigs, aes(fill=Sigat001, xmin=phystart/1000000, xmax=phystop/1000000, ymin=0, ymax=0.5)) 
 BSSS_H <- geom_point(data=alldata, aes(x=physpos/1000000, y=HetSS), colour="blue", size=1)
 BSCB1_H <- geom_point(data=alldata, aes(x=physpos/1000000, y=HetNSS), colour="red2", size=1)
-png('../FigS2A.png')
-ggplot(data=alldata) + blocks + BSSS_H + BSCB1_H + facet_grid(LG~.) + ylab("Heterozygosity") + xlab("Position (Mb)") + theme_bw() +
-scale_fill_manual(values=c("pink1", "skyblue2", "green"), name="P < 0.001") + theme(panel.border=element_rect(size=0.1))  + theme(axis.text.y=element_text(size=8))
-dev.off()
+#png('../FigS2A.png')
+physplot = ggplot(data=alldata) + blocks + BSSS_H + BSCB1_H + facet_grid(LG~.) + ylab("Heterozygosity") + xlab("Position (Mb)") + theme_bw() +
+  scale_fill_manual(values=c("pink1", "skyblue2", "green"), name="P < 0.001") + theme(panel.border=element_rect(size=0.1))  + theme(axis.text.y=element_text(size=6)) +
+  ggtitle("Physical Space")
+#dev.off()
 
 blocks=geom_rect(data=sigs, aes(fill=Sigat001, xmin=genstart, xmax=genstop, ymin=0, ymax=0.5)) 
 BSSS_H <- geom_point(data=alldata, aes(x=genpos, y=HetSS), colour="blue", size=1)
 BSCB1_H <- geom_point(data=alldata, aes(x=genpos, y=HetNSS), colour="red2", size=1)
-png('../FigS2B.png')
-ggplot(data=alldata) + blocks + BSSS_H + BSCB1_H + facet_grid(LG~.) + ylab("Heterozygosity") + xlab("Position (cM)") + theme_bw() +
-  scale_fill_manual(values=c("pink1", "skyblue2", "green"), name="P < 0.001") + theme(panel.border=element_rect(size=0.1))  + theme(axis.text.y=element_text(size=8))
-dev.off()
+#png('../FigS2B.png')
[…]

### 2015-07-03 response.tex
working on response

@@
 
+\response{We thank the editor and reviewers for their time and helpful comments.
+We have addressed these and other comments specifically in the text below.
+Text in the manuscript modified in response to the reviewers has been highlighted blue.
+}
+
 On the statistical side, candidate selected regions are picked out and effectively 1-sided tests are reported for the rare allele that has increased in frequency.
@@
 
-\response{blah}
+\response{We feel that while an experiment-wide p-value would be important if we were performing a formal genome scan, our purpose here was simply to show that while our overall power to distinguish selection from drift may be low, there are nonetheless individual genomic regions that show patterns of fixation suggestive of selection.
+We report the \% of simulations for each region that showed lower heterozygosity in simulations, so a simple bonferroni or other calculation could be calculated by a reader concerned we are not sufficiently conservative.
+\jri{Do we want to do an FDR for the p-values in Figure 3?}}
 
@@
 
-In the introduction the authors describe the concept of heterotic groups and 
-state that these lines maximize performance and heterosis when crossed.  I would 
-caution that this should be clarified to heterosis for yield rather than for all 
-traits. Research by other groups (ie Flint-Garcia et al 2009) suggests that 
-heterosis can vary for different traits and that the definition of heterotic 
-groups is primarily based on yield heterosis, not heterosis for other traits.
+In the introduction the authors describe the concept of heterotic groups and state that these lines maximize performance and heterosis when crossed.
+I would caution that this should be clarified to heterosis for yield rather than for all traits.
+Research by other groups (ie Flint-Garcia et al 2009) suggests that heterosis can vary for different traits and that the definition of heterotic groups is primarily based on yield heterosis, not heterosis for other traits.
 
-\response{blah}
+\response{This is a good point and we have added a few words in the intro in the hopes of making this clearer.}
 
@@
 population structure in breeding programs from the Ross-Ibarra group.  I 
-hesitate to even make this suggestion (do to potential flaws) but it might be 
-appropriate to also include Jiao et al nat Genetics 2012 here as this study also 
-looked at a large dataset of SNPs in modern maize.
+hesitate to even make this suggestion (do to potential flaws) but it might be appropriate to also include Jiao et al nat Genetics 2012 here as this study also looked at a large dataset of SNPs in modern maize.
 
-\response{While \citet{jiao2012genome} have decided against including citation of the Jiao et al paper, mostly because we are unsure of the veracity of the results (see \citet{lorenz2013phylogenetic}) and concerned by the fact that a wholesale recoding of their genotyping data appeared to have little impact on their conclusions.}
+\response{We had avoided citing \citet{jiao2012genome} because we were  unsure of the veracity of the results (see \citet{lorenz2013phylogenetic}) and concerned by the fact that a wholesale recoding of their genotyping data appeared to have little impact on their conclusions.
+We have now added a citation, however, as it is certainly true that they performed similar analyses.}
 
-I was quite intrigued by the difference in distance between cycles 4 and then 8-
-16.  The authors suggest that this could be due to some differences in genetic 
-maintenance of the populations.  I am curious if there were any particular 
-environmental influences during that time period that may have led to stronger 
-selection or if there were any differences in agronomic practices that have 
-occurred during the long-term experiment (ie irrigation, fertilization, plant 
-density, etc).  it would be helpful to know if any of these factors may have 
-contributed.
+I was quite intrigued by the difference in distance between cycles 4 and then 8-16.
+The authors suggest that this could be due to some differences in genetic maintenance of the populations.
+I am curious if there were any particular environmental influences during that time period that may have led to stronger selection or if there were any differences in agronomic practices that have occurred during the long-term experiment (ie irrigation, fertilization, plant density, etc).
+it would be helpful to know if any of these factors may have contributed.
 
-\response{blah}
+\response{\jri{Jode? Justin? I don't have the answer here, but suspect it is ``no''}}
 
-It was a bit surprising to me that there was no differentiation among the 
-founder inbreds of the two populations.  If PCA was run only on the founders 
-(without any other lines) do you get separation of the two groups or not?
+It was a bit surprising to me that there was no differentiation among the founder inbreds of the two populations.
+If PCA was run only on the founders (without any other lines) do you get separation of the two groups or not?
 
-\response{blah}
+\response{\jri{Justin, have you done this? Easy to do?}}
 
-Figure 3 is quite useful and helps to show large physical regions that have lost 
-heterozygosity.  The authors discuss the physical-genetic difference.  It would 
-be useful to have a supplemental figure that shows these same values (cycle 16 
-heterozygosity on y-axis) versus genetic position on x-axis.  I am curious if 
-there are ever regions of multiple cM on chromosome arms that become fixed.  
-This would be a nice complementary figure.  These are provided for some 
-chromosomes in 4B and 5B but would be useful to see genome wide.
+Figure 3 is quite useful and helps to show large physical regions that have lost heterozygosity.
+The authors discuss the physical-genetic difference.
+It would be useful to have a supplemental figure that shows these same values (cycle 16 heterozygosity on y-axis) versus genetic position on x-axis.
+I am curious if there are ever regions of multiple cM on chromosome arms that become fixed.
+This would be a nice complementary figure.
+These are provided for some chromosomes in 4B and 5B but would be useful to see genome wide.
 
-\response{blah}
+\response{Heterozygosity plotted on a genetic map was shown in the original supplemental figure 2 (which the reviewer pointed out we failed to reference).
+We have switched this figure for the original figure 2.
+As shown in the new figure 3, there are a few regions on chromosome arms that have been fixed and span multiple cM.
+\jri{do we need to note this in the main text?
+is it an intesting observation, maybe implying rapid selection?}}
 
-The large regions of fixation are quite intriguing.  Is there any evidence for 
-major flowering time or disease resistance QTL in these regions?
+The large regions of fixation are quite intriguing.  Is there any evidence for major flowering time or disease resistance QTL in these regions?
 
-\response{blah}
+\response{\jri{Justin, any ideas? Do we do some formal correspondence with NAM results? Or say ``geez this is hard because QTL are big and we're not sure we'd believe the results?''}
 
-I am somewhat baffled as to why the Fig S1 was put in supplemental materials.  A 
-central aspect of this study is this issue of genetic drift relative to 
-selection and there are two pages of results text for this section.  However, 
-this useful figure is only in supplemental materials.  I would strongly 
-encourage moving this figure to the actual manuscript.  If space is an issue 
-then only use figu 4 or fig 5 as these are quite redundant with each other in 
-terms of conceptual views. 
-The legend to fig S1 could use additional detail.  I am guessing that the black 
-lines are 95\% CI but this is not clear.  Please explain the black lines.
+I am somewhat baffled as to why the Fig S1 was put in supplemental materials.
+A central aspect of this study is this issue of genetic drift relative to selection and there are two pages of results text for this section.
+However, this useful figure is only in supplemental materials.
+I would strongly encourage moving this figure to the actual manuscript.
+If space is an issue then only use figu 4 or fig 5 as these are quite redundant with each other in terms of conceptual views.
+The legend to fig S1 could use additional detail.  I am guessing that the black lines are 95\% CI but this is not clear.  Please explain the black lines.
 
-\response{blah}
+\response{Following the reviewer's suggestion we have moved this figure to the main text; the figure legend now explains the black lines.
+The old figures 4 and 5 have been combined into a single multi-panel plot.}
 
@@
 
-\response{blah}
+\response{Figure S2 and Figure 3 have been switched as described above; the supplmental figure is now referenced in the caption to the (new) Figure 3.}
 
-The last paragraph of the results section assesses the fixation of particular 
-haplotypes over time.  I think it could be very useful to visualize this data 
-for each of the large blocks.  This could be done by using a graph of allele 
-frequency over time for each region.  Right now these results are presented in 
-text form for only two regions but it would be very interesting to see these 
-results for all eight regions shown in table 1.  What is the fixation at each 
-cycle in the observed data and how often was this region fixed in simulations.  
+The last paragraph of the results section assesses the fixation of particular haplotypes over time.
+I think it could be very useful to visualize this data for each of the large blocks.
+This could be done by using a graph of allele frequency over time for each region.
+Right now these results are presented in text form for only two regions but it would be very interesting to see these results for all eight regions shown in table 1.
+What is the fixation at each cycle in the observed data and how often was this region fixed in simulations.
 I appreciate the authors point that these fixations do occur at low rates in
-simulations and since they are studying a genome-wide pattern a number of low 
-frequency events are expected to occur by chance.
+simulations and since they are studying a genome-wide pattern a number of low frequency events are expected to occur by chance.
 

### 2015-07-03 discussion.tex
working on response

@@
 
-Several key inbreds in the ‘stiff-stalk’ heterotic group, B73, B37 and B14, were derived from the BSSS population \citep{darrah19861985, 1986; troyer1999background}. 
+Several key inbreds in the ‘stiff-stalk’ heterotic group, B73, B37 and B14, were derived from the BSSS population \citep{darrah1986, troyer1999background}. 
 B37 and B14 were derived from cycle 0, and B73 was derived from a half-sib recurrent selection program also started with the BSSS population. 

### 2015-07-03 intro.tex
working on response

@@
 Although maize is naturally an outcrossing organism, modern breeding takes advantage of highly inbred lines that are used in controlled crosses to produce hybrids. 
-Hybrid maize, first developed in the early 20th century by \jri{East and Shull?}, was so successful that within a few decades it had completely replaced mass-selected open pollinated varieties in the United States \citep{crabb1947hybrid}.
+Hybrid maize, first developed in the early 20th century \citep{crow1998}, was so successful that within a few decades it had completely replaced mass-selected open pollinated varieties in the United States \citep{crabb1947hybrid}.
 The shift towards selection of inbred lines based on their ability to generate good hybrids – referred to as ‘combining ability’ – constituted an abrupt change from the open-pollinated mass selection that breeders practiced for millennia \citep{anderson1944sources, troyer1999background}.
-Maize inbred lines are now partitioned into separate ‘heterotic groups’ that maximize performance and hybrid vigor (heterosis) when inbreds from different heterotic groups are crossed with each other \citep{tracy2006historical}.
+Maize inbred lines are now partitioned into separate ‘heterotic groups’ that maximize performance and hybrid vigor (heterosis) \rev{for yield} when inbreds from different heterotic groups are crossed with each other \citep{tracy2006historical}.
 
@@
 Advances in high throughput genotyping and the development of a maize reference genome now enable the observation of maize population structure at high marker density across the whole genome \citep{ganal2011a-large,chia2012maize}.
-These studies have examined a broad spectrum of germplasm at various points in the history of maize to search for the signals of population structure and artificial selection \citep{Hufford2012b, van2012historical}.
+These studies have examined a broad spectrum of germplasm at various points in the history of maize to search for the signals of population structure and artificial selection \citep{Hufford2012b, van2012historical,jiao2012genome}.
 Although selective sweeps remaining from domestication are clearly visible, the impact of selection during modern breeding appears comparatively small in terms of its impact on genomic diversity \citep{Hufford2012b,van2012historical}, despite steady, heritable improvement in phenotype \citep{duvick2005contribution}.
@@
 This model of selection, named reciprocal recurrent selection, provides the generalized model for strategies commonly used in commercial maize breeding \citep{comstock1949breeding, duvick2004long}. 
-The Iowa RRS experiment proves especially relevant because lines derived from the BSSS population have had a major impact upon the development of commercial hybrids \citep{duvick2004long, darrah19861985}, the formation of modern heterotic groups \citep{troyer1999background; senior1998utility}, and the choice of a maize reference genome \citep{schnable2009the-b73-maize}.
+The Iowa RRS experiment proves especially relevant because lines derived from the BSSS population have had a major impact upon the development of commercial hybrids \citep{duvick2004long, darrah1986}, the formation of modern heterotic groups \citep{troyer1999background, senior1998utility}, and the choice of a maize reference genome \citep{schnable2009the-b73-maize}.
 \jri{something about heterosis}

### 2015-07-03 main.tex
working on response

@@
 %editing commands (please leave in for now)
-\newcommand{\jri}[1]{\textcolor{blue}{ \emph{\scriptsize  #1}} }
+\newcommand{\jri}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} }
 \definecolor{gerkegreen} {rgb} {0,0.6,0}
-\newcommand{\jpg}[1]{\textcolor{mattgreen}{ \em{\scriptsize  #1}} }
+\newcommand{\jpg}[1]{\textcolor{gerkegreen}{ \em{\scriptsize  #1}} }
 \newcommand{\citex}[1]{\textcolor{red}{\bf #1}}
+\newcommand{\rev}[1]{\textcolor{blue}{\bf #1}}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRO
-\input{intro}
+\section*{Introduction}
+Although maize is naturally an outcrossing organism, modern breeding takes advantage of highly inbred lines that are used in controlled crosses to produce hybrids. 
+Hybrid maize, first developed in the early 20th century \citep{crow1998}, was so successful that within a few decades it had completely replaced mass-selected open pollinated varieties in the United States \citep{crabb1947hybrid}.
+The shift towards selection of inbred lines based on their ability to generate good hybrids – referred to as ‘combining ability’ – constituted an abrupt change from the open-pollinated mass selection that breeders practiced for millennia \citep{anderson1944sources, troyer1999background}.
+Maize inbred lines are now partitioned into separate ‘heterotic groups’ that maximize performance and hybrid vigor (heterosis) \rev{for yield} when inbreds from different heterotic groups are crossed with each other \citep{tracy2006historical}.
+
+Multiple studies with molecular markers have indicated that these heterotic groups have diverged genetically over time to become highly structured and isolated populations, resulting in a dramatic restructuring of population genetic variation \citep{duvick2004long, ho2005extent, feng2006temporal}.
+Advances in high throughput genotyping and the development of a maize reference genome now enable the observation of maize population structure at high marker density across the whole genome \citep{ganal2011a-large,chia2012maize}.
+These studies have examined a broad spectrum of germplasm at various points in the history of maize to search for the signals of population structure and artificial selection \citep{Hufford2012b, van2012historical,jiao2012genome}.
+Although selective sweeps remaining from domestication are clearly visible, the impact of selection during modern breeding appears comparatively small in terms of its impact on genomic diversity \citep{Hufford2012b,van2012historical}, despite steady, heritable improvement in phenotype \citep{duvick2005contribution}.
+The lack of distinct selection signals from modern breeding may be due to population-specific selection; this interpretation is supported by the measured success in identifying targets of selection in individual experimental populations selected for a limited number of phenotypes \citep{hirsch2014insights, beissinger2014genome}.
+
+Here we take a detailed look at the dynamics of genome-wide patterns of diversity over time within an individual breeding program. 
+The Iowa Stiff Stalk Synthetic (BSSS) and the Iowa Corn Borer Synthetic No.
+1 (BSCB1) Reciprocal Recurrent Selection Program of the USDA-ARS at Ames, Iowa (hereafter referred to as the Iowa RRS) represents one of the best-documented public experiments on selection for combining ability and hybrid performance.
+BSSS and BSCB1 have been recurrently selected for improved cross-population hybrids \citep{penny1971twenty}. 
+\jri{add some text explaining recurrent selection}
+This model of selection, named reciprocal recurrent selection, provides the generalized model for strategies commonly used in commercial maize breeding \citep{comstock1949breeding, duvick2004long}. 
+The Iowa RRS experiment proves especially relevant because lines derived from the BSSS population have had a major impact upon the development of commercial hybrids \citep{duvick2004long, darrah1986}, the formation of modern heterotic groups \citep{troyer1999background, senior1998utility}, and the choice of a maize reference genome \citep{schnable2009the-b73-maize}.
+\jri{something about heterosis}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRO
 
-\input{methods}
+\section*{Methods}
+\subsection*{The BSSS and BSCB1 recurrent selection program}
+The Iowa RRS experiment began with founder inbred lines (Table S1) that were randomly mated to create the BSSS and BSCB1 ‘cycle 0’ base populations. 
+Each group was then recurrently selected for improved combining ability with the other group \citep{penny1971twenty,  keeratinijakal1993responses}. 
+Initially approximately 200-250 starting ‘S0’ plants within each population were simultaneously self-fertilized to generate ‘S1’ lines and crossed to a random sample (4-6) of plants from the other population to generate testcross seed.
+The material was evaluated for highly heritable traits including standability, disease resistance and ear rot resistance in the nursery at the time the testcross seed was made, and 100 testcrosses were then grown in replicated field trials and evaluated for grain yield, grain moisture, and standability.
+S1 seed of 10 selected families within each population were randomly mated to create a new ‘cycle n+1’ population. 
+After cycle 4, the testcrosses were carried out on the S1 lines themselves rather than at S0, which leads to another round of selfing prior to the creation of the next cycle.
+Ten lines (out of 100) were selected and advanced to form the next cycle between cycles 1 and 8, and twenty lines were selected between cycles 9 and 16 \citep{keeratinijakal1993responses}.
+Founder inbreds and samples from cycles 0, 4, 8, 12, and 16 were genotyped at 39,258 SNPs that passed a set of quality filters and could be assigned collinear genetic and physical map positions.
 
-\input{results}
+\subsection*{Plants and inbred lines used}
+The plants and inbred lines used in this experiment are listed in Table S1. 
+We genotyped 36 plants from each of the BSSS and BSCB1 populations at selection cycles 0, 4, 8, 12, and 16. 
+These plants represent descendants of the original populations which have been randomly mated to maintain seed. 
+We also genotyped the founder inbreds for each population, with the exception of lines F1B1, CI.617, WD456, and K230 for which seed was not available. 
+The data for founder line CI.540 was not used because the genotyped material was heterozygous. 
+A number of derived lines were also genotyped for calibrating phasing and imputation procedures (see below).
 
-\input{discussion}
+\subsection*{Genotype data}
+Plants from the cycles of selection, founders, and derived lines were grown in a greenhouse and tissue was collected at the 3 leaf stage. 
+Tissue was lyophilized, ground, and DNA extracted by a CTAB procedure \citep{saghai-maroof1984ribosomal}. 
+Samples were genotyped using the 24-sample Illumina MaizeSNP50 array \citep{ganal2011a-large} according to the Illumina Infinium protocol, and imaged on an Illumina BeadStation at the University of Missouri DNA core facility.
+Genotypes were determined with the GenomeStudio v2010.2 software using the manufacturer's MaizeSNP50\_B.egt cluster file. 
+\jri{this is an edit because url listed in pages doc no longer valid. correct that this is not a custom .egt file?}
+The design of the maize SNP50 Chip included a relatively small ascertainment panel of inbred lines, introducing a bias in the frequencies of SNPs included on the chip \citep{ganal2011a-large}. 
+However, because our simulations are based not on theoretical expectations but instead on sampling from the observed data at cycle 0, we expect ascertainment bias to have a minimal impact on our results.
+The effect of this ascertainment bias was shown to be minimal in another study of North American germplasm \citep{van2012historical}.
+
+We called 48,919 SNPs on the Illumina platform from the MaizeSNP50\_B.egt cluster file. Genotypes with quality scores of 50 or less were recoded as missing data. 
+Three plants were removed from the data due to an excess of missing data (the derived line B10, a cycle 0 plant from BSSS, and a cycle 8 plant from BSSS). 
+In addition, BSCB1 plant 31 from cycle 4 appeared switched with plant 31 from cycle 8 based on our principal component analysis (PCA), so we switched the labels for these two genotypes to correct the mistake.
+To avoid structure among the missing data, we removed any SNP that was coded as missing in more than 3 plants in either group of founders or any group of plants from a particular cycle and population.
+Preliminary analysis by PCA and heatmap plots of distance matrices revealed two additional likely mix-ups. 
+Plant 23 from BSSS cycle 8 was a clear outlier from the BSSS population as a whole and plant 2 from BSSS cycle 0 is likely a mislabeled plant from cycle 16.
+Since there was no evidence suggesting when mis-labelings occurred, each of these plants was removed from the analysis.
+
+\subsection*{Integrating the genetic and physical map}
+The interpretation of our results depends upon a genetic and physical map that is as accurate as possible. 
+We therefore took steps to improve the positions of the SNP markers on the genetic and physical maps relative to version 5A.59 of the maize genome assembly. 
+The probable physical position of each SNP based was obtained by comparing SNP context sequences to the genome sequence. 
+For this purpose, SNP context sequences were defined as the sequence 25 bp upstream of the SNP, the bp representing the SNP itself, followed by 25 bp downstream of the SNP, making a total sequence length of 51bp.
+When a single genomic location was queried by two separate probes on the array, we chose the probe with higher quality calls and dropped the other marker from the dataset. 
+To assign a genetic position for each SNP, we used a map derived from the B73xMo17 (IBM) mapping population similar to the IBM framework map in \citet{ganal2011a-large}. 
+This genetic map contains 4,217 framework SNP markers, which provides a much higher density than the map used to order the 5A.59 release of the maize genome sequence. 
+As a result, we identified several places in the genome where the physical positions were incorrect according to our genetic map. 
+These cases included both simple reversals of the physical map relative to the genetic map, and also the assignment of blocks of markers to the wrong linkage group, which we refer to as mis-mapped blocks.
+SNPs at these loci were reordered to match our genetic map. To maintain collinearity between the genetic and physical map, the physical positions of these SNPs were reassigned as follows. 
+Individual mis-mapped markers small reversals and mis-mapped blocks ($<10$ kb) were removed from the data. 
+Small rearrangements of this sort are more likely to represent mis-mapped paralagous sequence than true errors in the physical map. 
+When larger reversals were identified, we transposed the physical positions of the SNPs from one end of the segment to the other. 
+Mis-mapped blocks were often larger than the physical gap into which they were moved. 
+We therefore assigned the first SNP of the block to a position 10 kb downstream from the previous SNP on the correct linkage group. 
+We then recalculated genomic coordinates for the rest of the chromosome based on the marker distances within the translocated segment. 
+The last SNP of the block was also given a 10 kb cushion between itself and the next SNP on the correct linkage group.
+
+Only a portion of the SNPs on the array had genetically mapped positions. 
+‘Unmapped’ SNPs (those with a physical position but no genetic position) bounded by two mapped markers were moved along with their mapped neighbors if both anchoring mapped markers were also moved. 
+However, it is unclear whether unmapped SNPs just outside of these anchors should be kept in place or moved along with the adjacent SNPs. 
+Since most inversions were small relative to the genetic map (and would therefore still fall in the same window of a sliding window analysis), these SNPs were left in place. 
+However, markers bordering translocations were removed to ensure there were no markers mapped to the incorrect linkage group. 
+Unmapped SNPs were then assigned a genetic position by linear interpolation of genetic vs. physical distance using the approx() function in R \citep{rteam}. 
+The IBM genetic map distances were then converted to single-meiosis map distances using the formulae of \citet{winkler2003determination}. 
+Finally, SNPs located at physical positions outside of those bounded by the genetic map (such as the telomeres) were assigned the genetic position of their nearest mapped neighbor. 
+Since moved segments were arbitrarily joined 10 kb from their nearest genetic neighbor, we acknowledge that the physical positions of these markers are only estimates. 
+However, the estimated junctions are small relative to the genetic windows used for our analysis. 
+The final map used is provided as supplemental material (Supplementary File S1).
+
+\subsection*{Haplotype phasing}
+Although the genotypes of the plants from each population are unphased, the homozygous genotypes of the founders and derived inbreds provide excellent prior information for a probabilistic estimation of genotype phase in the populations.
+We therefore used fastPHASE \citep{scheet2006fast} to estimate the genotype phase of each plant.
+To estimate the error in phasing, we created test cases by combining the genotypes of two derived inbreds into a hypothetical “F1 hybrid” of unknown phase.
+This F1 was presented to fastPHASE with the rest of the data, except that its parent inbreds were removed. 
+Analyses of several hypothetical F1’s from different cycles of selection revealed very low phasing error rates (Table S3). 
+Therefore the phased genotypes of cycle 0 plants were used as the starting data for simulations (see below).
+
+\subsection*{Diversity and principal component analysis}
+Heterozygosity (H) was measured as:
+
+\begin{align*}
+H = 2p(1-p)
+\end{align*}
+
+where $p$ and $(1-p)$ are the frequencies of the two SNP alleles. 
+$F_{ST}$ \citep{hudson1992statistical} was calculated using the HBKpermute program in the analysis package (\url{https://github.com/molpopgen/analysis}) of the software library libsequence \citep{thornton2003libsequence}.
+All results were plotted using the R package ggplot2 \citep{wickham2009ggplot2}. 
+We conducted PCA by singular value decomposition, as described in \cite{mcvean2009genealogical}.
+
+\subsection*{Simulations}  
+Our simulation sought to model the effects of genetic drift in the Iowa RRS experiment independent of any selection, and our model thus closely followed the published methods of the Iowa RRS \citep{penny1971twenty,  keeratinijakal1993responses}.
+Starting individuals in each population were constructed by randomly sampling two distinct haplotypes with replacement from the phased haplotypes of cycle 0. 
+In the actual random mating scheme used in the Iowa RRS experiment, a single pairing could only contribute four gametes to the next generation (two kernels each from two ears), and our simulation reflects this.
+Advanced cycles were simulated by randomly mating gametes from self-fertilized plants of the previous cycle until 10 new individuals were created. 
+The first cycle involved two rounds of random mating, whereas all subsequent cycles used one round. 
+After cycle 5, the process employed two rounds of selfing instead of one. 
+After cycle 7, the population size was increased from 10 to 20. 
+At cycles 4, 8, 12, and 16, the plants were randomly mated to match the sample size of the observed data. 
+The genotypes of these simulated random matings are the final results of each simulation and were analyzed in the same way as the observed data.
+
+All recombination events in the simulations were carried out in R with the hypred software package (\url{cran.r-project.org/web/packages/hypred/}). 
+Simulations were executed in parallel on a computing cluster, with unique random number seeds drawn for each simulation. 
+Statistics were calculated for each simulation using the same formula as the experimental data. 
+We used non-overlapping sliding windows of equal genetic distance to account for the non-independence of markers in low-recombination regions when calculating measures of significance. 
+For the haplotype-based, single-locus simulations, recombination was simply replaced with binomial sampling of two alleles.
+
+\section*{Results}
+
+\subsection*{Population structure and genetic diversity}
+Founder inbreds and samples from cycles 0, 4, 8, 12, and 16 were genotyped at 39,258 SNPs that passed a set of quality filters and could be assigned collinear genetic and physical map positions (See Materials and Methods for details).
+Change in population structure throughout the Iowa RRS experiment can be observed visually by a principal component analysis (PCA). 
+A joint analysis of individuals from all the selection cycles (Figure \ref{fig:pca}) clearly separates the BSSS and BSCB1 populations along the first axis of variation, with increasing separation as the experiment progressed.
+The second axis of variation primarily separates the cycles from one another within each population. 
+There is no separation between the founders of the two populations, but at cycle 0, the BSSS population shows more divergence from the founders than does BSCB1. 
+This could be due to drift during either the population’s construction or subsequent maintenance. 
+Structure continued to develop within each population over the course of the experiment. 
+There is an especially wide gap between cycles 4 and 8, which correlates with the addition of an extra generation of self-pollination prior to selection at each cycle. 
+The distance between cycles then decreases dramatically after cycle 8, and corresponding to the increased effective population size (See Materials and Methods).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
+\begin{figure}[tb]   
+  \begin{center}
+   \vspace{-0mm}
+   \includegraphics[width=0.5\textwidth]{fig1}
+   \renewcommand{\baselinestretch}{0.9}
+   \vspace{-3mm}
+   \caption{Principal component analysis of the SNP data from Iowa RRS.
+The axes represent the first two eigenvectors from an analysis of cycles 0-16, with projection of the founder lines onto the vector space.
+The variation explained by each eigenvector is given in parentheses on the axes.
+The populations steadily diverge at increasing cycles, with less distinction visible between the founder groups.
+The comparatively large distance between cycles 4 and 8 corresponds to a switch from one to two generations of selfing at each cycle.
+The smaller separation between cycles 8-16 corresponds to an increase in effective population size from ten to twenty.
+The BSSS cycle 0 population has drifted away from the BSSS founders, despite the absence of intentional selection during the creation and maintenance of cycle 0.
+} 
+\vspace{-6mm}
+    \label{fig:pca}
+  \end{center}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
+
+No new genetic material was intentionally introduced into either population after the experiment's inception, so the substantial increase in genetic distance could only arise from the loss of genetic diversity within each population.
+Consistent with previous studies of the Iowa RRS \citep{messmer1991genetic, labate1997molecular,  hagdorn2003molecular, hinze2005population}, genome-wide genetic diversity (expected heterozygosity, H) decreases steadily across cycles of selection in both populations (Figure \ref{fig:decline}).
+\jri{do all these citations make sense here?
+I just copied these from the abstract}
+The loss of heterozygosity is smaller when the two populations are considered together, indicating the loss of different alleles within BSCB1 and BSSS. 
+This genetic differentiation is reflected by the tenfold increase in $F_{ST}$ between the founder lines and the populations at cycle 16.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
+\begin{figure}[tb]   
+  \begin{center}
+   \vspace{-0mm}
+   \includegraphics[width=0.5\textwidth]{fig2}
+   \renewcommand{\baselinestretch}{0.9}
+   \vspace{-3mm}
+   \caption{Heterozygosity and $F_{ST}$ across cycles. $H$ (left panel) and $F_{ST}$ (right panel) plotted as a function of selection cycle in each population.
+} 
+\vspace{-6mm}
+    \label{fig:decline}
+  \end{center}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
+
+We noticed an irregular increase in the number of polymorphic markers between BSSS cycles 4 and 8 (Table \ref{tab:s2}). 
+All of these newly polymorphic markers were present at extremely low frequency and were spread among various individuals. 
+This may represent a series of minor alleles that were not captured in our sample of cycle 4 individuals and thus appeared to ‘resurface’ at cycle 8. 
+Alternatively, the pattern may be the result of minor contamination at some point in the population’s history. 
+It was observed that an allele of the sugary gene associated with sweet corn appeared in the population at this time (O.S. Smith, personal communication), suggesting contamination may be the cause. 
+However, the low frequency of the ‘new’ alleles (typically only one or two alleles out of 72 possible in 36 diploid samples) means their effect on population diversity is minimal.  
+We did not attempt to incorporate this contamination into our simulation approaches, as it only makes our tests for low heterozygosity slightly more conservative.
+
+\subsection*{Fixation of large genomic regions}
+Figure \ref{fig:heterotic} shows heterozygosity varying along the genome at cycle 16 of each population. 
+Of particular note are extremely large pericentromeric regions of zero or near-zero heterozygosity spanning tens of megabases. 
+These regions experience low rates of meiotic recombination, which creates an expanded physical map relative to their genetic length \citep{ganal2011a-large}. 
+In general, the majority of fixed haplotype segments are small ($<2$ cM) in genetic space regardless of their physical size; one exception is 7 cM region on chromosome 1 in the BSCB1 population.
+
+\begin{figure*}[tb]   
+  \begin{center}
+\includegraphics[width=\linewidth]{figs2_jri}
+   \caption{Heterozygosity ($H$) at cycle 16 across all ten chromosomes in each population, calculated on 15-marker sliding windows with 5 marker steps.
+Heterozygosity values in BSSS (blue dots) and BSCB1 (red dots) are superimposed in one panel.
+2 cM windows of heterozygosity lower than 10 of 10,000 simulations ($P<0.001$) are shaded in light blue (BSSS) or pink (BSCB1).
+Two regions genome-wide show significantly low heterozygosity in both populations and are shaded green.
+(A) Physical map.
+(B) Genetic map.
+\rev{The same data plotted are plotted separately for each population in Figure \ref{fig:s2}.}}
+    \label{fig:heterotic}
+  \end{center}
+\end{figure*}
+
+The sheer physical size of the pericentromeric regions yields extremely high marker density on the genetic map, allowing for clear resolution of haplotype phasing and recombination breakpoints. 
+To further examine the fixation in these regions, we computationally imputed haplotype phase in the BSSS and BSCB1 populations and used the phased data to track haplotype frequencies and founder of origin.
+In most cases, these fixed haplotypes can be traced back to single founder inbreds. 
+For example, in BSSS, a 60 Mb (1.2 cM) region of chromosome 9 became fixed by cycle 12 and traces back to the founder Os420, and in BSCB1, a 60 Mb (0.7 cM) region became fixed on chromosome 4 and traces back to A340 (Figure \ref{fig:genphys}).
+Table \ref{tab:fix} gives a summary of the large genomic regions that have become fixed or nearly fixed by cycle 16. 
+These regions represent blocks of linked loci that show no evidence of recombination since at least the development of the founding inbred lines in the 1920’s and 1930’s.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
+\begin{figure*}[tb]   
+  \begin{center}
+  % \vspace{-0mm}
+   \includegraphics[width=\linewidth]{fig4_combined}
+ %  \renewcommand{\baselinestretch}{0.9}
+ %  \vspace{-3mm}
+   \caption{Heterozygosity ($H$) in each cycle across chromosome 9 of the BSSS (left) and chromosome 4 of the BSCB1 (right) plotted on the physical (top) and genetic (bottom) map.
+$H$ is calculated on 15-marker sliding windows, with 5 marker steps between each calculation.
+Each data point is color-coded based on a linear transformation of recombination rate (red = low recombination rate).
+Shaded regions represent 2 cM windows found to have heterozygosity values significantly lower than expected by simulation at a given cycle.
+Light shading denotes significance at $P<0.0025$, whereas the darker shading represents significance at $P<0.001$.
+Comparing the physical and genetic maps reveals that the regions of fixation are extremely small genetically, but some contain much of the physical content of the chromosome.
+} 
+\vspace{-6mm}
+    \label{fig:genphys}
+  \end{center}
+\end{figure*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
+
+\begin{table*}[t]
+\caption{Ancestry of haplotypes fixed in the cycle 16 population.}
+\fontsize{10}{12}\sf
+\centering
+\label{tab:fix}
+\begin{tabular}{llllll}
+Population &	 Chr.& Interval (cM) & Interval (Mb) &	Founder & Derived Lines \\
+BSSS       & 3          & 40.2-41.4            & 67.7-123              & CI187-2        & B94                               \\
+BSSS       & 3          & 43-48.5              & 129.2-157.1           & NDa            & B89, B94                          \\
+BSSS       & 4          & 39.7-42.1            & 39.9-82.7             & CI187-2        & B89, B94, B67, B72, B39, B43      \\
+BSSS       & 9          & 30.6-33.5            & 20.8-26.6             & Oh7b           & B89, B94, B43, B17, B72, B84, B67 \\
+BSSS       & 9          & 34.3-35.6c           & 30.8-90.4c            & Os420          & B89, B94                          \\
+BSCB1      & 2          & 50.2-51.8            & 80.6-114.5            & CC5            & B90, B91, B95, B97, B99           \\
+BSCB1      & 4          & 42.1-42.8            & 82.7-140              & NDa            & B90, B95, B97                     \\
+BSCB1      & 8          & 46.1-50.7            & 125.1-145.6           & P8             & B90, B97, B91, B99, B54           \\\hline 
+\end{tabular}
+\end{table*}
+%a ND = Not Determined (Either a recombinant haplotype or originates from an un-genotyped founder)
+%
+%b Although Oh7 is a BSCB1 founder, it is a descendant of CI.540, an un-genotyped BSSS founder. BSSS segments matching Oh7 presumably derive from CI.540
+%
+%c  Founders Ind_B2 (BSSS), CI187-2 (BSSS), R4 (BSCB1), and I205 (BSCB1) are all IBD at this region of chromosome 9
+
+\subsection*{The role of genetic drift}
+There is clear evidence of phenotypic improvement in response to selection in the Iowa RRS populations (Smith, 1983; Keeratinijakal and Lamkey, 1993; Schnicker and Lamkey, 1993; Holthaus and Lamkey, 1995; Brekke et al., 2011a; Brekke et al., 2011b; Edwards, 2011) and large changes in genetic structure indicated by molecular markers.
+A central issue for these maize populations and others like them is whether the changes observed at the molecular level are caused directly by selection on phenotype, or indirectly due to the genetic drift that selection imposes through inbreeding and small effective population sizes.
+To gauge the roles of selection and drift, we conducted simulations of the crossing and selection schemes used in the RRS experiment.
+Selection was executed at random in each simulation, so the patterns observed across simulations represent the expected distribution of effects caused only by recombination and genetic drift. 
+We conducted 10,000 simulations, with genetic recombination modeled on a framework map derived from the IBM population \citep{lee2002expanding}, a cross between inbred lines B73 (a BSSS derived line used to construct the reference genome) and Mo17 (an inbred with some co-ancestry to both the BSSS and BSCB1 founder germplasm).
+We sampled individuals from each cycle equal in number to the observed samples, measured heterozygosity, and compared the results to the observed data.
+
+\begin{figure*}[tb]   
+   \includegraphics[width=0.7\linewidth]{Fig_S1_combined.pdf}
+   \caption{Heterozygosity in each population, observed vs.
+simulated data.
+Heterozygosity was calculated as the average across all markers genome-wide  in the BSSS (A)  and BSCB1 (B) populations.
+The observed data is marked by the red line, the simulations by gray dots, and the median of the simulations by a green dot.
+\rev{Black lines represent the 95\% quantiles of the simulated data.} \jri{is that interpretation of the black lines correct?}}
+    \label{fig:compare_to_sims}
+\end{figure*}
+
+Averaged across the genome, the vast majority of the reduction in diversity observed in both populations can be attributed to genetic drift. 
+Nonetheless, we do observe deviations from the simulated values (Figure \ref{fig:compare_to_sims}). 
+The observed data show higher than expected heterozygosity at cycle 4, suggesting selection for heterozygosity in early cycles, a deviation from simulated and actual breeding practices, or an under-sampling of the diversity present in the original cycle 0 populations.
+\jri{where is this excess of diversity?
+if it's all pericentromeric can we argue for the first hypothesis?}
+As the cycles progress, heterozygosity falls more rapidly than expected in both populations, and the observed values at cycle 16 are significantly lower than the simulated data.  
+Simulations across a number of different marker densities were consistent with this result (data not shown).
+
+To examine the behavior of specific regions, we also calculated the observed and simulated results for each 2 cM segment of the genome. 
+The dynamics observed across most of the genome are largely insensitive to changes in window size (we tested from 2-4 cM, data not shown), and are consistent with strong genetic drift imposed by selection practices.
+\jri{did we test smaller regions?
+not enough markers?}
+A subset of loci were flagged as significant, and these loci almost always overlapped regions of fixation or near-zero heterozygosity in one population. 
+However, in these regions the values obtained by simulation are often quite low, such that the extreme simulated values sometimes lie only a few percentage points away from the observed value (Supplementary File S2), indicating that drift alone can explain most of the drop in diversity.
+
+Since the population size of the Iowa RRS is small (10-20), many biallelic SNPs should fix by chance regardless of their starting minor allele frequencies.
+The deviation from simulation arises not from changes in allele frequencies per se, but rather from the fixation of linked markers across larger than expected genetic distances.
+The validity of significance cutoffs therefore depend on the accuracy of our genetic map, and the maize genetic map is known to vary among genetic backgrounds across short genetic distances ($<5$cM) (McMullen et al., 2009).
+\jri{need to cite either \citep{bauer2013intraspecific} who shows there's lots of variation or maybe also \citep{rodgers2015recombination} who argues it's stable (but stable vs.
+variable is somewhat subjective}
+Deviations of the observed and simulated results could be due to selection, variation or inaccuracy in the genetic map, or a combination of these factors.
+
+To explore the roles of selection and drift independent of the genetic map, we returned to the large regions of fixation in the centromeres which showed no recombination across the full RRS experiment.
+Given the lack of recombination, each of these regions can be analyzed by the simulation of a single locus, and the high density of markers allows the clear resolution of the individual haplotypes. 
+We used the computationally phased data to measure the frequency of the fixed haplotype at each cycle, and assessed the probability of observing the fixation event given the initial frequency. 
+The BSSS chromosome 9 haplotype fixed at cycle 16 was at low frequency at cycle 0 (7 out of 68 haplotypes), but increased rapidly in frequency by cycle 8 (66 out of 70 haplotypes). 
+Simulation of the haplotype as a single locus in the RRS experiment produces this increase in frequency in only 3.9\% of 1000 independent simulations, whereas the haplotype was lost in over 80\% of the simulations.
+In BSCB1, a 30 Mb, 1.6 cM region of chromosome 2 became nearly fixed by cycle 8 (67/70) despite a prevalence of 4/72 at cycle 0, which occurred 1.5\% of the time by simulation. 
+Although these results suggest that selection may have pushed these haplotypes to fixation, the fact that fixation of such a rare haplotype still occurred in some simulations speaks to the strong genetic drift imposed upon the BSSS and BSCB1 populations.
+\jri{do you think we need to do more to explain here or above?
+in the previous paragraph we argue biallelic loci should fix by chance regardless of starting allele freq.}
+Interestingly, each of these two genomic regions harbored a different cycle 0 haplotype at higher frequency, but these higher-frequency haplotypes were subsequently lost within the RRS population. 
+In other cases, the haplotypes which eventually fixed were at moderate frequency in the cycle 0 populations and drift to fixation in the majority of simulations.
+
+\section*{Discussion}
+
+Our analysis of diversity in the Iowa RRS experiment reveals a steady loss of diversity in the BSSS and BSCB1 populations as they became increasingly differentiated from one another over time.
+Principal component analysis shows that as the effective population size and the rates of inbreeding were altered, the rates of change in population structure were altered as well. 
+These patterns of population structure, diversity, and differentiation between BSSS and BSCB1 can be largely reproduced by simulation without any selection, supporting the hypothesis that the majority of the genetic structure observed can be attributed to genetic drift alone, despite effective selection for phenotypic improvement.
+This drift was driven by inbreeding through self-pollination and the small effective population sizes used to select a limited number of high-performing, potentially related individuals at each cycle.
+
+Reciprocal recurrent selection serves as a model for the method of hybrid maize improvement employed broadly in North America \citep{duvick2004long}.
+The trends observed in the Iowa RRS are also a common theme in similar recurrent selection experiments \citep{romay2012effect,lamkey2014relative}.
+Increased population structure due to increased inbreeding and lowered effective population size are visible in the modern ‘stiff-stalk’ and ‘non stiff-stalk’ heterotic groups as well \citep{van2012historical}.
+Given the lack of strong selection signals in each case, genetic drift has most likely played a large role in the current genetic structure of modern maize.
+
+Several key inbreds in the ‘stiff-stalk’ heterotic group, B73, B37 and B14, were derived from the BSSS population \citep{darrah1986, troyer1999background}. 
+B37 and B14 were derived from cycle 0, and B73 was derived from a half-sib recurrent selection program also started with the BSSS population. 
+We examined these three inbreds at the pericentromeric regions listed in Table \ref{tabfix} \jri{correct?}, and found that in most cases they carry different haplotypes from those that rose to high frequency in the RRS experiment.
+Near isogenic lines that substitute these centromeric haplotypes from key inbreds could be used to determine if the haplotypes at high frequency in cycle 16 of BSSS provide a phenotypic advantage when crossed with BSCB1.
+
+Both our study and previous work \citep{labate1997molecular} identified genetic differentiation between cycle 0 of BSSS and the founder inbreds. 
+The loss of diversity in the cycle 0 population likely led to the loss of some rare alleles and haplotypes. 
+The BSSS was maintained for several years before RRS was initiated, and the differentiation could have occurred due to low effective population size during maintenance. 
+If so, then this drift will have impacted the trajectory of the RRS program and of the key inbreds derived from BSSS. 
+However, cycle 0 has also undergone maintenance since the beginning of the RRS experiment, so some genetic drift may have occurred during this time as well.
+
+Although drift can explain most of the genetic structure genome-wide, phenotypic data provide clear evidence that selection has altered the frequencies of favorable alleles in the BSSS and BSCB1 populations.
+Numerous experiments have shown that the selected populations and the hybrids formed from them are superior to their cycle 0 ancestors \citep{smith1983evaluation, keeratinijakal1993responses, schnicker1993interpopulation, holthaus1995population}.
+Improvement occurred not only for hybrid yield, but concurrently for plant architecture and tolerance to high-density planting in both hybrids and inbreds \citep{brekke2011selection, brekke2011selectionb, edwards2011changes, lauer2012morphological}.
+Since these same trends are observed across wider North American germplasm \citep{tollenaar1989genetic}, identification and characterization of the loci conferring phenotypic improvement in BSSS and BSCB1 could reveal some of the genetic mechanisms by which maize yield has steadily improved over the decades.
+We find that the genomic regions of extremely low diversity evident at cycle 16 are unlikely to be produced by simulation, and heterozygosity falls more than expected across the genome as a whole.  
+However, simulated and observed values are often close, so overall the signature of selection is difficult to detect at any given locus. 
+This is due in large part to a limitation of statistical power imposed by inbreeding and low effective population sizes.
+\citet{lorenz2015selection} analyze a similar selection experiment with comparable effective population size and also find little little statistical evidence for selection on individual loci in spite of genetic gain for the trait of interest.
+Nonetheless, we show that an identity-by-descent, haplotype based approach provides additional power, as it can distinguish between the fixation of rare and common haplotypes. 
+However, the results of single locus simulations are sensitive to sampling error and drift caused by population maintenance. 
+In this case, it is difficult to assess significance with only a single population to test.
+This type of analysis is more effective across several replicated populations, which can control for genetic drift due to the independence between the selections in each replicate \citep{lamkey2014relative}.
+
+\citet{van2012historical} assessed variation in and tested for selection across a wider range of North American maize germplasm using the same SNP array. 
+Although selected loci were identified in that study, the overall effects on genomic diversity and haplotype patterns were relatively small compared with the impact of domestication. 
+That study was a wide survey across many germplasm sources. 
+The limited overall impact of selection could arise because different haplotypes and loci are selected in different breeding programs. 
+In support of this idea, we find that the most likely targets of selection between the BSSS and BSCB1 populations occur at non-overlapping loci. 
+This non-overlap between putatively selected loci in complementary populations has also been observed in commercial breeding programs \citep{feng2006temporal}.
+
+The observation that the same targets of selection are not observed in the opposing heterotic populations bears implications for the genetic mechanisms responsible for heterosis and the success of maize hybrids.
+Classic overdominance models of heterosis predict that at a single locus, two distinct alleles confer heterozygote advantage when combined, while the dominance model predicts that heterosis is driven by dominance effects and the complementation of linked alleles in low-recombination regions (dominance or pseudo-overdominance).
+In the case of true overdominance, we expect selection should lead to decreased heterozygosity at a locus in both populations as complementary haplotypes are fixed in each group. 
+We find no evidence of this genetic phenomenon. 
+The observed pattern instead favors a dominance model, where fixation of a haplotype in one population simply selects against that same haplotype in the other population. 
+Because most deleterious alleles are rare in both heterotic groups \citep{Mezmouk2014}, most haplotypes in the second population will have a different suite of deleterious variants and will complement the fixed haplotype reasonably well, such that selection will have little impact on diversity in the second population.
+Our data are consistent with such a process having been important for hybrid improvement in the Iowa RSS experiment, and the lack of extreme differentiation seen over time across US Corn Belt germplasm is consistent with a role for such selection in maize hybrid improvement in general \citep{van2012historical}.
+
+Our results are consistent with the dominance and pseudo-overdominance models, but we caution that the exact outcome of particular models will depend strongly on the effects of selection, drift, and the frequencies of beneficial alleles.
+Disentangling these factors remains a major challenge that will require careful simulation of maize breeding using a wider range of parameters. 
+This is especially true because in a model of hybrid complementation, genetic drift in one population can alter the selective value of alleles in the other population. 
+When drift and selection interact in such a manner, tests for selection that attempt to independently partition the effects of each force may not provide the full picture.
+Furthermore, uncertainties in the fine-scale relationship between genetic and physical distance make it difficult to assign significance in forward-simulation approaches.
+In the end, the best test for selection of specific genomic regions will ultimately be conducted by phenotypic observation in the field in balanced tests of different haplotypes. 
+In the case of individual breeding programs, genome-wide genotyping such as we conducted here can identify the lines carrying the recombinant haplotypes and introgressions necessary to conduct such an experiment.
 
 \begin{acknowledgments}
-We thank O.
-S.
-Smith and members of the Ross-Ibarra lab for comments on earlier versions of the manuscript.
+We thank Oscar ``Howie'' Smith and members of the Ross-Ibarra lab for comments on earlier versions of the manuscript.
 J.P.G received support for this research as a Merck Fellow of the Life Sciences Research Foundation.

### 2015-07-03 methods.tex
supp figures, tables
working on response

@@
 Genotypes were determined with the GenomeStudio v2010.2 software using the manufacturer's MaizeSNP50\_B.egt cluster file. 
-\jri{this is an edit because url listed in pages doc no longer valid.
-correct that this is not a custom .egt file?}
+\jri{this is an edit because url listed in pages doc no longer valid. correct that this is not a custom .egt file?}
 The design of the maize SNP50 Chip included a relatively small ascertainment panel of inbred lines, introducing a bias in the frequencies of SNPs included on the chip \citep{ganal2011a-large}. 
@@
 
-We called 48,919 SNPs on the Illumina platform from the MaizeSNP50_B.egt cluster file. Genotypes with quality scores of 50 or less were recoded as missing data. 
+We called 48,919 SNPs on the Illumina platform from the MaizeSNP50\_B.egt cluster file. Genotypes with quality scores of 50 or less were recoded as missing data. 
 Three plants were removed from the data due to an excess of missing data (the derived line B10, a cycle 0 plant from BSSS, and a cycle 8 plant from BSSS). 
@@
 Finally, SNPs located at physical positions outside of those bounded by the genetic map (such as the telomeres) were assigned the genetic position of their nearest mapped neighbor. 
-\jri{so everything beyond 0 gets a 0 position on the map instead of negative?}
 Since moved segments were arbitrarily joined 10 kb from their nearest genetic neighbor, we acknowledge that the physical positions of these markers are only estimates. 

### 2015-07-03 results.tex
working on response

@@
 The deviation from simulation arises not from changes in allele frequencies per se, but rather from the fixation of linked markers across larger than expected genetic distances.
-The validity of significance cutoffs therefore depend on the accuracy of our genetic map, and the maize genetic map is known to vary among genetic backgrounds across short genetic distances (< 5cM) (McMullen et al., 2009).
+The validity of significance cutoffs therefore depend on the accuracy of our genetic map, and the maize genetic map is known to vary among genetic backgrounds across short genetic distances ($<5$cM) (McMullen et al., 2009).
 \jri{need to cite either \citep{bauer2013intraspecific} who shows there's lots of variation or maybe also \citep{rodgers2015recombination} who argues it's stable (but stable vs.

### 2015-07-03 sup1.tex
working on response

@@
 
-\begin{figure}[tb]   
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
+\begin{figure*}   
   \begin{center}
-   \includegraphics[width=\linewidth]{figs1_jri}
-   \caption{Heterozygosity in each population, observed vs.
-simulated data.
-Heterozygosity was calculated as the average across all markers genome-wide  in the BSSS (A)  and BSCB1 (B) populations.
-The observed data is marked by the red line, the simulations by gray dots, and the median of the simulations by a green dot.
+   \vspace{-0mm}
+   \includegraphics[\linewidth]{fig3}
+   %\renewcommand{\baselinestretch}{0.9}
+   \vspace{-3mm}
+   \caption{ Heterozygosity ($H$) at cycle16 across all ten chromosomes in each population.
+$H$ is calculated on 15-marker sliding windows with 5 marker steps.
+Each point is plotted at the midpoint of the 15-marker window.
 } 
-    \label{fig:s1}
-  \end{center}
-\end{figure}
-
-\pagebreak
-
-\begin{figure*}[tb]   
-  \begin{center}
-\includegraphics[width=\linewidth]{figs2_jri}
-   \caption{Heterozygosity ($H$) at cycle 16 across all ten chromosomes in each population, calculated on 15-marker sliding windows with 5 marker steps as in Figure 3.
-Heterozygosity values in BSSS (blue dots) and BSCB1 (red dots) are superimposed in one panel.
-2 cM windows of low heterozygosity that were reproduced in less than 10 of 10,000 simulations (P<0.001) are shaded in light blue (BSSS) or pink (BSCB).
-At this cutoff, there is no overlap between highlighted windows in the two populations.
-(A) Physical map.
-(B) Genetic map.
-}
+\vspace{-6mm}
     \label{fig:s2}
@@
 \end{figure*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
 

### 2015-07-03 response.tex
more responses

@@
 
-\newcommand{\jri}[1]{\textcolor{blue}{ \emph{\scriptsize  #1}} }
-\newcommand{\jristop}{\textcolor{red}{\hline \bf JRI STOPPED HERE\\ \hline}}
-\newcommand{\st}[1]{\textcolor{red}{ #1}}
-\newcommand{\comst}[1]{\textcolor{red}{ \em{\scriptsize  (#1)}} }
-\definecolor{mattgreen} {rgb} {0,0.6,0}
-\newcommand{\mbh}[1]{\textcolor{mattgreen}{ \em{\scriptsize  #1}} }
-\definecolor{peterpurple}{rgb}{.6,0,.6}
-\newcommand{\plr}[1]{\textcolor{peterpurple}{ \emph{\scriptsize (#1)}} }
+\newcommand{\jri}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} }
+\newcommand{\citex}[1]{\textcolor{red}{\bf #1}}
+\newcommand{\blah}{\texcolor{red}{\bf blah}}
 
@@
 
-\response{We feel that while an experiment-wide p-value would be important if we were performing a formal genome scan, our purpose here was simply to show that while our overall power to distinguish selection from drift may be low, there are nonetheless individual genomic regions that show patterns of fixation suggestive of selection.
-We report the \% of simulations for each region that showed lower heterozygosity in simulations, so a simple bonferroni or other calculation could be calculated by a reader concerned we are not sufficiently conservative.
+\response{In the last paragraph of the results we report 1-sided tests for a few large haplotypes.
+We agree with the reviewer that an experiment-wide p-value might be useful if we were performing a formal genome scan, but because our purpose was simply to show that although the  overall power to distinguish selection from drift may be low, there are nonetheless individual genomic regions that show patterns of fixation suggestive of selection.
 \jri{Do we want to do an FDR for the p-values in Figure 3?}}
@@
 
-\response{blah}
+\response{\blah}
 
@@
 
-\response{blah}
+\response{This is a good point and we have added this to the text.}
 
@@
 
-\response{This is a good point and we have added a few words in the intro in the hopes of making this clearer.}
+\response{This is a good point and we have added mention of this in the introduction.}
 
@@
 
-\response{We had avoided citing \citet{jiao2012genome} because we were  unsure of the veracity of the results (see \citet{lorenz2013phylogenetic}) and concerned by the fact that a wholesale recoding of their genotyping data appeared to have little impact on their conclusions.
-We have now added a citation, however, as it is certainly true that they performed similar analyses.}
+\response{While we agree with the reviewer's hesitation (see \citet{lorenz2013phylogenetic}), we have added a citation to  \citet{jiao2012genome}.}
 
@@
 
-\response{blah}
+\response{Some of these regions are visible in Figure 4. We have now included plots similar to figure 4 for all chromosomes in the supplement. \jri{Justin is this doable?}}
 
-The authors interpret the rate of heterozygosity in terms of selection and 
-drift.  I am assuming that they are interpreting selection to mean the selection 
-for advantageous alleles in hybrid test combinations.  However, I wonder about 
-the inadvertent selection to maintain heterozygosity?  In work from McMullen and 
-Buckler it was argued that the gain in vigour due to heterozygosity of 
-pericentromeric regions was larger than expected and that this may be due to 
-increased ability to cross and maintain these plants.  Was this concept included 
-in the modeling of the data?  In other words, while drift will tend to increase 
-homozygosity in a population there may be counteracting forces of heterosis 
-effects that will maintain heterozygosity in the same population.  Just curious 
-about this concept.
+The authors interpret the rate of heterozygosity in terms of selection and drift.
+I am assuming that they are interpreting selection to mean the selection for advantageous alleles in hybrid test combinations.
+However, I wonder about the inadvertent selection to maintain heterozygosity?
+In work from McMullen and Buckler it was argued that the gain in vigour due to heterozygosity of pericentromeric regions was larger than expected and that this may be due to increased ability to cross and maintain these plants.
+Was this concept included in the modeling of the data?
+In other words, while drift will tend to increase homozygosity in a population there may be counteracting forces of heterosis effects that will maintain heterozygosity in the same population.
+Just curious about this concept.
 
-\response{blah}
+\response{Yes, this is indeed a possibilitiy and one we mentioned in the text as a possible explanation for the fact that the observed data show higher heterozygosity than the simulations in early generations.
+We have now added citations to \citet{McMullen2009} and \citet{Gore2009a} in this sentence to help highlight the point.}
 
@@
 experiment was done with a maize (or teosinte) population that had never been 
-passed through an inbred “bottleneck” that there was be some alleles of major 
+passed through an inbred ``bottleneck'' that there was be some alleles of major 
 effect that would be purged quickly and that the role of drift and selection 
@@
 
-\response{blah}
+\response{This is an excellent point.
+We currently have experiments underway to test preciesly this point using selfed progeny of both landrace and teosinte populations.
+We have added discussion of this point to the text.}
 
@@
 Girke et al.
-in "The genomic impacts of drift and selection for hybrid performance in maize" report a retrospective examination of genomic patterns of diversity in a maize experimental population.
-The data
-set consists of genotyping of ~38,000 SNPs genotyped in individuals from various phases of the selection 
-experiment. While the manuscript consists primarily of population genetic analysis, the authors attempted 
+in ``The genomic impacts of drift and selection for hybrid performance in maize'' report a retrospective examination of genomic patterns of diversity in a maize experimental population.
+The data set consists of genotyping of ~38,000 SNPs genotyped in individuals from various phases of the selection experiment.
+While the manuscript consists primarily of population genetic analysis, the authors attempted
 to better integrate a popular maize genetic map with the current reference genome.
@@
 2?) recombination events per arm generation was included in the simulation, with probability of 
-recombination at a location determine by genetic map length. Something like this needs to be stated. Also 
-related to the simulations, there are multiple references in the text to empirical results as "deviation from 
+recombination at a location determine by genetic map length. Something like this needs to be stated.
+
+\response{We have added three sentences to the methods describing additional details of the recombination model.
+More information about the specific functions as well as detailed code is available at the url link to the hypred R package.}
+
+Also related to the simulations, there are multiple references in the text to empirical results as "deviation from 
 simulation" (e.g., p 15) when I suspect what the authors wish to say is that the results differ from neutral 
@@
 
-\response{blah}
+\response{We have removed the phrase ``deviation from simulation'' throughout, replacing it with phrases mentioning differences from neutral simulations.}
 
@@
 
-\response{blah}
+\response{\jri{Jode?}}
 
@@
 
-\response{blah}
+\response{\jri{Jode? Justin?}}
 
@@
 - Careful copy editing is needed. Heterozygsoity in the figures and figure captions is referred to variously 
-as "H" with and without italics and also as heterozygosity spelled out, sometimes within th esame figure or 
-figure caption. Simiarily, Fst needs to be presented in a consistent manner. Also edit citations for 
-consistent capitalization and in=text citations for consistent spacing. Also:
+as "H" with and without italics and also as heterozygosity spelled out, sometimes within the same figure or 
+figure caption. Simiarily, Fst needs to be presented in a consistent manner.
+
+\response{We have changed text and labels to improve consistency in using heterozygosity vs ``H''.}
+
+Also edit citations for consistent capitalization and in=text citations for consistent spacing.
+
+\response{We have moved the ms from word to latex; we believe this has fixed many of these issues. We have also corrected bibtex entries in several places to improve consistency of capitalization.}
+
+Also:
 On page 7, line 10: A sentence should not begin with a number.
+
+\response{Changed.}
+
 Page 10, line 14: Citation format is incorrect.
+
+\response{Fixed.}
+
 Page 11, line 16: Number of SNPs passing filters should be reported in ‘Materials and Methods’
+
+\response{This is mentioned in the very beginning of the results, where we felt it would be easier to find for readers that may skim the methods rapidly.}
+
 Figure 2 caption: What is “H combined” and how was it calculated?
+
+\response{This is simply total heterozygosity pooling across all plants. Text explaining this has been added to the figure legend.}
+
 Figure 4 and 5: Please include the simulated data (mean + variance) so the pattern is more apparent.
 
-\response{blah}
+\response{\blah}
 
@@
 "Physical Map" and the cycles as "Cycle."
+
+\response{We have tried to improve figure labelling throughout.
+In places instead of adding labels (e.g.
+Fig 4 mentioned by the reviewer) we have added additional explanatory text to the caption in hopes of clarifying.}
 
@@
 
-\response{blah}
+\response{Four plants were found that appeared noticeably out of place on the PCA.
+For two of these we have insufficent evidence to know whether they were switched, mislabelled, contamination, etc.
+These were removed from the analysis.
+The remaining two were clearly a label switch in which ``plant 31'' from cycle 4 was labelled as cycle 8 and ``plant 31'' from cycle 8 was labelled as cycle 4.
+Because the labels corresponded to the position swap seen in the PCA, we were confident in correcting the labels for these plants and kept them in the analysis.
+}
 
@@
 
-\response{blah}
+\response{Details of the chip and validation of the markers are published in \citet{ganal2011a-large} and do not seem appropriate to repeat here.
+All SNPs have an assigned position on the reference genome, but our inspection of the genetic map led us to move some of these markers as detailed in the text.}
 
@@
 
-\response{blah}
+\response{\jri{Justin?}We have added this number to the text.}
 
@@
 
-\response{blah}
+\response{Yes. While a multi-generation Ne (e.g. harmonic mean of N) will not immediately double, the single-generation Ne has effectively doubled.}
 
@@
 
-\response{blah}
+\response{Done}
 
@@
 
-\response{blah}
+\response{Hudson et al.
+1992 Fst is based on pairwise differences between sequences.
+We can thus calculate this from SNPs in a window (or individual SNPs) without knowing anything about the number of bp.
+}
 
@@
 
-\response{blah}
+\response{We have deleted ``line'' here in order to clarify. No distinction was intended.}
 
@@
 
-\response{blah}
+\response{``selection practices'' has been replaced with ``experimental design'' for clarity}
 

### 2015-07-03 main.tex
more responses

@@
 \newcommand{\jpg}[1]{\textcolor{gerkegreen}{ \em{\scriptsize  #1}} }
-\newcommand{\citex}[1]{\textcolor{red}{\bf #1}}
+
 \newcommand{\rev}[1]{\textcolor{blue}{\bf #1}}
@@
 These plants represent descendants of the original populations which have been randomly mated to maintain seed. 
-We also genotyped the founder inbreds for each population, with the exception of lines F1B1, CI.617, WD456, and K230 for which seed was not available. 
+We also genotyped the founder inbreds for each population, with the exception of F1B1, CI.617, WD456, and K230 for which seed was not available. 
 The data for founder line CI.540 was not used because the genotyped material was heterozygous. 
@@
 
-We called 48,919 SNPs on the Illumina platform from the MaizeSNP50\_B.egt cluster file. Genotypes with quality scores of 50 or less were recoded as missing data. 
+\rev{We called 48,919 SNPs} on the Illumina platform from the MaizeSNP50\_B.egt cluster file. Genotypes with quality scores of 50 or less were recoded as missing data. 
 Three plants were removed from the data due to an excess of missing data (the derived line B10, a cycle 0 plant from BSSS, and a cycle 8 plant from BSSS). 
@@
 These cases included both simple reversals of the physical map relative to the genetic map, and also the assignment of blocks of markers to the wrong linkage group, which we refer to as mis-mapped blocks.
-SNPs at these loci were reordered to match our genetic map. To maintain collinearity between the genetic and physical map, the physical positions of these SNPs were reassigned as follows. 
+The \rev{XXX} SNPs at these loci were reordered to match our genetic map. 
+To maintain collinearity between the genetic and physical map, the physical positions of these SNPs were reassigned as follows. 
 Individual mis-mapped markers small reversals and mis-mapped blocks ($<10$ kb) were removed from the data. 
@@
 The genotypes of these simulated random matings are the final results of each simulation and were analyzed in the same way as the observed data. 
+\rev{Simulated recombination was carried out in R with the hypred software package (\url{cran.r-project.org/web/packages/hypred/}).
+The number of crossovers between two parental gametes is drawn from a Poisson distribution with $\labmda=L$, where L is the length of the chromosome in Morgans.
+Crossover breakpoints are drawn from a uniform distribution over the interval $(0,L)$}
 
-All recombination events in the simulations were carried out in R with the hypred software package (\url{cran.r-project.org/web/packages/hypred/}). 
+The meiosis is simulated according to the count-location model that assumes no interference and agrees with the Haldane mapping function.[c]1 Here the number of crossing over on a random meiotic product of a chro- mosome is Poisson distributed with parameter λ = L (L being the length of the chromosome in Morgan).
+The locations of the crossing over follow a uniform distribution over the interval (0, L).
+
 Simulations were executed in parallel on a computing cluster, with unique random number seeds drawn for each simulation. 
@@
    \vspace{-3mm}
-   \caption{Heterozygosity and $F_{ST}$ across cycles. $H$ (left panel) and $F_{ST}$ (right panel) plotted as a function of selection cycle in each population.
+   \caption{Heterozygosity \rev{(H, left panel) and $F_{ST}$ (right panel)} plotted as a function of selection cycle in each population.
+\rev{Heterozygosity is shown for each population separately and then for the total population pooling all plants.}
 } 
@@
 \includegraphics[width=\linewidth]{figs2_jri}
-   \caption{Heterozygosity ($H$) at cycle 16 across all ten chromosomes in each population, calculated on 15-marker sliding windows with 5 marker steps.
+   \caption{Heterozygosity at cycle 16 across all ten chromosomes in each population, calculated on 15-marker sliding windows with 5 marker steps.
 Heterozygosity values in BSSS (blue dots) and BSCB1 (red dots) are superimposed in one panel.
@@
 (B) Genetic map.
-\rev{The same data plotted are plotted separately for each population in Figure \ref{fig:s2}.}}
+\rev{The same data are plotted separately for each population in Figure \ref{fig:s2}.}}
     \label{fig:heterotic}
@@
  %  \vspace{-3mm}
-   \caption{Heterozygosity ($H$) in each cycle across chromosome 9 of the BSSS (left) and chromosome 4 of the BSCB1 (right) plotted on the physical (top) and genetic (bottom) map.
-$H$ is calculated on 15-marker sliding windows, with 5 marker steps between each calculation.
+   \caption{\rev{Heterozygosity} in each cycle across chromosome 9 of the BSSS (left) and chromosome 4 of the BSCB1 (right) plotted on the physical (top) and genetic (bottom) map.
+\rev{Each panel consists of five plots representing cycles 0,4,8,12, and 16 of the experiment.} \rev{Heterozygosity} is calculated on 15-marker sliding windows, with 5 marker steps between each calculation.
 Each data point is color-coded based on a linear transformation of recombination rate (red = low recombination rate).
@@
 Averaged across the genome, the vast majority of the reduction in diversity observed in both populations can be attributed to genetic drift. 
-Nonetheless, we do observe deviations from the simulated values (Figure \ref{fig:compare_to_sims}). 
-The observed data show higher than expected heterozygosity at cycle 4, suggesting selection for heterozygosity in early cycles, a deviation from simulated and actual breeding practices, or an under-sampling of the diversity present in the original cycle 0 populations.
-\jri{where is this excess of diversity?
+Nonetheless, we do observe \rev{differences from values generated under our neutral simulations} (Figure \ref{fig:compare_to_sims}). 
+The observed data show higher than expected heterozygosity at cycle 4, \rev{which could be explained by several factors, including differences between simulated and actual breeding practices, under-sampling of diversity in cycle 0, selection acting to increase the frequency of initially rare haplotypes, or even selection acting to maintain heterozygosity \citep[e.g.][]{McMullen2009,Gore2009a} } \jri{where is this excess of diversity?
 if it's all pericentromeric can we argue for the first hypothesis?}
+
 As the cycles progress, heterozygosity falls more rapidly than expected in both populations, and the observed values at cycle 16 are significantly lower than the simulated data.  
@@
 To examine the behavior of specific regions, we also calculated the observed and simulated results for each 2 cM segment of the genome. 
-The dynamics observed across most of the genome are largely insensitive to changes in window size (we tested from 2-4 cM, data not shown), and are consistent with strong genetic drift imposed by selection practices.
+The dynamics observed across most of the genome are largely insensitive to changes in window size (we tested from 2-4 cM, data not shown), and are consistent with strong genetic drift imposed by \rev{the experimental design}.
 \jri{did we test smaller regions?
@@
 Since the population size of the Iowa RRS is small (10-20), many biallelic SNPs should fix by chance regardless of their starting minor allele frequencies.
-The deviation from simulation arises not from changes in allele frequencies per se, but rather from the fixation of linked markers across larger than expected genetic distances.
+\rec{Observed differences from the simulated neutral expectation} thus do not arise from changes in allele frequencies per se, but rather from the fixation of linked markers across larger than expected genetic distances.
 The validity of significance cutoffs therefore depend on the accuracy of our genetic map, and the maize genetic map is known to vary among genetic backgrounds across short genetic distances ($<5$cM) (McMullen et al., 2009).
@@
 variable is somewhat subjective}
-Deviations of the observed and simulated results could be due to selection, variation or inaccuracy in the genetic map, or a combination of these factors.
+\rev{Differences between} observed and simulated results could be due to selection, variation or inaccuracy in the genetic map, or a combination of these factors.
 
@@
 \section*{Discussion}
+\jri{need to mention that the complementation we are seeing is not large effect deleterious that were likely purged during early inbreeding but instead many weakly deleterious things.}
 
@@
 
+\subsection*{Heterosis}
+
 The observation that the same targets of selection are not observed in the opposing heterotic populations bears implications for the genetic mechanisms responsible for heterosis and the success of maize hybrids.

### 2015-07-03 main.tex
...and now it compiles.

@@
 \usepackage{multirow}
+\usepackage{amsmath}
 
@@
 \rev{Simulated recombination was carried out in R with the hypred software package (\url{cran.r-project.org/web/packages/hypred/}).
-The number of crossovers between two parental gametes is drawn from a Poisson distribution with $\labmda=L$, where L is the length of the chromosome in Morgans.
+The number of crossovers between two parental gametes is drawn from a Poisson distribution with $\lambda = L$, where $L$ is the length of the chromosome in Morgans.
 Crossover breakpoints are drawn from a uniform distribution over the interval $(0,L)$}
-
-The meiosis is simulated according to the count-location model that assumes no interference and agrees with the Haldane mapping function.[c]1 Here the number of crossing over on a random meiotic product of a chro- mosome is Poisson distributed with parameter λ = L (L being the length of the chromosome in Morgan).
-The locations of the crossing over follow a uniform distribution over the interval (0, L).
 
@@
   \begin{center}
-\includegraphics[width=\linewidth]{figs2_jri}
+\includegraphics[width=\linewidth]{fig_S2_combined}
    \caption{Heterozygosity at cycle 16 across all ten chromosomes in each population, calculated on 15-marker sliding windows with 5 marker steps.
@@
 Since the population size of the Iowa RRS is small (10-20), many biallelic SNPs should fix by chance regardless of their starting minor allele frequencies.
-\rec{Observed differences from the simulated neutral expectation} thus do not arise from changes in allele frequencies per se, but rather from the fixation of linked markers across larger than expected genetic distances.
+\rev{Observed differences from the simulated neutral expectation} thus do not arise from changes in allele frequencies per se, but rather from the fixation of linked markers across larger than expected genetic distances.
 The validity of significance cutoffs therefore depend on the accuracy of our genetic map, and the maize genetic map is known to vary among genetic backgrounds across short genetic distances ($<5$cM) (McMullen et al., 2009).

### 2015-07-06 response.tex
responses

@@
 \newcommand{\citex}[1]{\textcolor{red}{\bf #1}}
-\newcommand{\blah}{\texcolor{red}{\bf blah}}
+\newcommand{\blah}{\textcolor{red}{\bf blah}}
 
@@
 \fancyhf{}
-\fancyfoot[CE,CO]{\thepage}
+\fancyfoot[C,CO]{\thepage}
 \renewcommand{\headrulewidth}{0pt}
@@
 
-\response{\blah}
+\response{While we are not making a formal hypothesis test of the overdominance model (which seems difficult to do as it would depend on the number of loci proposed to be overdominant), we believe the preponderance of the evidence supports a simple dominance model.
+While we can not rule out any overdominance -- and have now added text citing one such example -- the genome-wide pattern does not appear to us to support such a model as the major driver of heterosis for yield.}
 
@@
 \section*{Reviews}
-\subsection*{Reviewer 1}
+\subsection{Reviewer 1}
 I really enjoyed reading this manuscript.  I think this is a well done study 
@@
 
-\response{\blah}
+\response{The abstract has been extensivley edited to better reflect our findings, and references in the abstract have been removed.}
 
@@
 
-\response{\jri{Jode? Justin? I don't have the answer here, but suspect it is ``no''}}
+\response{The major difference is the change to 20 (instead of 10) lines at cycle 8.
+Machine harvest was introduced at cycle 6, but we have no reason to believe this would have dramatically changed diversity patterns.
+From our reading of the literature and consultation with breeders who have run the experiment we are not aware of any other significant changes.
+}
 
@@
 
-\response{\jri{Justin, have you done this? Easy to do?}}
+\response{PCA of just the founder lines does not show differentiation corresponding to the two populations. \jri{Justin, can you do this quick and dirty to confirm?}}
 
@@
 
-\response{\jri{Justin, any ideas? Do we do some formal correspondence with NAM results? Or say ``geez this is hard because QTL are big and we're not sure we'd believe the results?''}
+\response{\jri{Justin, any ideas? Do we do some formal correspondence with NAM results? Or say ``geez this is hard because QTL are big and we're not sure we'd believe the results?''}}
 
@@
 
-\subsection*{Reviewer 2}
+\subsection{Reviewer 2}
 Girke et al.
@@
 
-\response{\jri{Jode?}}
+\response{\jri{Justin? Jode?}}
 
@@
 
-\response{\blah}
+\response{This was an oversight on our part; we have now incorporated citation to some of this literature and hopefully clarified the discussion.}
 
@@
 
-\response{\blah}
+\response{We have endeavored to streamline the discussion where possible while incorporating additional citations (as above) and discussion of load as mentioend by reviewer 1.}
 
@@
 
-\response{\blah}
+\response{We were unable to devise a means of adding simulated data to the plot in the new Figure 4 without substantially obfuscating the main points of the figure.
+We have however marked individual regions which have lower heterozygosity than seen in our simulations.}
 

### 2015-07-06 main.tex
responses

@@
 Modern maize breeding relies upon selection in inbreeding populations to improve the performance of cross-population hybrids. 
-The United States Department of Agriculture - Agricultural Research Service (USDA-ARS) reciprocal recurrent selection experiment between the Iowa Stiff Stalk Synthetic (BSSS) and the Iowa Corn Borer Synthetic No.
+The United States Department of Agriculture reciprocal recurrent selection experiment between the Iowa Stiff Stalk Synthetic (BSSS) and the Iowa Corn Borer Synthetic No.
 1 (BSCB1) populations represents one of the longest standing models of selection for hybrid performance.
-To investigate the genomic impact of this selection program, we used the Illumina MaizeSNP50 high-density SNP array to determine genotypes of progenitor lines and over 600 individuals across multiple cycles of selection.
-Consistent with previous research \citep{messmer1991genetic, labate1997molecular,  hagdorn2003molecular, hinze2005population}, we found that genetic diversity within each population steadily decreases, with a corresponding increase in population structure.
-High marker density also enabled the first view of haplotype ancestry, fixation and recombination within this historic maize experiment. 
-Extensive regions of haplotype fixation within each population are visible in the pericentromeric regions, where large blocks trace back to single founder inbreds. 
-Simulation attributes most of the observed reduction in genetic diversity to genetic drift. 
-Signatures of selection were difficult to observe in the background of this strong genetic drift, but heterozygosity in each population has fallen more than expected. 
-Regions of haplotype fixation represent the most likely targets of selection, but as observed in other germplasm selected for hybrid performance \citep{feng2006temporal}, there is no overlap between the most likely targets of selection in the two populations.
+To investigate the genomic impact of this selection program, \rev{we genotyped the progenitor lines and over 600 individuals across multiple cycles of selection using a genome-wide panel of $\sim 40,000$ SNPs.
+We first confirm previous results showing a steady temporal decrease in genetic diversity within populations and a corresponding increase in differentiation between populations.  
+Thanks to detailed historical information on experimental design, we were able to perform extensive simulations using founder haplotypes to replicate the experiment in the absence of selection.
+These simulations demonstrate that while most of the observed reduction in genetic diversity can be attributed to genetic drift,   
+heterozygosity in each population has fallen more than expected. 
+We then take advantage of our high-density genotype data to identify extensive regions of haplotype fixation and trace haplotype ancestry to single founder inbred lines.
+The vast majority of regions showing such evidence of selection differe between the two populations, providing evidence against the overdominance model as a means of explaining selection for hbyrid vigor.
+}
 We discuss how this pattern is likely to occur during selection for hybrid performance, and how it poses challenges for dissecting the impacts of modern breeding and selection on the maize genome. }
@@
 BSSS and BSCB1 have been recurrently selected for improved cross-population hybrids \citep{penny1971twenty}. 
-\jri{add some text explaining recurrent selection}
+\jri{Jode/Justin add text explaining recurrent selection and populations}
 This model of selection, named reciprocal recurrent selection, provides the generalized model for strategies commonly used in commercial maize breeding \citep{comstock1949breeding, duvick2004long}. 
@@
 Since there was no evidence suggesting when mis-labelings occurred, each of these plants was removed from the analysis.
+\rev{The final genotyping data set is available at \jri{Justin, please send me this and I can put online}.}
 
@@
   \begin{center}
-\includegraphics[width=\linewidth]{fig_S2_combined}
+\includegraphics[width=0.7\linewidth]{fig_S2_combined}
    \caption{Heterozygosity at cycle 16 across all ten chromosomes in each population, calculated on 15-marker sliding windows with 5 marker steps.
@@
   % \vspace{-0mm}
-   \includegraphics[width=\linewidth]{fig4_combined}
+   \includegraphics[width=0.8\linewidth]{fig4_combined}
  %  \renewcommand{\baselinestretch}{0.9}
@@
 Selection was executed at random in each simulation, so the patterns observed across simulations represent the expected distribution of effects caused only by recombination and genetic drift. 
-We conducted 10,000 simulations, with genetic recombination modeled on a framework map derived from the IBM population \citep{lee2002expanding}, a cross between inbred lines B73 (a BSSS derived line used to construct the reference genome) and Mo17 (an inbred with some co-ancestry to both the BSSS and BSCB1 founder germplasm).
-We sampled individuals from each cycle equal in number to the observed samples, measured heterozygosity, and compared the results to the observed data.
+We conducted 10,000 simulations, \rev{modeling recombination using the IBM genetic map \citep{lee2002expanding}, which is based on a cross between between a BSSS-derived inbred line used to construct the reference genome (B73) and an inbred with  with co-ancestry from both the BSSS and BSCB1 founder germplasm (Mo17).}
 
@@
 
-To examine the behavior of specific regions, we also calculated the observed and simulated results for each 2 cM segment of the genome. 
-The dynamics observed across most of the genome are largely insensitive to changes in window size (we tested from 2-4 cM, data not shown), and are consistent with strong genetic drift imposed by \rev{the experimental design}.
-\jri{did we test smaller regions?
-not enough markers?}
+To examine the behavior of specific regions, we compared observed and simulated results for each 2 cM segment of the genome. 
+The dynamics observed across most of the genome are largely insensitive to window size (we tested from 2-4 cM, data not shown), and are consistent with strong genetic drift imposed by \rev{the experimental design}.
 A subset of loci were flagged as significant, and these loci almost always overlapped regions of fixation or near-zero heterozygosity in one population. 
-However, in these regions the values obtained by simulation are often quite low, such that the extreme simulated values sometimes lie only a few percentage points away from the observed value (Supplementary File S2), indicating that drift alone can explain most of the drop in diversity.
+\rev{Simulated values in these regions are often quite low as well (Supplementary File S2), however, indicating that drift alone can explain most of the drop in diversity.}
 
@@
 \rev{Observed differences from the simulated neutral expectation} thus do not arise from changes in allele frequencies per se, but rather from the fixation of linked markers across larger than expected genetic distances.
-The validity of significance cutoffs therefore depend on the accuracy of our genetic map, and the maize genetic map is known to vary among genetic backgrounds across short genetic distances ($<5$cM) (McMullen et al., 2009).
-\jri{need to cite either \citep{bauer2013intraspecific} who shows there's lots of variation or maybe also \citep{rodgers2015recombination} who argues it's stable (but stable vs.
-variable is somewhat subjective}
+The validity of significance cutoffs therefore depend on the accuracy of our genetic map.
+While the maize genetic map is known to vary among genetic backgrounds across short distances \citep{McMullen2009}, broad-scale patterns of recombination appear relatively stable across diverse germplasm \citep[][but see \citet{bauer2013intraspecific}]{rodgers2015recombination}.
 \rev{Differences between} observed and simulated results could be due to selection, variation or inaccuracy in the genetic map, or a combination of these factors.
@@
 Although these results suggest that selection may have pushed these haplotypes to fixation, the fact that fixation of such a rare haplotype still occurred in some simulations speaks to the strong genetic drift imposed upon the BSSS and BSCB1 populations.
-\jri{do you think we need to do more to explain here or above?
-in the previous paragraph we argue biallelic loci should fix by chance regardless of starting allele freq.}
 Interestingly, each of these two genomic regions harbored a different cycle 0 haplotype at higher frequency, but these higher-frequency haplotypes were subsequently lost within the RRS population. 
@@
 \section*{Discussion}
-\jri{need to mention that the complementation we are seeing is not large effect deleterious that were likely purged during early inbreeding but instead many weakly deleterious things.}
+\jri{need to mention that the complementation we are seeing is not large effect deleterious that were likely purged during early inbreeding but instead many weakly deleterious things.
+need to add Drosophila and other cites.
+need to add Sprague 1983.
+then need to shorten/tighten up.}
 
@@
 
-Reciprocal recurrent selection serves as a model for the method of hybrid maize improvement employed broadly in North America \citep{duvick2004long}.
-The trends observed in the Iowa RRS are also a common theme in similar recurrent selection experiments \citep{romay2012effect,lamkey2014relative}.
-Increased population structure due to increased inbreeding and lowered effective population size are visible in the modern ‘stiff-stalk’ and ‘non stiff-stalk’ heterotic groups as well \citep{van2012historical}.
-Given the lack of strong selection signals in each case, genetic drift has most likely played a large role in the current genetic structure of modern maize.
+\rev{The trends observed in the Iowa RRS are reflected in similar recurrent selection experiments \citep{romay2012effect,lamkey2014relative}.
+Reciprocal recurrent selection serves as a model for the method of hybrid maize improvement \citep{duvick2004long}, and similar patterns of diversity and population structure can be seen broadly across North American maize germplasm \citep{van2012historical}.
+Given the lack of strong selection signals in each case, genetic drift has most likely played a large role in the current genetic structure of modern maize. }
 
@@
 B37 and B14 were derived from cycle 0, and B73 was derived from a half-sib recurrent selection program also started with the BSSS population. 
-We examined these three inbreds at the pericentromeric regions listed in Table \ref{tabfix} \jri{correct?}, and found that in most cases they carry different haplotypes from those that rose to high frequency in the RRS experiment.
+We examined these three inbreds at the pericentromeric regions listed in Table \ref{tab:fix} \jri{correct?}, and found that in most cases they carry different haplotypes from those that rose to high frequency in the RRS experiment.
 Near isogenic lines that substitute these centromeric haplotypes from key inbreds could be used to determine if the haplotypes at high frequency in cycle 16 of BSSS provide a phenotypic advantage when crossed with BSCB1.
 
-Both our study and previous work \citep{labate1997molecular} identified genetic differentiation between cycle 0 of BSSS and the founder inbreds. 
-The loss of diversity in the cycle 0 population likely led to the loss of some rare alleles and haplotypes. 
-The BSSS was maintained for several years before RRS was initiated, and the differentiation could have occurred due to low effective population size during maintenance. 
-If so, then this drift will have impacted the trajectory of the RRS program and of the key inbreds derived from BSSS. 
-However, cycle 0 has also undergone maintenance since the beginning of the RRS experiment, so some genetic drift may have occurred during this time as well.
+Both our study and previous work \citep{labate1997molecular} identified genetic differentiation between cycle 0 of BSSS and the founder inbreds, \rev{likely reflecting the loss of some rare alleles.
+If this differentiation reflects drift during population maintenance before initiation of the RRS, it may have impacted the trajectory of the RRS program and of key inbreds derived from BSSS. 
+However, cycle 0 has also undergone maintenance since the beginning of the RRS experiment, so some} drift may have occurred during this time as well.
 
@@
 Classic overdominance models of heterosis predict that at a single locus, two distinct alleles confer heterozygote advantage when combined, while the dominance model predicts that heterosis is driven by dominance effects and the complementation of linked alleles in low-recombination regions (dominance or pseudo-overdominance).
-In the case of true overdominance, we expect selection should lead to decreased heterozygosity at a locus in both populations as complementary haplotypes are fixed in each group. 
-We find no evidence of this genetic phenomenon. 
+In the case of true overdominance, we selection should lead to decreased heterozygosity at a locus in both populations as complementary haplotypes are fixed in each group \rev{\citep[e.g.][]{guo2014maize}}.
+We find \rev{little evidence to support this genetic phenomenon, finding only two 2cM windows genome-wide in which both populations show significantly reduced heterozygosity.} 
 The observed pattern instead favors a dominance model, where fixation of a haplotype in one population simply selects against that same haplotype in the other population. 

### 2015-07-07 Fig4/allchroms.R
corrected figure 4, made a 'figure 4' for each chromosome

@@ -1,0 +1,95 @@
+#!/usr/bin/Rscript
+
+rm(list=ls())
+library(ggplot2)
+library(dplyr)
+
+loadcycle = function(cycle, mychrom){
+  infile = paste("../Fig3/data/H_cycle", cycle, "_LG", mychrom, ".txt", sep='')
+  myfile = read.table(infile, header=T)
+  return(myfile)
+}
+
+loadchrom = function(chrom){
+  cycles = c(0,4,8,12,16)
+  tables = lapply(cycles, loadcycle, mychrom=chrom) 
+  results = do.call(rbind, tables)
+  return(results)
+}
+
+chromdata_H = lapply(1:10, loadchrom)
+
+simresults = read.table("../data/2cm_sim_collated_results.txt", header=T)
+
+
+#The code to make plots is ugly because ggplot has some unpredictable (at least to me) behavior when wrapped in functions due to scope issues.
+####code to make plot for chromosome 9 BSSS.  ####
+for (x in 1:10) {
+  Chrom = x
+
+Hetgrp="BSSS"
+Hdata = chromdata_H[[Chrom]]
+mytitle = paste("Chromosome ", Chrom, ", ", Hetgrp, sep="" )
+sigregions001 = filter(simresults, Sigat001 == "BSSS" | Sigat001 == "Both") %>% filter(LG==Chrom)
+sigregions0025 = filter(simresults, Sigat0025 == "BSSS" | Sigat0025 == "Both") %>% filter(LG==Chrom)
+sigat001rect_phys <-  geom_rect(data=sigregions001, aes(xmin=phystart/1000000, xmax=phystop/1000000, ymin=0, ymax=0.5), fill="lightskyblue") 
+#sigat0025rect_phys <- geom_rect(data=sigregions0025, aes(xmin=phystart/1000000, xmax=phystop/1000000, ymin=0, ymax=0.5), fill="red") 
+Hplot_phys <- geom_point(size = 1, data=Hdata, aes(x=physpos/1000000, y=HetSS, colour=sqrt(rate^(1/4)))) 
[…]

### 2015-07-07 Fig4/fig4.R

@@ -32,19 +32,19 @@ simresults = read.table("../data/2cm_sim_collated_results.txt", header=T)
   
   sigregions001 = filter(simresults, Sigat001 == "BSSS" | Sigat001 == "Both") %>% filter(LG==Chrom)
   sigregions0025 = filter(simresults, Sigat0025 == "BSSS" | Sigat0025 == "Both") %>% filter(LG==Chrom)
-	sigat001rect_phys <-  geom_rect(data=sigregions001, aes(xmin=phystart/1000000, xmax=phystop/1000000, ymin=0, ymax=0.5), fill="lightskyblue1") 
-	sigat0025rect_phys <- geom_rect(data=sigregions0025, aes(xmin=phystart/1000000, xmax=phystop/1000000, ymin=0, ymax=0.5), fill="lightskyblue") 
+	sigat001rect_phys <-  geom_rect(data=sigregions001, aes(xmin=phystart/1000000, xmax=phystop/1000000, ymin=0, ymax=0.5), fill="lightskyblue") 
+#	sigat0025rect_phys <- geom_rect(data=sigregions0025, aes(xmin=phystart/1000000, xmax=phystop/1000000, ymin=0, ymax=0.5), fill="lightskyblue") 
 	Hplot_phys <- geom_point(size = 1, data=Hdata, aes(x=physpos/1000000, y=HetSS, colour=sqrt(rate^(1/4)))) 
-	myphysplot=ggplot(data=Hdata) + facet_grid(Cycle~.) + sigat001rect_phys + sigat0025rect_phys + Hplot_phys  + xlab("Position(Mb)")+ ylab("Heterozygosity") + scale_colour_gradient2(low="red", mid="orange", high="yellow", midpoint=median(Hdata$rate^(1/4))) + theme_bw() + ggtitle("BSSS, Chromosome 9")
+	myphysplot=ggplot(data=Hdata) + facet_grid(Cycle~.) + sigat001rect_phys + Hplot_phys  + xlab("Position(Mb)")+ ylab("Heterozygosity") + scale_colour_gradient2(low="red", mid="orange", high="yellow", midpoint=median(Hdata$rate^(1/4))) + theme_bw() + ggtitle("BSSS, Chromosome 9")
 
[…]

### 2015-07-07 main.tex
fig4 changed, removed chr9

@@
   % \vspace{-0mm}
-   \includegraphics[width=0.8\linewidth]{fig4_combined}
+   \includegraphics[width=0.8\linewidth]{Fig4/fig4_combined}
  %  \renewcommand{\baselinestretch}{0.9}
  %  \vspace{-3mm}
-   \caption{\rev{Heterozygosity} in each cycle across chromosome 9 of the BSSS (left) and chromosome 4 of the BSCB1 (right) plotted on the physical (top) and genetic (bottom) map.
+   \caption{\rev{Heterozygosity} in each cycle across chromosome 4 of the BSSS (left) and  BSCB1 (right) plotted on the physical (top) and genetic (bottom) map.
 \rev{Each panel consists of five plots representing cycles 0,4,8,12, and 16 of the experiment.} \rev{Heterozygosity} is calculated on 15-marker sliding windows, with 5 marker steps between each calculation.

### 2015-07-07 main.tex
figures modified, added to supp.

@@
 \usepackage{amsmath}
+\usepackage{capt-of}
 
@@
 
+\newcommand\Image[3][]{%
+  \tabular[b]{@{}c@{}}\includegraphics[#1]{#2}\\
+    #3
+  \endtabular}
 \renewcommand{\textfraction}{0.01}
@@
   % \vspace{-0mm}
-   \includegraphics[width=0.8\linewidth]{Fig4/fig4_combined}
+   \includegraphics[width=0.8\linewidth]{Fig4/chrom4_combined}
  %  \renewcommand{\baselinestretch}{0.9}
@@
    \caption{\rev{Heterozygosity} in each cycle across chromosome 4 of the BSSS (left) and  BSCB1 (right) plotted on the physical (top) and genetic (bottom) map.
-\rev{Each panel consists of five plots representing cycles 0,4,8,12, and 16 of the experiment.} \rev{Heterozygosity} is calculated on 15-marker sliding windows, with 5 marker steps between each calculation.
+\rev{Each panel consists of five plots representing cycles 0,4,8,12, and 16 of the experiment.
+Heterozygosity} is calculated on 15-marker sliding windows, with 5 marker steps between each calculation.
 Each data point is color-coded based on a linear transformation of recombination rate (red = low recombination rate).
-Shaded regions represent 2 cM windows found to have heterozygosity values significantly lower than expected by simulation at a given cycle.
-Light shading denotes significance at $P<0.0025$, whereas the darker shading represents significance at $P<0.001$.
+Shaded regions represent 2 cM windows with heterozygosity values significantly lower than expected by simulation at a given cycle.
+\rev{Light and darker shading denotes significance at $P<0.0025$ and $P<0.001$, respectively.
+Plots for other chromosomes are shown in in Figure \ref{fig:others2}}.
 Comparing the physical and genetic maps reveals that the regions of fixation are extremely small genetically, but some contain much of the physical content of the chromosome.

### 2015-07-15 remapping/remapping.R
added count of mismapped markers

@@ -1,0 +1,49 @@
+rm(list=ls())
+require(dplyr)
+require(stringr)
+load("../data/new_full_set.RData")
+
+fromchris <- read.delim("../dataforupload/55kv2_sorted.tab", stringsAsFactors = F, header = T)
+
+remap <- read.delim("../dataforupload/cluster_b_remapped_calls.txt", stringsAsFactors = F, header = T)
+
+df <- data.frame(Chrom = new_full_set$LG,
+                 Position = new_full_set$position,
+                 Name = as.character(new_full_set$markernames),
+                 Code = as.character(new_full_set$markercodes),
+                 Gen = new_full_set$F2map,
+                 stringsAsFactors = F)
+
+rm(new_full_set)
+
+original = select(fromchris, Name, Code=dbSNP, Chrom_Original=chr.v2, Position_Original=pos.v2) 
+
+test = left_join(df, original, by=c("Name", "Code")) %>% 
+  mutate(Chrom_Original = as.integer(str_replace(Chrom_Original, "chr", ""))) %>%
+  mutate(Position_Original = as.integer(Position_Original))
+
+#how many markers have moved chromosomes?
+sum(test$Chrom != test$Chrom_Original)
+#[1] 15
+samechrom = filter(test, Chrom == Chrom_Original)
+sum(abs(samechrom$Position - samechrom$Position_Original>0) )
+#[1] 1585
+
+moved = c()
+for (mychrom in 1:10) {
+  currchrom = filter(samechrom, Chrom==mychrom)
+  tots = sum(abs(currchrom$Position - currchrom$Position_Original>0) )
+  moved = c(moved, tots)
+}
+
+moved
+#[1]   20   41  221    2  106   90   39    2 1063    1
+
+
+#look at chr9 out of curiosity
+chrom9 = filter(samechrom, Chrom==9) 
+chrom9_gen = left_join(chrom9, df)
+moved_9 = filter(chrom9_gen, Position != Position_Original) %>%
+  mutate(difference = abs(Position - Position_Original))
+ggplot(data=moved_9) + aes(x=Position, y=Gen) + geom_point() #+ geom_hline(yintercept=138.4)
+ggplot(data=moved_9) + aes(x=Position, y=Position_Original) + geom_point()

### 2015-07-31 main.tex
added count of mismapped markers
Updated genetic positions of Tbl_1 in main.tex

@@
 These cases included both simple reversals of the physical map relative to the genetic map, and also the assignment of blocks of markers to the wrong linkage group, which we refer to as mis-mapped blocks.
-The \rev{XXX} SNPs at these loci were reordered to match our genetic map. 
 To maintain collinearity between the genetic and physical map, the physical positions of these SNPs were reassigned as follows. 
@@
 The last SNP of the block was also given a 10 kb cushion between itself and the next SNP on the correct linkage group. 
+Among the SNPs used for analyses, 15 were mis-mapped to a different linkage group 1,585 were moved within a linkage group.
 
@@
 Population &	 Chr.& Interval (cM) & Interval (Mb) &	Founder & Derived Lines \\
-BSSS       & 3          & 40.2-41.4            & 67.7-123              & CI187-2        & B94                               \\
-BSSS       & 3          & 43-48.5              & 129.2-157.1           & NDa            & B89, B94                          \\
-BSSS       & 4          & 39.7-42.1            & 39.9-82.7             & CI187-2        & B89, B94, B67, B72, B39, B43      \\
-BSSS       & 9          & 30.6-33.5            & 20.8-26.6             & Oh7b           & B89, B94, B43, B17, B72, B84, B67 \\
-BSSS       & 9          & 34.3-35.6c           & 30.8-90.4c            & Os420          & B89, B94                          \\
-BSCB1      & 2          & 50.2-51.8            & 80.6-114.5            & CC5            & B90, B91, B95, B97, B99           \\
-BSCB1      & 4          & 42.1-42.8            & 82.7-140              & NDa            & B90, B95, B97                     \\
-BSCB1      & 8          & 46.1-50.7            & 125.1-145.6           & P8             & B90, B97, B91, B99, B54           \\\hline 
+BSSS       & 3          & 53.6-55.3            & 67.7-123              & CI187-2        & B94                               \\
+BSSS       & 3          & 57.3-64.8              & 129.2-157.1           & NDa            & B89, B94                          \\
+BSSS       & 4          & 52.8-55.6            & 39.9-82.7             & CI187-2        & B89, B94, B67, B72, B39, B43      \\
+BSSS       & 9          & 41.1-44.8            & 20.8-26.6             & Oh7b           & B89, B94, B43, B17, B72, B84, B67 \\
+BSSS       & 9          & 45.7-47.5c           & 30.8-90.4c            & Os420          & B89, B94                          \\
+BSCB1      & 2          & 67-67.5            & 80.6-114.5            & CC5            & B90, B91, B95, B97, B99           \\
+BSCB1      & 4          & 55.6-57            & 82.7-140              & NDa            & B90, B95, B97                     \\
+BSCB1      & 8          & 61.5-67.7            & 125.1-145.6           & P8             & B90, B97, B91, B99, B54           \\\hline 
 \end{tabular}

### 2015-07-31 remap_table1/remap.rmd
Updated genetic positions of Tbl_1 in main.tex

---
output: 
  html_document: 
    highlight: tango
    theme: cosmo
---
#Mapping the fixed haplotype regions to genetic coordinates.

```{r loadup, message=FALSE, warning=FALSE}
rm(list=ls())
library(dplyr)
snps = load("../data/new_full_set.RData")

#turn the dataset object into a dataframe
df <- data.frame(Chrom = new_full_set$LG,
                 Position = new_full_set$position,
                 Name = as.character(new_full_set$markernames),
                 Code = as.character(new_full_set$markercodes),
                 Gen = new_full_set$F2map,
                 stringsAsFactors = F)
```
##BSSS3.1
```{r}
chr3 = filter(df, Chrom==3)
head(chr3[order(abs(67.7e6-chr3$Position)),])
```
###Starting cM: 53.6 
```{r}
head(chr3[order(abs(chr3$Position-123e6)),], n=10)
```
###Ending cM:  55.3

##BSSS3.2
```{r}
head(chr3[order(abs(129.2e6-chr3$Position)),])
```
###Starting cM: 57.3
```{r}
head(chr3[order(abs(157.1e6-chr3$Position)),], n=10)
```
###Ending cM: 64.8

##BSSS4
```{r}
chr4 = filter(df, Chrom==4)
head(chr4[order(abs(39.9e6-chr4$Position)),],n=10)
```
###Starting cM: 52.8
```{r}
head(chr4[order(abs(82.7e6-chr4$Position)),],n=10)
```
###Ending cM: 55.6

##BSSS9.1
```{r}
chr9 = filter(df, Chrom==9)
head(chr9[order(abs(20.8e6-chr9$Position)),],n=10)
```
###Starting cM: 41.1
```{r}
head(chr9[order(abs(26.6e6-chr9$Position)),],n=10)
```
###Ending cM: 44.8

##BSSS9.2
```{r}
head(chr9[order(abs(30.8e6-chr9$Position)),],n=10)
```
###Starting cM: 45.7
```{r}
head(chr9[order(abs(90.4e6-chr9$Position)),],n=10)
```
###Ending cM: 47.5

#BSCB2
```{r}
chr2 = filter(df, Chrom==2)
head(chr2[order(abs(80.6e6-chr2$Position)),],n=10)
```
###Starting cM: 67
```{r}
head(chr2[order(abs(114.5e6-chr2$Position)),],n=10)
```
###Ending cM: 67.5

#BSCB4
```{r}
head(chr4[order(abs(82.7e6-chr4$Position)),],n=10)
```
###Starting cM: 55.6
```{r}
head(chr4[order(abs(140e6-chr4$Position)),],n=10)
```
###Ending cM: 57

##BSCB8
```{r}
chr8= filter(df, Chrom==8)
head(chr8[order(abs(125.1e6-chr8$Position)),],n=10)
```
###Starting cM: 61.5
```{r}
head(chr8[order(abs(145.6e6-chr8$Position)),],n=10)
```
###Ending cM: 67.7

### 2015-07-31 main.tex
fixed the table cuz i rock

@@
 \usepackage{capt-of}
+\usepackage{threeparttable}
 
@@
 \centering
+   \begin{threeparttable}
+
 \label{tab:fix}
@@
 BSSS       & 3          & 53.6-55.3            & 67.7-123              & CI187-2        & B94                               \\
-BSSS       & 3          & 57.3-64.8              & 129.2-157.1           & NDa            & B89, B94                          \\
+BSSS       & 3          & 57.3-64.8              & 129.2-157.1           & ND\tnote{\emph{a}}           & B89, B94                          \\
 BSSS       & 4          & 52.8-55.6            & 39.9-82.7             & CI187-2        & B89, B94, B67, B72, B39, B43      \\
-BSSS       & 9          & 41.1-44.8            & 20.8-26.6             & Oh7b           & B89, B94, B43, B17, B72, B84, B67 \\
-BSSS       & 9          & 45.7-47.5c           & 30.8-90.4c            & Os420          & B89, B94                          \\
+BSSS       & 9          & 41.1-44.8            & 20.8-26.6             & Oh7\tnote{\emph{b}}           & B89, B94, B43, B17, B72, B84, B67 \\
+BSSS       & 9          & 45.7-47.5\tnote{\emph{c}}           & 30.8-90.4             & Os420          & B89, B94                          \\
 BSCB1      & 2          & 67-67.5            & 80.6-114.5            & CC5            & B90, B91, B95, B97, B99           \\
-BSCB1      & 4          & 55.6-57            & 82.7-140              & NDa            & B90, B95, B97                     \\
+BSCB1      & 4          & 55.6-57            & 82.7-140              & ND\tnote{\emph{a}}            & B90, B95, B97                     \\
 BSCB1      & 8          & 61.5-67.7            & 125.1-145.6           & P8             & B90, B97, B91, B99, B54           \\\hline 
 \end{tabular}
+   \begin{tablenotes}
+      \footnotesize
+      \item[\emph{a}]{ND = Not Determined (Either a recombinant haplotype or originates from an un-genotyped founder)}
+      \item[\emph{b}]{Although Oh7 is a BSCB1 founder, it is a descendant of CI.540, an un-genotyped BSSS founder. BSSS segments matching Oh7 presumably derive from CI.540}            
+      \item[\emph{c}]{Founders Ind\_B2 (BSSS), CI187-2 (BSSS), R4 (BSCB1), and I205 (BSCB1) are all IBD at this region of chromosome 9}
+   \end{tablenotes}
+\end{threeparttable}
 \end{table*}
-%a ND = Not Determined (Either a recombinant haplotype or originates from an un-genotyped founder)
-%
-%b Although Oh7 is a BSCB1 founder, it is a descendant of CI.540, an un-genotyped BSSS founder. BSSS segments matching Oh7 presumably derive from CI.540
-%
-%c  Founders Ind_B2 (BSSS), CI187-2 (BSSS), R4 (BSCB1), and I205 (BSCB1) are all IBD at this region of chromosome 9
 

### 2015-08-14 response.tex
fdr

@@
 
-\response{In the last paragraph of the results we report 1-sided tests for a few large haplotypes.
-We agree with the reviewer that an experiment-wide p-value might be useful if we were performing a formal genome scan, but because our purpose was simply to show that although the  overall power to distinguish selection from drift may be low, there are nonetheless individual genomic regions that show patterns of fixation suggestive of selection.
-\jri{Do we want to do an FDR for the p-values in Figure 3?}}
+\response{We have calculated the false discovery rate for the simulation p-values reported in Fig.
+3.
+At a p-value cutoff of 0.001, FDR is 2.7\% for the BSSS and 1.7\% for the BSCB1.
+This is now reported in Fig.
+3 as a genome-wide FDR of $<3\%$.}
 
@@
 
-\response{Some of these regions are visible in Figure 4. We have now included plots similar to figure 4 for all chromosomes in the supplement. \jri{Justin is this doable?}}
+\response{Some of these regions are visible in Figure 4. We have now included plots similar to figure 4 for all chromosomes in the supplement.}
 

### 2015-08-14 main.tex
fdr

@@
 Heterozygosity values in BSSS (blue dots) and BSCB1 (red dots) are superimposed in one panel.
-2 cM windows of heterozygosity lower than 10 of 10,000 simulations ($P<0.001$) are shaded in light blue (BSSS) or pink (BSCB1).
-Two regions genome-wide show significantly low heterozygosity in both populations and are shaded green.
+2 cM windows of heterozygosity lower than 10 of 10,000 simulations ($P<0.001$) are shaded in light blue (BSSS) or pink (BSCB1), \rev{and correspond to a genome-wide false discovery rate of $<3\%$ in both populations.} Two regions genome-wide show significantly low heterozygosity in both populations and are shaded green.
 (A) Physical map.
@@
 The dynamics observed across most of the genome are largely insensitive to window size (we tested from 2-4 cM, data not shown), and are consistent with strong genetic drift imposed by \rev{the experimental design}.
-A subset of loci were flagged as significant, and these loci almost always overlapped regions of fixation or near-zero heterozygosity in one population. 
+A subset of loci were flagged as significant (Figure \ref{fig:heterotic}), and these loci almost always overlapped regions of fixation or near-zero heterozygosity in one population. 
 \rev{Simulated values in these regions are often quite low as well (Supplementary File S2), however, indicating that drift alone can explain most of the drop in diversity.}
@@
 In the case of true overdominance, we selection should lead to decreased heterozygosity at a locus in both populations as complementary haplotypes are fixed in each group \rev{\citep[e.g.][]{guo2014maize}}.
-We find \rev{little evidence to support this genetic phenomenon, finding only two 2cM windows genome-wide in which both populations show significantly reduced heterozygosity.} 
+We find \rev{little evidence to support this genetic phenomenon, finding only two 2 cM windows genome-wide in which both populations show significantly reduced heterozygosity.} 
 The observed pattern instead favors a dominance model, where fixation of a haplotype in one population simply selects against that same haplotype in the other population. 

### 2015-08-16 response.tex
added thoughts to the response and managed not to break latex

@@
 \newcommand{\jri}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} }
+\newcommand{\jpg}[1]{\textcolor{magenta}{ \emph{\scriptsize  #1}} }
 \newcommand{\citex}[1]{\textcolor{red}{\bf #1}}
@@
 
-\response{PCA of just the founder lines does not show differentiation corresponding to the two populations. \jri{Justin, can you do this quick and dirty to confirm?}}
+\response{PCA of just the founder lines does not show differentiation corresponding to the two populations. \jri{Justin, can you do this quick and dirty to confirm?}
+\jpg{Yep, I did it a while back. I don't still have the plot but there was clear structure along that axis.}}
 
@@
 
-\response{\jri{Justin, any ideas? Do we do some formal correspondence with NAM results? Or say ``geez this is hard because QTL are big and we're not sure we'd believe the results?''}}
+\response{\jri{Justin, any ideas? Do we do some formal correspondence with NAM results? Or say ``geez this is hard because QTL are big and we're not sure we'd believe the results?''}
+\jpg{I'm sure we can find them, just due to the size of the regions, but I don't think it would tell us much.
+I would be in favor of adding some more to the discussion about how the pressure for synchronous flowering time may have driven a lot of what we see.
+That said I could dig through Wallace et al.
+to see what is there.
+Thoughts?}}
 

### 2015-08-16 main.tex
comments

@@
 \usepackage{threeparttable}
+\usepackage[colorinlistoftodos]{todonotes} % comments in margins
+\definecolor{flame}{rgb}{0.89, 0.35, 0.13}
+
+\setlength{\marginparwidth}{1.5cm}
+\newcommand{\jri}[1]{\todo[size=\scriptsize, color=flame]{#1}}
+\definecolor{gerkegreen} {rgb} {0,0.6,0}
+\newcommand{\jpg}[1]{\todo[size=\scriptsize, color=gerkegreen]{#1}}
+\newcommand{\rev}[1]{\textcolor{blue}{\bf #1}}
 
@@
 }
-
-%editing commands (please leave in for now)
-\newcommand{\jri}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} }
-\definecolor{gerkegreen} {rgb} {0,0.6,0}
-\newcommand{\jpg}[1]{\textcolor{gerkegreen}{ \em{\scriptsize  #1}} }
-
-\newcommand{\rev}[1]{\textcolor{blue}{\bf #1}}
 
@@
 The observed data is marked by the red line, the simulations by gray dots, and the median of the simulations by a green dot.
-\rev{Black lines represent the 95\% quantiles of the simulated data.} \jri{is that interpretation of the black lines correct?}}
+\rev{Black lines represent the 95\% quantiles of the simulated data.} }
     \label{fig:compare_to_sims}
@@
 Averaged across the genome, the vast majority of the reduction in diversity observed in both populations can be attributed to genetic drift. 
-Nonetheless, we do observe \rev{differences from values generated under our neutral simulations} (Figure \ref{fig:compare_to_sims}). 
+Nonetheless, we do observe \rev{differences from values generated under our neutral simulations} (Figure \ref{fig:compare_to_sims})\jri{is my interpretation of the black lines in figure correct?}. 
 The observed data show higher than expected heterozygosity at cycle 4, \rev{which could be explained by several factors, including differences between simulated and actual breeding practices, under-sampling of diversity in cycle 0, selection acting to increase the frequency of initially rare haplotypes, or even selection acting to maintain heterozygosity \citep[e.g.][]{McMullen2009,Gore2009a} } \jri{where is this excess of diversity?

### 2015-08-16 sup1.tex
...and now it compiles.
figures modified, added to supp.
comments

@@
    \vspace{-0mm}
-   \includegraphics[\linewidth]{fig3}
+   \includegraphics[width=0.7\linewidth]{fig3}
    %\renewcommand{\baselinestretch}{0.9}
@@
 \end{figure*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
+\begin{center}
+\Image[width=0.7\linewidth]{Fig4/chrom1_combined}{} \\
+\Image[width=0.7\linewidth]{Fig4/chrom2_combined}{} 
+\newpage
+\Image[width=0.7\linewidth]{Fig4/chrom3_combined}{} \\
+\Image[width=0.7\linewidth]{Fig4/chrom5_combined}{} 
+\newpage
+\Image[width=0.7\linewidth]{Fig4/chrom6_combined}{} \\
+\Image[width=0.7\linewidth]{Fig4/chrom7_combined}{} 
+\newpage
+\Image[width=0.7\linewidth]{Fig4/chrom8_combined}{} \\
+\Image[width=0.7\linewidth]{Fig4/chrom9_combined}{} 
+\newpage
+\Image[width=0.7\linewidth]{Fig4/chrom10_combined}{}  
+\captionof{figure}{Heterozygosity in each cycle across chromosomes of the BSSS (left) and BSCB1 (right) plotted on the physical (top) and genetic (bottom) map. Details are as in Fig. 4.}
+    \label{fig:others2}
+\end{center}
+
+%\begin{figure*}   
+%  \begin{center}
+%   \vspace{-0mm}
+%   \includegraphics[width=0.7\linewidth]{Fig4/chrom2_combined.pdf}
+%   \includegraphics[width=0.7\linewidth]{Fig4/chrom3_combined.pdf}
+%   \includegraphics[width=0.7\linewidth]{Fig4/chrom5_combined.pdf}
+%   \includegraphics[width=0.7\linewidth]{Fig4/chrom6_combined.pdf}
+%   \includegraphics[width=0.7\linewidth]{Fig4/chrom7_combined.pdf}
+%   \includegraphics[width=0.7\linewidth]{Fig4/chrom8_combined.pdf}
+%   \includegraphics[width=0.7\linewidth]{Fig4/chrom9_combined.pdf}
+%   \includegraphics[width=0.7\linewidth]{Fig4/chrom10_combined.pdf}
+%   %\renewcommand{\baselinestretch}{0.9}
+%   \vspace{-3mm}
+%   \caption{ Heterozygosity in each cycle across chromosomes of the BSSS (left) and BSCB1 (right) plotted on the physical (top) and genetic (bottom) map. Details are as in Fig. 4.
+%} 
+%\vspace{-6mm}
+%    \label{fig:s2}
+%  \end{center}
+%\end{figure*}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
@@
 \begin{table}
-\caption{Heterozygosity of SNPs in cycle 8 but not cycle 4 of the BSSS. \jri{are these numbers supposed to be the same? I checked the Pages doc and they are the same there too.}}
+\caption{Heterozygosity of SNPs in cycle 8 but not cycle 4 of the BSSS. }
 \begin{tabular}{  l | l | l | }
@@
 \end{table}
+\jri{are  numbers in table supposed to be the same? I checked the Pages doc and they are the same there too.}
 

### 2015-08-16 response.tex
some changes, still bib not working

--- 2015-08-14 response.tex
@@
 
-\newcommand{\jri}[1]{\textcolor{red}{ \emph{\scriptsize  #1}} }
+\usepackage[colorinlistoftodos]{todonotes} % comments in margins
+\definecolor{flame}{rgb}{0.89, 0.35, 0.13}
+\setlength{\marginparwidth}{1.5cm}
+\newcommand{\jri}[1]{\todo[size=\scriptsize, color=flame]{#1}}
+\definecolor{gerkegreen} {rgb} {0,0.6,0}
+\newcommand{\jpg}[1]{\todo[size=\scriptsize, color=gerkegreen]{#1}}
+\newcommand{\rev}[1]{\textcolor{blue}{\bf #1}}
 \newcommand{\citex}[1]{\textcolor{red}{\bf #1}}
-\newcommand{\blah}{\textcolor{red}{\bf blah}}
 
@@
 
-\response{PCA of just the founder lines does not show differentiation corresponding to the two populations. \jri{Justin, can you do this quick and dirty to confirm?}}
+\response{Projections of all lines onto a PCA of the founders shows no differentiation corresponding to BSSS and BSCB1. }
 
@@
 As shown in the new figure 3, there are a few regions on chromosome arms that have been fixed and span multiple cM.
-\jri{do we need to note this in the main text?
-is it an intesting observation, maybe implying rapid selection?}}
+\jri{do we need to note in the main text?
+an interesting observation, maybe implying rapid selection?}}
 
@@
 
-\response{\jri{Justin, any ideas? Do we do some formal correspondence with NAM results? Or say ``geez this is hard because QTL are big and we're not sure we'd believe the results?''}}
+\response{\jri{Justin, any ideas? Do we do some formal correspondence with NAM results? }
+\jpg{I'm sure we can find them, due to the size of the regions, but won't tell us much.
+I would be in favor of adding some more to the discussion about how pressure for synchronous flowering time may have driven a lot of what we see.
+But I could dig through Wallace et al.
+to see what is there.
+Thoughts?}}
 
@@
 \response{Four plants were found that appeared noticeably out of place on the PCA.
-For two of these we have insufficent evidence to know whether they were switched, mislabelled, contamination, etc.
+For two of these we have insufficient evidence to know whether they were switched, mislabelled, contamination, etc.
 These were removed from the analysis.

### 2015-08-16 main.tex
some changes, still bib not working

@@
 \definecolor{flame}{rgb}{0.89, 0.35, 0.13}
-
 \setlength{\marginparwidth}{1.5cm}
@@
 Since there was no evidence suggesting when mis-labelings occurred, each of these plants was removed from the analysis.
-\rev{The final genotyping data set is available at \jri{Justin, please send me this and I can put online}.}
+\rev{The final genotyping data set is available at .} \jri{Justin, please send me this and I can put online}
 
@@
 The second axis of variation primarily separates the cycles from one another within each population. 
-There is no separation between the founders of the two populations, but at cycle 0, the BSSS population shows more divergence from the founders than does BSCB1. 
+There is no separation between the founders of the two populations, \rev{and projection of later cycles onto a PCA of the founders shows no distinction between BSSS and BSCB1.
+At cycle 0, however,} the BSSS population shows more divergence from the founders than does BSCB1. 
 This could be due to drift during either the population’s construction or subsequent maintenance. 
@@
 
-\bibliography{references.bib}
+\bibliography{references}
 \bibliographystyle{geneticsT2}

### 2015-08-16 main.tex
refs working

@@
 \usepackage{threeparttable}
+
 \usepackage[colorinlistoftodos]{todonotes} % comments in margins
@@
 \subsection*{Diversity and principal component analysis}
-Heterozygosity (H) was measured as:
-
-\begin{align*}
-H = 2p(1-p)
-\end{align*}
-
-where $p$ and $(1-p)$ are the frequencies of the two SNP alleles. 
+Heterozygosity (H) was measured as $H = 2p(1-p)$, where $p$ and $(1-p)$ are the frequencies of the two SNP alleles. 
 $F_{ST}$ \citep{hudson1992statistical} was calculated using the HBKpermute program in the analysis package (\url{https://github.com/molpopgen/analysis}) of the software library libsequence \citep{thornton2003libsequence}.
@@
 The second axis of variation primarily separates the cycles from one another within each population. 
-There is no separation between the founders of the two populations, \rev{and projection of later cycles onto a PCA of the founders shows no distinction between BSSS and BSCB1.
+There is no separation between the founders of the two populations, \rev{and projection of later cycles onto a PCA of the founders shows no distinction between BSSS and BSCB1 (Figure \ref{fig:sfounders}).
 At cycle 0, however,} the BSSS population shows more divergence from the founders than does BSCB1. 
@@
 
-\bibliography{references}
 \bibliographystyle{geneticsT2}
+\bibliography{references.bib}
 

### 2015-08-19 main.tex
revisions to intro from Jode/Justin

--- 2015-08-16 main.tex
@@
 
-Multiple studies with molecular markers have indicated that these heterotic groups have diverged genetically over time to become highly structured and isolated populations, resulting in a dramatic restructuring of population genetic variation \citep{duvick2004long, ho2005extent, feng2006temporal}.
+\rev{Although the founders of these heterotic groups were not initially differentiated, m}ultiple studies with molecular markers have indicated that these heterotic groups have diverged genetically over time to become highly structured and isolated populations, resulting in a dramatic restructuring of population genetic variation \citep{duvick2004long, ho2005extent, feng2006temporal}.
 Advances in high throughput genotyping and the development of a maize reference genome now enable the observation of maize population structure at high marker density across the whole genome \citep{ganal2011a-large,chia2012maize}.
@@
 Although selective sweeps remaining from domestication are clearly visible, the impact of selection during modern breeding appears comparatively small in terms of its impact on genomic diversity \citep{Hufford2012b,van2012historical}, despite steady, heritable improvement in phenotype \citep{duvick2005contribution}.
-The lack of distinct selection signals from modern breeding may be due to population-specific selection; this interpretation is supported by the measured success in identifying targets of selection in individual experimental populations selected for a limited number of phenotypes \citep{hirsch2014insights, beissinger2014genome}.
+The lack of distinct selection signals \rev{in broad germplasm collections may be due to a population specific selection within germplasm subgroups, which would be further obscured by migration due to the exchange of material among breeders.} \textcolor{red}{cite Mikel here}  \rev{In this case selection will be easier to detect in populations maintained in an individual program.} This \rev{possibliity} is supported by the measured success in identifying targets of selection in individual experimental populations \rev{under directional selection for specific phenotypes such as seed size and ear number.} \citep{hirsch2014insights, beissinger2014genome}.
 
-Here we take a detailed look at the dynamics of genome-wide patterns of diversity over time within an individual breeding program. 
-The Iowa Stiff Stalk Synthetic (BSSS) and the Iowa Corn Borer Synthetic No.
-1 (BSCB1) Reciprocal Recurrent Selection Program of the USDA-ARS at Ames, Iowa (hereafter referred to as the Iowa RRS) represents one of the best-documented public experiments on selection for combining ability and hybrid performance.
-BSSS and BSCB1 have been recurrently selected for improved cross-population hybrids \citep{penny1971twenty}. 
-\jri{Jode/Justin add text explaining recurrent selection and populations}
-This model of selection, named reciprocal recurrent selection, provides the generalized model for strategies commonly used in commercial maize breeding \citep{comstock1949breeding, duvick2004long}. 
-The Iowa RRS experiment proves especially relevant because lines derived from the BSSS population have had a major impact upon the development of commercial hybrids \citep{duvick2004long, darrah1986}, the formation of modern heterotic groups \citep{troyer1999background, senior1998utility}, and the choice of a maize reference genome \citep{schnable2009the-b73-maize}.
-\jri{something about heterosis}
+\rev{In this study, we examine a different closed experiment whose phenotypic target and selection method -Reciprocal Recurrent Selection- closely mirrors methods used in the generation of modern maize hybrids \citep{comstock1949breeding, duvick2004long}.
+Proposed by Comstock et al.
+(1949), reciprocal recurrent selection is a procedure to simultaneously improve two populations for their ability to produce high yielding hybrids when individuals from the two populations are crossed.
+In this method, the two populations are kept separate and lines from each population are advanced to the next cycle of selection based on the performance of the line or its close relatives in a testcross with random individuals from the opposing population.
+Selection is therefore carried out on lines within each population, but on a cross-population hybrid phenotype.
+After selection the selected lines within each population are intermated to generate lines for the next cycle.
+In this way the two populations simultaneously evolve improved combining ability with one another.
+The USDA-ARS at Ames, Iowa has conducted a Reciprocal Recurrent Selection Program (hereafter referred to as the Iowa RRS) with the Iowa Stiff Stalk Synthetic (BSSS) and Iowa Corn Borer Synthetic No.
+1 (BSCB1) populations for 18 cycles of selection\citep{penny1971twenty}} \textcolor{red}{cite Jode 2011 here}.
+\rev{This program represents one of the best-documented public experiments on selection for combining ability and hybrid performance.
+Reminiscent of modern maize germplasm as a whole, BSSS and BSCB1 have evolved from undifferentiated founders as they have been selected for cross-population hybrid yield.
+BSSS and BSCB1 have both been maintained as closed populations with no intentional migration (although evidence suggests low level of unintentional gene flow early in the BSSS population).
+Extensive records on population sizes, breeding methods, selection differentials, and seed are available for all cycles of selection.
+This set of resources on a closed population selected for hybrid yield make the Iowa RRS an ideal test case for the study of the genomic impact of intensive hybrid breeding in maize.} The Iowa RRS experiment \rev{provides additional historical relevance} because lines derived from the BSSS population have had a major impact upon the development of commercial hybrids \citep{duvick2004long, darrah1986}, the formation of modern heterotic groups \citep{troyer1999background, senior1998utility}, and the choice of a maize reference genome \citep{schnable2009the-b73-maize}.
+\jri{something about heterosis} \jpg{what are you thinking for heterosis here?}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRO

### 2015-08-19 response.tex
refs working
jode text

@@
 
-\response{Projections of all lines onto a PCA of the founders shows no differentiation corresponding to BSSS and BSCB1. }
+\response{Projections of all lines onto a PCA of the founders shows no differentiation corresponding to BSSS and BSCB1.
+The main eigenvector only explains $\approx 6\%$ of the variation in the founders and appears to separate Reid and Lancaster lines.
+The BSSS and BSCB1 plants form two clusters near zero along this axis.
+This figure has been included in the supplement.
+}
 
@@
 
-\response{\jri{Justin? Jode?}}
+\response{We have added a paragraph in the introduction explaining the RRS method and populations in increased detail.\jri{this enough or do we add more elsewhere?}}
 
@@
 
-\response{\jri{Jode? Justin?}}
+\response{We have noted in the new introductory paragraph that yield was the main focus of selection.\jri{add other phenotypes elsewhere?}}
 

### 2015-08-19 main.tex
jode text

--- 2015-08-16 main.tex
@@
 
-Multiple studies with molecular markers have indicated that these heterotic groups have diverged genetically over time to become highly structured and isolated populations, resulting in a dramatic restructuring of population genetic variation \citep{duvick2004long, ho2005extent, feng2006temporal}.
+\rev{While the founders of these heterotic groups were not initially differentiated, m}ultiple studies with molecular markers have indicated that these heterotic groups have diverged genetically over time to become highly structured and isolated populations, resulting in a dramatic restructuring of population genetic variation \citep{duvick2004long, ho2005extent, feng2006temporal}.
 Advances in high throughput genotyping and the development of a maize reference genome now enable the observation of maize population structure at high marker density across the whole genome \citep{ganal2011a-large,chia2012maize}.
@@
 Although selective sweeps remaining from domestication are clearly visible, the impact of selection during modern breeding appears comparatively small in terms of its impact on genomic diversity \citep{Hufford2012b,van2012historical}, despite steady, heritable improvement in phenotype \citep{duvick2005contribution}.
-The lack of distinct selection signals from modern breeding may be due to population-specific selection; this interpretation is supported by the measured success in identifying targets of selection in individual experimental populations selected for a limited number of phenotypes \citep{hirsch2014insights, beissinger2014genome}.
+The lack of distinct selection signals \rev{in broad germplasm collections may be due to population-specific selection within germplasm subgroups; in this case selection should be easier to detect in populations maintained in an individual program.} This \rev{possibliity} is supported by the measured success in identifying targets of selection in individual experimental populations \rev{under directional selection for specific phenotypes such as seed size and ear number} \citep{hirsch2014insights, beissinger2014genome}.
 
-Here we take a detailed look at the dynamics of genome-wide patterns of diversity over time within an individual breeding program. 
-The Iowa Stiff Stalk Synthetic (BSSS) and the Iowa Corn Borer Synthetic No.
-1 (BSCB1) Reciprocal Recurrent Selection Program of the USDA-ARS at Ames, Iowa (hereafter referred to as the Iowa RRS) represents one of the best-documented public experiments on selection for combining ability and hybrid performance.
-BSSS and BSCB1 have been recurrently selected for improved cross-population hybrids \citep{penny1971twenty}. 
-\jri{Jode/Justin add text explaining recurrent selection and populations}
-This model of selection, named reciprocal recurrent selection, provides the generalized model for strategies commonly used in commercial maize breeding \citep{comstock1949breeding, duvick2004long}. 
-The Iowa RRS experiment proves especially relevant because lines derived from the BSSS population have had a major impact upon the development of commercial hybrids \citep{duvick2004long, darrah1986}, the formation of modern heterotic groups \citep{troyer1999background, senior1998utility}, and the choice of a maize reference genome \citep{schnable2009the-b73-maize}.
-\jri{something about heterosis}
+\rev{In this study, we examine a different closed experiment whose phenotypic target and selection method --- reciprocal recurrent aelection --- closely mirrors methods used in the generation of modern maize hybrids \citep{comstock1949breeding, duvick2004long}.
+Proposed by \citet{comstock1949breeding}, reciprocal recurrent selection is a procedure to simultaneously improve two populations for their ability to produce high yielding hybrids when individuals from the two populations are crossed.
+In this method, the two populations are kept separate and lines from each population are selected based on the phenotype of their cross-population hybrids. 
+Selected lines within each population are then intermated to generate lines for the next cycle. 
+In this way, the two populations simultaneously evolve improved combining ability with one another. 
+The USDA-ARS at Ames, Iowa has conducted a reciprocal recurrent selection program (hereafter referred to as the Iowa RRS) with the Iowa Stiff Stalk Synthetic (BSSS) and Iowa Corn Borer Synthetic No.
+1 (BSCB1) populations for 18 cycles of selection\citep{penny1971twenty}.
+This program represents one of the best-documented public experiments on selection for combining ability and hybrid performance. 
+Extensive records on population sizes, breeding methods, selection differentials, and seed are available for all cycles of selection. 
+This set of resources on a closed population selected for hybrid yield make the Iowa RRS an ideal test case for the study of the genomic impact of intensive hybrid breeding in maize.} 
+The Iowa RRS experiment \rev{provides additional historical relevance} because lines derived from the BSSS population have had a major impact on the development of commercial hybrids \citep{duvick2004long, darrah1986}, the formation of modern heterotic groups \citep{troyer1999background, senior1998utility}, and the choice of a maize reference genome \citep{schnable2009the-b73-maize}.
+\jri{something about heterosis} \jpg{what are you thinking for heterosis here?}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRO

### 2015-08-19 main.tex
methods

@@
 The USDA-ARS at Ames, Iowa has conducted a reciprocal recurrent selection program (hereafter referred to as the Iowa RRS) with the Iowa Stiff Stalk Synthetic (BSSS) and Iowa Corn Borer Synthetic No.
-1 (BSCB1) populations for 18 cycles of selection\citep{penny1971twenty}.
+1 (BSCB1) populations for 18 cycles of selection \citep{penny1971twenty}.
 This program represents one of the best-documented public experiments on selection for combining ability and hybrid performance. 
@@
 \subsection*{The BSSS and BSCB1 recurrent selection program}
-The Iowa RRS experiment began with founder inbred lines (Table S1) that were randomly mated to create the BSSS and BSCB1 ‘cycle 0’ base populations. 
-Each group was then recurrently selected for improved combining ability with the other group \citep{penny1971twenty, keeratinijakal1993responses}. 
-Initially approximately 200-250 starting ‘S0’ plants within each population were simultaneously self-fertilized to generate ‘S1’ lines and crossed to a random sample (4-6) of plants from the other population to generate testcross seed.
-The material was evaluated for highly heritable traits including standability, disease resistance and ear rot resistance in the nursery at the time the testcross seed was made, and 100 testcrosses were then grown in replicated field trials and evaluated for grain yield, grain moisture, and standability.
-S1 seed of 10 selected families within each population were randomly mated to create a new ‘cycle n+1’ population. 
-After cycle 4, the testcrosses were carried out on the S1 lines themselves rather than at S0, which leads to another round of selfing prior to the creation of the next cycle.
-Ten lines (out of 100) were selected and advanced to form the next cycle between cycles 1 and 8, and twenty lines were selected between cycles 9 and 16 \citep{keeratinijakal1993responses}.
-Founder inbreds and samples from cycles 0, 4, 8, 12, and 16 were genotyped at 39,258 SNPs that passed a set of quality filters and could be assigned collinear genetic and physical map positions.
+\rev{The Iowa RRS program was initiated in 1949 \citep{penny1971twenty}. 
+The BSSS population was formed from 16 inbred lines selected for stalk strength in 1933 and 1934 (Sprague, 1946) and the BSCB1 population was formed from a set of 12 inbred lines \citep{hallauer1974registration}.}
+Founder lines (Table S1) from each population were randomly mated to create the `cycle 0' base population.
+\rev{Testcross progenies were formed by self-pollinating 100 individuals in each population and simultaneously crossing the same 100 individuals onto 10 plants (as females) in the reciprocal population (e.g., a BSSS plant was self-pollinated and crossed to 10 females in BSCB1).
+Seed from the 10 females pollinated by a single male plant was bulked, and the 200 testcross progenies were grown in replicated multi-environment yield trials.  
+Ten testcross families were selected from among the BSSS males (crossed to BSCB1 as a tester) and BSCB1 males (crossed to BSSS as a tester) based primarily on grain yield.  
+Self-pollinated seed from the 10 selected males in each population were planted in the following season and plant-to-plant crosses made between each of the 45 possible pairs; seed from these crosses formed the cycle 1 populations.}
+
+\rev{Additional changes to this procedure occurred in later cycles \citep{penny1971twenty,keeratinijakal1993responses}.
+In cycles six through eight, individual S1 plants (progeny of one generation of self-pollination) were used as males instead of noninbred plants.  
+Beginning in cycle eight, the number of testcross progenies selected to form the next cycle of selection was doubled to 20.  
+Beginning in cycle 10, the method of producing testcross progenies for evaluation was changed to generate a single set of reciprocal full sib families between the populations rather than two sets of half-sib families.
+\jri{i have commented out a couple lines i think are too much detail.
+uncomment them if you think they are worth including.} %Specifically, progeny rows resulting from self-pollination of individuals within each population were planted in a breeding nursery.
+%An individual S1 plant in BSSS was self-pollinated and crossed to four females in an S1-row of BSCB1 while a fifth plant in a BSCB1 row was also self-pollinated and crossed to four females in the S1-row of BSSS, resulting in reciprocal crossing of a BSSS S1-row and BSCB1 S1-row.
+%Seed of eight females was bulked to form a single reciprocal hybrid.  
+This procedure has been continued to the present, with reciprocal hybrid testcross progeny derived from the cycle 18 population in 2014 and evaluated in 2015 for formation of cycle 19.  
+The phenotypes evaluated for selection have been consistent across all 19 cycles of selection: grain yield (dry matter yield of maize grain per acre) has been the most important, but selection was also applied for reduced moisture content at harvest and reduced root and stalk lodging.
+A formal selection index was implemented beginning with cycle nine.  
+Plant densities have increased consistently across cycles of selection as the populations have become consistently better adapted to high plant density with continued selection \citep{brekke2011selection}.}
 
@@
 However, because our simulations are based not on theoretical expectations but instead on sampling from the observed data at cycle 0, we expect ascertainment bias to have a minimal impact on our results.
-The effect of this ascertainment bias was shown to be minimal in another study of North American germplasm \citep{van2012historical}.
+%The effect of this ascertainment bias was shown to be minimal in another study of North American germplasm \citep{van2012historical}.
 

### 2015-08-19 response.tex
response edit
response:

@@
 
-\response{\jri{Justin, any ideas? Do we do some formal correspondence with NAM results? }
-\jpg{I'm sure we can find them, due to the size of the regions, but won't tell us much.
-I would be in favor of adding some more to the discussion about how pressure for synchronous flowering time may have driven a lot of what we see.
-But I could dig through Wallace et al.
-to see what is there.
-Thoughts?}}
+\response{We agree these regions are of interest, but the large size of these regions is such that there is almost certainly overlap with flowering time or resistance QTL.
+We would be hesitant to draw any conclusions from such overlap as similar overlap is likely to occur for a number of other traits of less interest, and evaluating significance or enrichment would be difficult.
+We have generally avoided GO or other enrichment analyses as the aim of the paper was not in identifying specific candidate genes or characterizing their function.}
 
@@
 
-\response{We have added a paragraph in the introduction explaining the RRS method and populations in increased detail.\jri{this enough or do we add more elsewhere?}}
+\response{We have mentioned the RRS method and populations a bit more in the introduction and included new paragraphs in the methods explaining the RRS method and populations in increased detail.}
 
@@
 
-\response{We have noted in the new introductory paragraph that yield was the main focus of selection.\jri{add other phenotypes elsewhere?}}
+\response{We have noted in the new paragraphs in the introduction and methods that yield was the main target of selection but that other traits (moisture content at harvest and reduced root and stalk lodging) were also selected upon.}
 
@@
 as "H" with and without italics and also as heterozygosity spelled out, sometimes within the same figure or 
-figure caption. Simiarily, Fst needs to be presented in a consistent manner.
-
-\response{We have changed text and labels to improve consistency in using heterozygosity vs ``H''.}
-
+figure caption. Similarily, Fst needs to be presented in a consistent manner. 
 Also edit citations for consistent capitalization and in=text citations for consistent spacing.
 
-\response{We have moved the ms from word to latex; we believe this has fixed many of these issues. We have also corrected bibtex entries in several places to improve consistency of capitalization.}
+\response{We have moved the ms from word to latex; we believe this has fixed many of these issues.
+We have remade figures and hopefully are now consistent in use of Fst and H.
+We have also corrected bibtex entries in several places to improve consistency of capitalization.}
 

### 2015-08-19 main.tex
typos, thank you kate

@@
 We then take advantage of our high-density genotype data to identify extensive regions of haplotype fixation and trace haplotype ancestry to single founder inbred lines.
-The vast majority of regions showing such evidence of selection differe between the two populations, providing evidence against the overdominance model as a means of explaining selection for hbyrid vigor.
+The vast majority of regions showing such evidence of selection differ between the two populations, providing evidence against the overdominance model as a means of explaining selection for hybrid vigor.
 }
@@
 Classic overdominance models of heterosis predict that at a single locus, two distinct alleles confer heterozygote advantage when combined, while the dominance model predicts that heterosis is driven by dominance effects and the complementation of linked alleles in low-recombination regions (dominance or pseudo-overdominance).
-In the case of true overdominance, we selection should lead to decreased heterozygosity at a locus in both populations as complementary haplotypes are fixed in each group \rev{\citep[e.g.][]{guo2014maize}}.
+In the case of true overdominance,  selection should lead to decreased heterozygosity at a locus in both populations as complementary haplotypes are fixed in each group \rev{\citep[e.g.][]{guo2014maize}}.
 We find \rev{little evidence to support this genetic phenomenon, finding only two 2 cM windows genome-wide in which both populations show significantly reduced heterozygosity.} 

### 2015-08-20 main.tex
higlighted text with number of SNPs in mismapped blocks
changed final SNP number by two. The last individual thrown out must have changed the filter

@@
 Since there was no evidence suggesting when mis-labelings occurred, each of these plants was removed from the analysis.
-\rev{The final genotyping data set is available at .} \jri{Justin, please send me this and I can put online}
+\rev{The final genotyping data set contained 39,261 SNPs and is available at .} \jri{Justin, please send me this and I can put online}
 
@@
 The last SNP of the block was also given a 10 kb cushion between itself and the next SNP on the correct linkage group. 
-Among the SNPs used for analyses, 15 were mis-mapped to a different linkage group 1,585 were moved within a linkage group.
+\rev Among the SNPs used for analyses, 15 were mis-mapped to a different linkage group and 1,585 were moved within a linkage group.
 
@@
 \subsection*{Population structure and genetic diversity}
-Founder inbreds and samples from cycles 0, 4, 8, 12, and 16 were genotyped at 39,258 SNPs that passed a set of quality filters and could be assigned collinear genetic and physical map positions (See Materials and Methods for details).
+Founder inbreds and samples from cycles 0, 4, 8, 12, and 16 were genotyped at 39,261 SNPs that passed a set of quality filters and could be assigned collinear genetic and physical map positions (See Materials and Methods for details).
 Change in population structure throughout the Iowa RRS experiment can be observed visually by a principal component analysis (PCA). 

### 2015-08-20 datatoshare/writeout.R
wrote the dataset to csv

@@ -1,0 +1,21 @@
+rm(list = ls())
+
+load("../data/new_full_set.RData")
+
+sample_info <- data.frame(Name = new_full_set$lines,
+                          Population = new_full_set$group,
+                          Type = new_full_set$status,
+                          Cycle = new_full_set$cycle)
+
+sample_info$Type <- as.character(sample_info$Type)
+sample_info$Type[sample_info$Type == "Outbred"] = "Population Sample"
+
+markerinfo <- data.frame(Name = new_full_set$markernames,
+                         ID = new_full_set$markercodes,
+                         Chromosome = new_full_set$LG,
+                         Position = new_full_set$position,
+                         Genetic_Position = new_full_set$F2map)
+
+write.csv(sample_info, file = "sample_info.csv", quote = F, row.names = F)
+write.csv(markerinfo, file = "SNP_info.csv", quote = F, row.names = F)
+write.table(new_full_set$calls, file = "genotype_data.csv", quote = F, sep = ',')

### 2015-08-26 main.tex
editorial changes to methods, especially phys vs. genetic map

@@
 
-\rev{In this study, we examine a different closed experiment whose phenotypic target and selection method --- reciprocal recurrent aelection --- closely mirrors methods used in the generation of modern maize hybrids \citep{comstock1949breeding, duvick2004long}.
+\rev{In this study, we examine a different closed experiment whose phenotypic target and selection method --- reciprocal recurrent selection --- closely mirrors methods used in the generation of modern maize hybrids \citep{comstock1949breeding, duvick2004long}.
 Proposed by \citet{comstock1949breeding}, reciprocal recurrent selection is a procedure to simultaneously improve two populations for their ability to produce high yielding hybrids when individuals from the two populations are crossed.
@@
 Beginning in cycle 10, the method of producing testcross progenies for evaluation was changed to generate a single set of reciprocal full sib families between the populations rather than two sets of half-sib families.
-\jri{i have commented out a couple lines i think are too much detail.
-uncomment them if you think they are worth including.} %Specifically, progeny rows resulting from self-pollination of individuals within each population were planted in a breeding nursery.
+%Specifically, progeny rows resulting from self-pollination of individuals within each population were planted in a breeding nursery.
 %An individual S1 plant in BSSS was self-pollinated and crossed to four females in an S1-row of BSCB1 while a fifth plant in a BSCB1 row was also self-pollinated and crossed to four females in the S1-row of BSSS, resulting in reciprocal crossing of a BSSS S1-row and BSCB1 S1-row.
@@
 The phenotypes evaluated for selection have been consistent across all 19 cycles of selection: grain yield (dry matter yield of maize grain per acre) has been the most important, but selection was also applied for reduced moisture content at harvest and reduced root and stalk lodging.
-A formal selection index was implemented beginning with cycle nine.  
+%A formal selection index was implemented beginning with cycle nine.  
 Plant densities have increased consistently across cycles of selection as the populations have become consistently better adapted to high plant density with continued selection \citep{brekke2011selection}.}
@@
 Genotypes were determined with the GenomeStudio v2010.2 software using the manufacturer's MaizeSNP50\_B.egt cluster file. 
-\jri{this is an edit because url listed in pages doc no longer valid. correct that this is not a custom .egt file?}
+%\jri{this is an edit because url listed in pages doc no longer valid. correct that this is not a custom .egt file?}
 The design of the maize SNP50 Chip included a relatively small ascertainment panel of inbred lines, introducing a bias in the frequencies of SNPs included on the chip \citep{ganal2011a-large}. 
@@
 \subsection*{Integrating the genetic and physical map}
-The interpretation of our results depends upon a genetic and physical map that is as accurate as possible. 
+%The interpretation of our results depends upon a genetic and physical map that is as accurate as possible. 
+\rev Our simulation approach requires co-linear genetic and physical maps.
 We therefore took steps to improve the positions of the SNP markers on the genetic and physical maps relative to version 5A.59 of the maize genome assembly.
@@
 The last SNP of the block was also given a 10 kb cushion between itself and the next SNP on the correct linkage group. 
-\rev Among the SNPs used for analyses, 15 were mis-mapped to a different linkage group and 1,585 were moved within a linkage group.
-
-Only a portion of the SNPs on the array had genetically mapped positions. 
-‘Unmapped’ SNPs (those with a physical position but no genetic position) bounded by two mapped markers were moved along with their mapped neighbors if both anchoring mapped markers were also moved. 
-However, it is unclear whether unmapped SNPs just outside of these anchors should be kept in place or moved along with the adjacent SNPs. 
+\rev Non-framework SNPs which had a physical position but no genetic coordinates were moved along with their framework neighbors if the nearest flanking framework markers were also moved.
+However, it is unclear whether non-framework SNPs just outside of these anchors should be kept in place or moved along with the adjacent SNPs. 
 Since most inversions were small relative to the genetic map (and would therefore still fall in the same window of a sliding window analysis), these SNPs were left in place. 
 However, markers bordering translocations were removed to ensure there were no markers mapped to the incorrect linkage group. 
-Unmapped SNPs were then assigned a genetic position by linear interpolation of genetic vs. physical distance using the approx() function in R \citep{rteam}. 
+\rev Among the SNPs used for analyses, 15 were mis-mapped to a different linkage group and 1,585 were moved within a linkage group.
+%Only a portion of the SNPs on the array had genetically mapped positions.
+
+\rev Approximate genetic positions for non-framework SNPs were interpolated based on their physical positions with the approx() function in R \citep{rteam} with the 4,217 framework SNPs used as a reference.
+%Unmapped SNPs were then assigned a genetic position by linear interpolation of genetic vs. physical distance using the approx() function in R \citep{rteam}. 
 The IBM genetic map distances were then converted to single-meiosis map distances using the formulae of \citet{winkler2003determination}. 

### 2015-08-26 main.tex
editorial changes to 'integrating genetic and physical map'

@@
 %The interpretation of our results depends upon a genetic and physical map that is as accurate as possible. 
-\rev Our simulation approach requires co-linear genetic and physical maps.
+\rev{Our simulation approach requires co-linear genetic and physical maps.}
 We therefore took steps to improve the positions of the SNP markers on the genetic and physical maps relative to version 5A.59 of the maize genome assembly. 
@@
 However, markers bordering translocations were removed to ensure there were no markers mapped to the incorrect linkage group. 
-\rev Among the SNPs used for analyses, 15 were mis-mapped to a different linkage group and 1,585 were moved within a linkage group.
+\rev{Among the SNPs used for analyses, 15 were mis-mapped to a different linkage group and 1,585 were moved within a linkage group.}
 %Only a portion of the SNPs on the array had genetically mapped positions.
 
-\rev Approximate genetic positions for non-framework SNPs were interpolated based on their physical positions with the approx() function in R \citep{rteam} with the 4,217 framework SNPs used as a reference.
+\rev{Approximate genetic positions for non-framework SNPs were interpolated based on their physical positions with the approx() function in R \citep{rteam} with the 4,217 framework SNPs used as a reference.}
 %Unmapped SNPs were then assigned a genetic position by linear interpolation of genetic vs. physical distance using the approx() function in R \citep{rteam}. 

### 2015-08-26 main.tex
updated genetic position numbers in the text.

@@
 This F1 was presented to fastPHASE with the rest of the data, except that its parent inbreds were removed. 
-Analyses of several hypothetical F1’s from different cycles of selection revealed very low phasing error rates (Table S3). 
+Analyses of several hypothetical F1’s from different cycles of selection revealed very low phasing error rates (Table S3).  \jpg{S2 now?}
 Therefore the phased genotypes of cycle 0 plants were used as the starting data for simulations (see below).
@@
 Consistent with previous studies of the Iowa RRS \citep{messmer1991genetic, labate1997molecular,  hagdorn2003molecular, hinze2005population}, genome-wide genetic diversity (expected heterozygosity, H) decreases steadily across cycles of selection in both populations (Figure \ref{fig:decline}).
-\jri{do all these citations make sense here?
+%\jri{do all these citations make sense here?
 I just copied these from the abstract}
@@
 
-We noticed an irregular increase in the number of polymorphic markers between BSSS cycles 4 and 8 (Table \ref{tab:s2}). 
+We noticed an irregular increase in the number of polymorphic markers between BSSS cycles 4 and 8 (Table \ref{tab:s2}). \jpg{S3 now?}
 All of these newly polymorphic markers were present at extremely low frequency and were spread among various individuals. 
@@
 These regions experience low rates of meiotic recombination, which creates an expanded physical map relative to their genetic length \citep{ganal2011a-large}. 
-In general, the majority of fixed haplotype segments are small ($<2$ cM) in genetic space regardless of their physical size; one exception is 7 cM region on chromosome 1 in the BSCB1 population.
+In general, the majority of fixed haplotype segments are small ($<2$ cM) in genetic space regardless of their physical size; one exception is an 8 cM region on chromosome 1 in the BSCB1 population.
 
@@
 In most cases, these fixed haplotypes can be traced back to single founder inbreds. 
-For example, in BSSS, a 60 Mb (1.2 cM) region of chromosome 9 became fixed by cycle 12 and traces back to the founder Os420, and in BSCB1, a 60 Mb (0.7 cM) region became fixed on chromosome 4 and traces back to A340 (Figure \ref{fig:genphys}).
+For example, in BSSS, a 60 Mb (~2 cM) region of chromosome 9 became fixed by cycle 12 and traces back to the founder Os420, and in BSCB1, a 60 Mb (~3 cM) region became fixed on chromosome 4 and traces back to A340 (Figure \ref{fig:genphys}).
 Table \ref{tab:fix} gives a summary of the large genomic regions that have become fixed or nearly fixed by cycle 16. 

### 2015-08-26 main.tex
updated Figure4 caption to reflect only shading at P<0.001

@@
 Each data point is color-coded based on a linear transformation of recombination rate (red = low recombination rate).
-Shaded regions represent 2 cM windows with heterozygosity values significantly lower than expected by simulation at a given cycle.
-\rev{Light and darker shading denotes significance at $P<0.0025$ and $P<0.001$, respectively.
-Plots for other chromosomes are shown in in Figure \ref{fig:others2}}.
+\rev{Shaded regions represent 2 cM windows with heterozygosity values significantly lower than expected by simulation at a given cycle at $P<0.001$.} Plots for other chromosomes are shown in in Figure \ref{fig:others2}}.
 Comparing the physical and genetic maps reveals that the regions of fixation are extremely small genetically, but some contain much of the physical content of the chromosome.

### 2015-08-26 main.tex
Fig 5 caption now states 99% quantile

@@
 The observed data is marked by the red line, the simulations by gray dots, and the median of the simulations by a green dot.
-\rev{Black lines represent the 95\% quantiles of the simulated data.} }
+\rev{Black lines represent the 99\% and 1\% quantiles of the simulated data.} }
     \label{fig:compare_to_sims}
@@
 Averaged across the genome, the vast majority of the reduction in diversity observed in both populations can be attributed to genetic drift. 
-Nonetheless, we do observe \rev{differences from values generated under our neutral simulations} (Figure \ref{fig:compare_to_sims})\jri{is my interpretation of the black lines in figure correct?}. 
+Nonetheless, we do observe \rev{differences from values generated under our neutral simulations} (Figure \ref{fig:compare_to_sims})
 The observed data show higher than expected heterozygosity at cycle 4, \rev{which could be explained by several factors, including differences between simulated and actual breeding practices, under-sampling of diversity in cycle 0, selection acting to increase the frequency of initially rare haplotypes, or even selection acting to maintain heterozygosity \citep[e.g.][]{McMullen2009,Gore2009a} } \jri{where is this excess of diversity?

### 2015-08-26 main.tex
updated another cM value in the text

@@
 Simulation of the haplotype as a single locus in the RRS experiment produces this increase in frequency in only 3.9\% of 1000 independent simulations, whereas the haplotype was lost in over 80\% of the simulations.
-In BSCB1, a 30 Mb, 1.6 cM region of chromosome 2 became nearly fixed by cycle 8 (67/70) despite a prevalence of 4/72 at cycle 0, which occurred 1.5\% of the time by simulation. 
+In BSCB1, a 30 Mb, a $<1$cM region of chromosome 2 became nearly fixed by cycle 8 (67/70) despite a prevalence of 4/72 at cycle 0, which occurred 1.5\% of the time by simulation. 
 Although these results suggest that selection may have pushed these haplotypes to fixation, the fact that fixation of such a rare haplotype still occurred in some simulations speaks to the strong genetic drift imposed upon the BSSS and BSCB1 populations.

### 2015-08-26 main.tex
discussion point about BSSS haps is correct

@@
 B37 and B14 were derived from cycle 0, and B73 was derived from a half-sib recurrent selection program also started with the BSSS population. 
-We examined these three inbreds at the pericentromeric regions listed in Table \ref{tab:fix} \jri{correct?}, and found that in most cases they carry different haplotypes from those that rose to high frequency in the RRS experiment.
+We examined these three inbreds at the pericentromeric regions listed in Table \ref{tab:fix}, and found that in most cases they carry different haplotypes from those that rose to high frequency in the RRS experiment.
 Near isogenic lines that substitute these centromeric haplotypes from key inbreds could be used to determine if the haplotypes at high frequency in cycle 16 of BSSS provide a phenotypic advantage when crossed with BSCB1.

### 2015-08-26 main.tex
fixed highlighting

@@
 The last SNP of the block was also given a 10 kb cushion between itself and the next SNP on the correct linkage group. 
-\rev Non-framework SNPs which had a physical position but no genetic coordinates were moved along with their framework neighbors if the nearest flanking framework markers were also moved.
-However, it is unclear whether non-framework SNPs just outside of these anchors should be kept in place or moved along with the adjacent SNPs. 
+\rev{Non-framework SNPs which had a physical position but no genetic coordinates were moved along with their framework neighbors if the nearest flanking framework markers were also moved.
+However, it is unclear whether non-framework SNPs just outside of these anchors should be kept in place or moved along with the adjacent SNPs.} 
 Since most inversions were small relative to the genetic map (and would therefore still fall in the same window of a sliding window analysis), these SNPs were left in place. 

### 2015-08-26 main.tex
minor edits with justin

@@
 The Iowa RRS experiment \rev{provides additional historical relevance} because lines derived from the BSSS population have had a major impact on the development of commercial hybrids \citep{duvick2004long, darrah1986}, the formation of modern heterotic groups \citep{troyer1999background, senior1998utility}, and the choice of a maize reference genome \citep{schnable2009the-b73-maize}.
-\jri{something about heterosis} \jpg{what are you thinking for heterosis here?}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRO
@@
 The BSSS population was formed from 16 inbred lines selected for stalk strength in 1933 and 1934 (Sprague, 1946) and the BSCB1 population was formed from a set of 12 inbred lines \citep{hallauer1974registration}.}
-Founder lines (Table S1) from each population were randomly mated to create the `cycle 0' base population.
+Founder lines (Table \ref{tab:founders}) from each population were randomly mated to create the `cycle 0' base population.
 \rev{Testcross progenies were formed by self-pollinating 100 individuals in each population and simultaneously crossing the same 100 individuals onto 10 plants (as females) in the reciprocal population (e.g., a BSSS plant was self-pollinated and crossed to 10 females in BSCB1).
@@
 \subsection*{Plants and inbred lines used}
-The plants and inbred lines used in this experiment are listed in Table S1. 
+The plants and inbred lines used in this experiment are listed in Table \ref{tab:founders}. 
 We genotyped 36 plants from each of the BSSS and BSCB1 populations at selection cycles 0, 4, 8, 12, and 16. 
@@
 Genotypes were determined with the GenomeStudio v2010.2 software using the manufacturer's MaizeSNP50\_B.egt cluster file. 
-%\jri{this is an edit because url listed in pages doc no longer valid. correct that this is not a custom .egt file?}
 The design of the maize SNP50 Chip included a relatively small ascertainment panel of inbred lines, introducing a bias in the frequencies of SNPs included on the chip \citep{ganal2011a-large}. 
 However, because our simulations are based not on theoretical expectations but instead on sampling from the observed data at cycle 0, we expect ascertainment bias to have a minimal impact on our results.
-%The effect of this ascertainment bias was shown to be minimal in another study of North American germplasm \citep{van2012historical}.
 
@@
 Since there was no evidence suggesting when mis-labelings occurred, each of these plants was removed from the analysis.
-\rev{The final genotyping data set contained 39,261 SNPs and is available at .} \jri{Justin, please send me this and I can put online}
+\rev{The final genotyping data set contained 39,261 SNPs and is available at \url{http://figshare.com/articles/Gerke_et_al_Iowa_RRS/1515061}.}
 
 \subsection*{Integrating the genetic and physical map}
-%The interpretation of our results depends upon a genetic and physical map that is as accurate as possible. 
 \rev{Our simulation approach requires co-linear genetic and physical maps.}
@@
 This F1 was presented to fastPHASE with the rest of the data, except that its parent inbreds were removed. 
-Analyses of several hypothetical F1’s from different cycles of selection revealed very low phasing error rates (Table S3).  \jpg{S2 now?}
+Analyses of several hypothetical F1’s from different cycles of selection revealed very low phasing error rates (Table \ref{tab:phase_error}).  
 Therefore the phased genotypes of cycle 0 plants were used as the starting data for simulations (see below).
@@
 Consistent with previous studies of the Iowa RRS \citep{messmer1991genetic, labate1997molecular,  hagdorn2003molecular, hinze2005population}, genome-wide genetic diversity (expected heterozygosity, H) decreases steadily across cycles of selection in both populations (Figure \ref{fig:decline}).
-%\jri{do all these citations make sense here?
-I just copied these from the abstract}
 The loss of heterozygosity is smaller when the two populations are considered together, indicating the loss of different alleles within BSCB1 and BSSS. 
@@
 
-We noticed an irregular increase in the number of polymorphic markers between BSSS cycles 4 and 8 (Table \ref{tab:s2}). \jpg{S3 now?}
+We noticed an irregular increase in the number of polymorphic markers between BSSS cycles 4 and 8 (Table \ref{tab:contamination}). 
 All of these newly polymorphic markers were present at extremely low frequency and were spread among various individuals. 
@@
 Comparing the physical and genetic maps reveals that the regions of fixation are extremely small genetically, but some contain much of the physical content of the chromosome.
-} 
+
 \vspace{-6mm}
@@
 The observed data show higher than expected heterozygosity at cycle 4, \rev{which could be explained by several factors, including differences between simulated and actual breeding practices, under-sampling of diversity in cycle 0, selection acting to increase the frequency of initially rare haplotypes, or even selection acting to maintain heterozygosity \citep[e.g.][]{McMullen2009,Gore2009a} } \jri{where is this excess of diversity?
-if it's all pericentromeric can we argue for the first hypothesis?}
+if it's all pericentromeric can we argue for the Gore2009 hypothesis?}
 
@@
 This drift was driven by inbreeding through self-pollination and the small effective population sizes used to select a limited number of high-performing, potentially related individuals at each cycle.
+\jri{drosophila here?
+lamkey again replication isn't enough}
 
@@
 Given the lack of strong selection signals in each case, genetic drift has most likely played a large role in the current genetic structure of modern maize. } 	
-
 Several key inbreds in the ‘stiff-stalk’ heterotic group, B73, B37 and B14, were derived from the BSSS population \citep{darrah1986, troyer1999background}. 
@@
 
-\subsection*{Heterosis}
-
 The observation that the same targets of selection are not observed in the opposing heterotic populations bears implications for the genetic mechanisms responsible for heterosis and the success of maize hybrids.
@@
 Because most deleterious alleles are rare in both heterotic groups \citep{Mezmouk2014}, most haplotypes in the second population will have a different suite of deleterious variants and will complement the fixed haplotype reasonably well, such that selection will have little impact on diversity in the second population.
+\jri{likely these are not strongly deleterious, but lots of weakly deleterious}
 Our data are consistent with such a process having been important for hybrid improvement in the Iowa RSS experiment, and the lack of extreme differentiation seen over time across US Corn Belt germplasm is consistent with a role for such selection in maize hybrid improvement in general \citep{van2012historical}.

### 2015-08-26 sup1.tex
refs working
addressed Jeff's question in sup1.tex
fixed label in Table S3
minor edits with justin

@@
 \renewcommand{\thetable}{S\arabic{table}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
+\begin{figure*}   
+  \begin{center}
+   \vspace{-0mm}
+   \includegraphics[width=0.7\linewidth]{pca_founders}
+   %\renewcommand{\baselinestretch}{0.9}
+   \vspace{-3mm}
+   \caption{ Principle component analysis of founder inbred lines.
+The names of founder inbreds are shown on the graph; all other points represent BSSS (blue) and BSCB1 (red) individuals projected onto the PCA of the founders.
+} 
+\vspace{-6mm}
+    \label{fig:sfounders}
+  \end{center}
+\end{figure*}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
 
@@
 	B97 & BSCB1 & 9 &  \\ 
-	    \label{tab:s1}  % caption is needed to make this work
+	    \label{tab:founders}  % caption is needed to make this work
 \end{tabular}
@@
 \end{tabular}
-		    \label{tab:s2}  % caption is needed to make this work
+		    \label{tab:contamination}  % caption is needed to make this work
 \end{table}
@@
 \begin{tabular}{  l | l | l | }
-\multicolumn{1}{l}{} & \multicolumn{1}{l}{BSSS} & \multicolumn{1}{l}{BSSS}   \\ \cline{2-3}
+\multicolumn{1}{l}{} & \multicolumn{1}{l}{BSSS} & \multicolumn{1}{l}{BSCB1}   \\ \cline{2-3}
 	Mean & 0.052 & 0.052   \\ \cline{2-3}
@@
 \end{table}
-\jri{are  numbers in table supposed to be the same? I checked the Pages doc and they are the same there too.}
 
@@
 \end{tabular}
-	\label{tab:s3}  % caption is needed to make this work
+	\label{tab:phase_error}  % caption is needed to make this work
 \end{table}

### 2015-08-26 Fig4/excess_het.R
looked at regions with excess H

@@ -1,0 +1,11 @@
+
+rm(list=ls())
+library(ggplot2)
+library(dplyr)
+
+simresults = read.table("../data/2cm_sim_collated_results.txt", header=T, stringsAsFactors = F) %>%
+  mutate(Genlength = genstop - genstart, Physlength = (phystop-phystart)/1000000)
+
+excess_BSSS <- simresults %>% filter(BSSS_upperP < 0.001)
+excess_BSCB <- simresults %>% filter(BSCB_upperP < 0.001)
+

### 2015-08-27 response.tex
minor changes

@@
 We have switched this figure for the original figure 2.
-As shown in the new figure 3, there are a few regions on chromosome arms that have been fixed and span multiple cM.
-\jri{do we need to note in the main text?
-an interesting observation, maybe implying rapid selection?}}
+As shown in the new figure 3, there are a few regions on chromosome arms that have been fixed and span multiple cM.}
 
@@
 
-\response{This was an oversight on our part; we have now incorporated citation to some of this literature and hopefully clarified the discussion.}
+\response{We have now incorporated mention of several Drosophila papers in the context of population size and drift in experimental evolution.
+The only remaining personal communication regards the observation of sugary alleles as evidence of contamination, for which there is no publication.
+}
 
@@
 
-\response{\jri{Justin?}We have added this number to the text.}
+\response{We have added this number to the text.}
 

### 2015-08-27 main.tex
minor changes

@@
 \section*{Discussion}
-\jri{need to mention that the complementation we are seeing is not large effect deleterious that were likely purged during early inbreeding but instead many weakly deleterious things.
-need to add Drosophila and other cites.
-need to add Sprague 1983.
-then need to shorten/tighten up.}
 
@@
 This drift was driven by inbreeding through self-pollination and the small effective population sizes used to select a limited number of high-performing, potentially related individuals at each cycle.
-\jri{drosophila here?
+\jri{drosophila here.
 lamkey again replication isn't enough}
@@
 The observed pattern instead favors a dominance model, where fixation of a haplotype in one population simply selects against that same haplotype in the other population. 
-Because most deleterious alleles are rare in both heterotic groups \citep{Mezmouk2014}, most haplotypes in the second population will have a different suite of deleterious variants and will complement the fixed haplotype reasonably well, such that selection will have little impact on diversity in the second population.
-\jri{likely these are not strongly deleterious, but lots of weakly deleterious}
+\rev{Although strongly deleterious variants were likely purged during the inbreeding process leading to the founder lines, many weakly deleterious alleles can be found segregating at low frequencies among inbreds \citep{Mezmouk2014}.}
+Because deleterious alleles will be rare in both populations, most haplotypes in the second population will have a different suite of deleterious variants and will complement the fixed haplotype reasonably well.
+\rev{Selection against homozygosity of the fixed haplotype will thus} have little impact on diversity in the second population.
+
 Our data are consistent with such a process having been important for hybrid improvement in the Iowa RSS experiment, and the lack of extreme differentiation seen over time across US Corn Belt germplasm is consistent with a role for such selection in maize hybrid improvement in general \citep{van2012historical}.

### 2015-08-27 response.tex
edits

@@
 
-\response{We have now incorporated mention of several Drosophila papers in the context of population size and drift in experimental evolution.
+\response{We have now incorporated mention of additional papers (Drosophila and E.
+coli) in the context of population size and drift in experimental evolution.
 The only remaining personal communication regards the observation of sugary alleles as evidence of contamination, for which there is no publication.

### 2015-08-27 main.tex
edits

@@
 \abstract{
-Modern maize breeding relies upon selection in inbreeding populations to improve the performance of cross-population hybrids. 
+Although maize is naturally an outcrossing organism, modern breeding takes advantage of highly inbred lines that are used in controlled crosses to produce hybrids. 
 The United States Department of Agriculture reciprocal recurrent selection experiment between the Iowa Stiff Stalk Synthetic (BSSS) and the Iowa Corn Borer Synthetic No.
-1 (BSCB1) populations represents one of the longest standing models of selection for hybrid performance.
+1 (BSCB1) populations represents one of the longest \rev{running experiments to understand the genetic bases} for hybrid performance.
 To investigate the genomic impact of this selection program, \rev{we genotyped the progenitor lines and over 600 individuals across multiple cycles of selection using a genome-wide panel of $\sim 40,000$ SNPs.
-We first confirm previous results showing a steady temporal decrease in genetic diversity within populations and a corresponding increase in differentiation between populations.  
+We confirmed previous results showing a steady temporal decrease in genetic diversity within populations and a corresponding increase in differentiation between populations.  
 Thanks to detailed historical information on experimental design, we were able to perform extensive simulations using founder haplotypes to replicate the experiment in the absence of selection.
@@
 heterozygosity in each population has fallen more than expected. 
-We then take advantage of our high-density genotype data to identify extensive regions of haplotype fixation and trace haplotype ancestry to single founder inbred lines.
-The vast majority of regions showing such evidence of selection differ between the two populations, providing evidence against the overdominance model as a means of explaining selection for hybrid vigor.
-}
+We then took advantage of our high-density genotype data to identify extensive regions of haplotype fixation and trace haplotype ancestry to single founder inbred lines.
+The vast majority of regions showing such evidence of selection differ between the two populations, providing evidence for the dominance model of heterosis.}
 We discuss how this pattern is likely to occur during selection for hybrid performance, and how it poses challenges for dissecting the impacts of modern breeding and selection on the maize genome. }
@@
 \section*{Introduction} 
-Although maize is naturally an outcrossing organism, modern breeding takes advantage of highly inbred lines that are used in controlled crosses to produce hybrids. 
 Hybrid maize, first developed in the early 20th century \citep{crow1998}, was so successful that within a few decades it had completely replaced mass-selected open pollinated varieties in the United States \citep{crabb1947hybrid}.
-The shift towards selection of inbred lines based on their ability to generate good hybrids – referred to as ‘combining ability’ – constituted an abrupt change from the open-pollinated mass selection that breeders practiced for millennia \citep{anderson1944sources, troyer1999background}.
+The shift towards development of inbred lines based on their ability to generate good hybrids – referred to as ‘combining ability’ – constituted an abrupt change from the open-pollinated mass selection that breeders practiced for millennia \citep{anderson1944sources, troyer1999background}.
 Maize inbred lines are now partitioned into separate ‘heterotic groups’ that maximize performance and hybrid vigor (heterosis) \rev{for yield} when inbreds from different heterotic groups are crossed with each other \citep{tracy2006historical}.
@@
 
-\rev{In this study, we examine a different closed experiment whose phenotypic target and selection method --- reciprocal recurrent selection --- closely mirrors methods used in the generation of modern maize hybrids \citep{comstock1949breeding, duvick2004long}.
-Proposed by \citet{comstock1949breeding}, reciprocal recurrent selection is a procedure to simultaneously improve two populations for their ability to produce high yielding hybrids when individuals from the two populations are crossed.
-In this method, the two populations are kept separate and lines from each population are selected based on the phenotype of their cross-population hybrids. 
-Selected lines within each population are then intermated to generate lines for the next cycle. 
+\rev{In this study, we examine a different experiment in which the  method and target of selection --- reciprocal recurrent selection for hybrid yield --- closely mirror those used in the generation of modern maize hybrids \citep{comstock1949breeding, duvick2004long}.
+Proposed by \citet{comstock1949breeding}, reciprocal recurrent selection simultaneously improves two populations for their ability to produce high yielding hybrids when individuals from the two populations are crossed.
+In this method, the two populations are isolated genetically and lines from each population are selected based on the phenotype of their cross-population hybrids. 
+The superio lines within each population are then intermated to generate lines for the next cycle. 
 In this way, the two populations simultaneously evolve improved combining ability with one another. 
@@
 Extensive records on population sizes, breeding methods, selection differentials, and seed are available for all cycles of selection. 
-This set of resources on a closed population selected for hybrid yield make the Iowa RRS an ideal test case for the study of the genomic impact of intensive hybrid breeding in maize.} 
-The Iowa RRS experiment \rev{provides additional historical relevance} because lines derived from the BSSS population have had a major impact on the development of commercial hybrids \citep{duvick2004long, darrah1986}, the formation of modern heterotic groups \citep{troyer1999background, senior1998utility}, and the choice of a maize reference genome \citep{schnable2009the-b73-maize}.
+This set of resources makes the Iowa RRS an ideal test case for the study of the genomic impact of hybrid breeding in maize.} 
+The Iowa RRS experiment \rev{provides additional relevance} because lines derived from the BSSS population have had a major impact on the development of commercial hybrids \citep{duvick2004long, darrah1986}, the formation of modern heterotic groups \citep{troyer1999background, senior1998utility}, and the choice of a maize reference genome \citep{schnable2009the-b73-maize}.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRO
@@
 Nonetheless, we do observe \rev{differences from values generated under our neutral simulations} (Figure \ref{fig:compare_to_sims})
-The observed data show higher than expected heterozygosity at cycle 4, \rev{which could be explained by several factors, including differences between simulated and actual breeding practices, under-sampling of diversity in cycle 0, selection acting to increase the frequency of initially rare haplotypes, or even selection acting to maintain heterozygosity \citep[e.g.][]{McMullen2009,Gore2009a} } \jri{where is this excess of diversity?
-if it's all pericentromeric can we argue for the Gore2009 hypothesis?}
+The observed data show higher than expected heterozygosity at cycle 4, \rev{which could be explained by several factors, including differences between simulated and actual breeding practices, under-sampling of diversity in cycle 0, selection acting to increase the frequency of initially rare haplotypes, or even selection acting to maintain heterozygosity \citep[e.g.][]{McMullen2009,Gore2009a} }
 
@@
 These patterns of population structure, diversity, and differentiation between BSSS and BSCB1 can be largely reproduced by simulation without any selection, supporting the hypothesis that the majority of the genetic structure observed can be attributed to genetic drift alone, despite effective selection for phenotypic improvement.
-This drift was driven by inbreeding through self-pollination and the small effective population sizes used to select a limited number of high-performing, potentially related individuals at each cycle.
-\jri{drosophila here.
-lamkey again replication isn't enough}
+\rev{Similar observations have been recently made in other reciprocal recurrent selection programs, even in comparisons of multiple replicated populations \citep{romay2012effect,lamkey2014relative.
+These patterns differ markedly, however, from similar experimental evolution in  systems such as \emph{Drosophila} \citep{turner2011population, burke2010genome} and \emph{E.
+coli} \citep{tenaillon2012molecular} in which effective population sizes are maintained at levels much higher than in the present work and the effects of selection on diversity are readily discernible.}
 
-\rev{The trends observed in the Iowa RRS are reflected in similar recurrent selection experiments \citep{romay2012effect,lamkey2014relative}.
 Reciprocal recurrent selection serves as a model for the method of hybrid maize improvement \citep{duvick2004long}, and similar patterns of diversity and population structure can be seen broadly across North American maize germplasm \citep{van2012historical}.
@@
 Both our study and previous work \citep{labate1997molecular} identified genetic differentiation between cycle 0 of BSSS and the founder inbreds, \rev{likely reflecting the loss of some rare alleles.
-If this differentiation reflects drift during population maintenance before initiation of the RRS, it may have impacted the trajectory of the RRS program and of key inbreds derived from BSSS. 
+If this differentiation reflects drift during population maintenance before initiation of the RRS, it may have impacted the trajectory of the RRS program and its derived inbreds. 
 However, cycle 0 has also undergone maintenance since the beginning of the RRS experiment, so some} drift may have occurred during this time as well.
@@
 Although drift can explain most of the genetic structure genome-wide, phenotypic data provide clear evidence that selection has altered the frequencies of favorable alleles in the BSSS and BSCB1 populations.
-Numerous experiments have shown that the selected populations and the hybrids formed from them are superior to their cycle 0 ancestors \citep{smith1983evaluation, keeratinijakal1993responses, schnicker1993interpopulation, holthaus1995population}.
-Improvement occurred not only for hybrid yield, but concurrently for plant architecture and tolerance to high-density planting in both hybrids and inbreds \citep{brekke2011selection, brekke2011selectionb, edwards2011changes, lauer2012morphological}.
-Since these same trends are observed across wider North American germplasm \citep{tollenaar1989genetic}, identification and characterization of the loci conferring phenotypic improvement in BSSS and BSCB1 could reveal some of the genetic mechanisms by which maize yield has steadily improved over the decades.
-We find that the genomic regions of extremely low diversity evident at cycle 16 are unlikely to be produced by simulation, and heterozygosity falls more than expected across the genome as a whole.  
-However, simulated and observed values are often close, so overall the signature of selection is difficult to detect at any given locus. 
-This is due in large part to a limitation of statistical power imposed by inbreeding and low effective population sizes.
-\citet{lorenz2015selection} analyze a similar selection experiment with comparable effective population size and also find little little statistical evidence for selection on individual loci in spite of genetic gain for the trait of interest.
-Nonetheless, we show that an identity-by-descent, haplotype based approach provides additional power, as it can distinguish between the fixation of rare and common haplotypes. 
-However, the results of single locus simulations are sensitive to sampling error and drift caused by population maintenance. 
-In this case, it is difficult to assess significance with only a single population to test.
-This type of analysis is more effective across several replicated populations, which can control for genetic drift due to the independence between the selections in each replicate \citep{lamkey2014relative}.
+Numerous experiments have shown that the selected populations and the hybrids formed from them exhibit genetic gain for hybrid yield as well plant architecture and tolerance to high-density planting \citep{smith1983evaluation, keeratinijakal1993responses, schnicker1993interpopulation, holthaus1995population, brekke2011selection, brekke2011selectionb, edwards2011changes, lauer2012morphological}.
+Since these same trends are observed across wider North American germplasm \citep{tollenaar1989genetic}, identification and characterization of loci conferring phenotypic improvement in BSSS and BSCB1 could reveal some of the genetic mechanisms by which maize yield has steadily improved over the decades.
+We find that heterozygosity falls more than expected across the genome as a whole, and, \rev{though drift imposes limitations on the power to detect selection at individual loci,} genomic regions of extremely low diversity evident at cycle 16 are unlikely to be produced by simulation.
+We show that an identity-by-descent, haplotype based approach provides additional power, as it can distinguish between the fixation of rare and common haplotypes.
 
-\citet{van2012historical} assessed variation in and tested for selection across a wider range of North American maize germplasm using the same SNP array. 
-Although selected loci were identified in that study, the overall effects on genomic diversity and haplotype patterns were relatively small compared with the impact of domestication. 
-That study was a wide survey across many germplasm sources. 
-The limited overall impact of selection could arise because different haplotypes and loci are selected in different breeding programs. 
+\citet{van2012historical} assessed variation in and tested for selection across a wide range of North American maize germplasm using the same SNP array. 
+In spite of the large genetic gains made in decades of maize breeding \citep{duvick2005contribution}, \citet{van2012historical} identified  few loci showing  selected loci were identified in that study, the overall effects on genomic diversity and haplotype patterns were relatively small compared with the effects of genetic drift.
+One explanation for the limited overall impact of selection could arise because different haplotypes and loci are selected in different breeding programs. 
 In support of this idea, we find that the most likely targets of selection between the BSSS and BSCB1 populations occur at non-overlapping loci. 

### 2015-08-27 main.tex
edits to discussion in progress

@@
 There is no separation between the founders of the two populations, \rev{and projection of later cycles onto a PCA of the founders shows no distinction between BSSS and BSCB1 (Figure \ref{fig:sfounders}).
-At cycle 0, however,} the BSSS population shows more divergence from the founders than does BSCB1. 
-This could be due to drift during either the population’s construction or subsequent maintenance. 
+At cycle 0, however,} the BSSS population shows more divergence from the founders than does BSCB1, likely due to drift during either the population’s construction or subsequent maintenance. 
 Structure continued to develop within each population over the course of the experiment. 
@@
 These patterns of population structure, diversity, and differentiation between BSSS and BSCB1 can be largely reproduced by simulation without any selection, supporting the hypothesis that the majority of the genetic structure observed can be attributed to genetic drift alone, despite effective selection for phenotypic improvement.
-\rev{Similar observations have been recently made in other reciprocal recurrent selection programs, even in comparisons of multiple replicated populations \citep{romay2012effect,lamkey2014relative.
+\rev{Similar observations have been recently made in other reciprocal recurrent selection programs, even in comparisons of multiple replicated populations \citep{romay2012effect,lamkey2014relative}.
 These patterns differ markedly, however, from similar experimental evolution in  systems such as \emph{Drosophila} \citep{turner2011population, burke2010genome} and \emph{E.
@@
 
-Reciprocal recurrent selection serves as a model for the method of hybrid maize improvement \citep{duvick2004long}, and similar patterns of diversity and population structure can be seen broadly across North American maize germplasm \citep{van2012historical}.
-Given the lack of strong selection signals in each case, genetic drift has most likely played a large role in the current genetic structure of modern maize. } 	
+Reciprocal recurrent selection serves as a model for the method of hybrid maize improvement \citep{duvick2004long} and similar patterns of diversity and population structure can be seen broadly across North American maize germplasm \citep{van2012historical}, suggesting , genetic drift has most likely played a large role in the current genetic structure of modern maize.
 Several key inbreds in the ‘stiff-stalk’ heterotic group, B73, B37 and B14, were derived from the BSSS population \citep{darrah1986, troyer1999background}. 
@@
 Near isogenic lines that substitute these centromeric haplotypes from key inbreds could be used to determine if the haplotypes at high frequency in cycle 16 of BSSS provide a phenotypic advantage when crossed with BSCB1.
+\jri{removed paragraph on drift at cycle 0.
+shortens discussion, felt out of place and we mention both possibilities already in the results}
 
-Both our study and previous work \citep{labate1997molecular} identified genetic differentiation between cycle 0 of BSSS and the founder inbreds, \rev{likely reflecting the loss of some rare alleles.
-If this differentiation reflects drift during population maintenance before initiation of the RRS, it may have impacted the trajectory of the RRS program and its derived inbreds. 
-However, cycle 0 has also undergone maintenance since the beginning of the RRS experiment, so some} drift may have occurred during this time as well.
+%Both our study and previous work \citep{labate1997molecular} identified genetic differentiation between cycle 0 of BSSS and the founder inbreds, \rev{likely reflecting the loss of some rare alleles.
+%If this differentiation reflects drift during population maintenance before initiation of the RRS, it may have impacted the trajectory of the RRS program and its derived inbreds. 
+%However, cycle 0 has also undergone maintenance since the beginning of the RRS experiment, so some} drift may have occurred during this time as well.
 
 Although drift can explain most of the genetic structure genome-wide, phenotypic data provide clear evidence that selection has altered the frequencies of favorable alleles in the BSSS and BSCB1 populations.
-Numerous experiments have shown that the selected populations and the hybrids formed from them exhibit genetic gain for hybrid yield as well plant architecture and tolerance to high-density planting \citep{smith1983evaluation, keeratinijakal1993responses, schnicker1993interpopulation, holthaus1995population, brekke2011selection, brekke2011selectionb, edwards2011changes, lauer2012morphological}.
+Numerous experiments have shown that the selected populations and the hybrids formed from them \rev{exhibit genetic gain for hybrid yield as well plant architecture and tolerance to high-density planting} \citep{smith1983evaluation, keeratinijakal1993responses, schnicker1993interpopulation, holthaus1995population, brekke2011selection, brekke2011selectionb, edwards2011changes, lauer2012morphological}.
 Since these same trends are observed across wider North American germplasm \citep{tollenaar1989genetic}, identification and characterization of loci conferring phenotypic improvement in BSSS and BSCB1 could reveal some of the genetic mechanisms by which maize yield has steadily improved over the decades.
 We find that heterozygosity falls more than expected across the genome as a whole, and, \rev{though drift imposes limitations on the power to detect selection at individual loci,} genomic regions of extremely low diversity evident at cycle 16 are unlikely to be produced by simulation.
-We show that an identity-by-descent, haplotype based approach provides additional power, as it can distinguish between the fixation of rare and common haplotypes.
+We further show that an identity-by-descent, haplotype-based approach provides additional power to identify selected regions, as it can distinguish between the fixation of rare and common haplotypes.
 

### 2015-08-27 main.tex
supp tables

@@
 \usepackage{subcaption}
+\usepackage{tabularx}
 \usepackage{geneticsT2}
@@
 \subsection*{Plants and inbred lines used}
-The plants and inbred lines used in this experiment are listed in Table \ref{tab:founders}. 
-We genotyped 36 plants from each of the BSSS and BSCB1 populations at selection cycles 0, 4, 8, 12, and 16. 
+The plants and inbred lines used in this experiment are listed in Tables \ref{tab:founders} and \ref{tab:derived}. 
+We genotyped 34-36 plants from each of the BSSS and BSCB1 populations at selection cycles 0, 4, 8, 12, and 16 (Table \ref{tab:plants}).
 These plants represent descendants of the original populations which have been randomly mated to maintain seed. 

### 2015-08-27 sup1.tex
edits to discussion in progress
supp tables

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FIGURE
+\clearpage
 
@@
 \end{center}
+
+\clearpage
 
@@
 
-\newpage
-
 \begin{table}
-\caption{Plants and Lines Genotyped in this study.
-Backgrounds of the founder inbreds are from \citet{hagdorn2003molecular}}
-\begin{tabular}{ | l | l | l | l | }
-
-	{\bf Plants from the selection cycles} &  &  &  \\ 
-	Population  & Cycle & \# plants &   \\ 
-	BSSS & 0 & 34 &  \\ 
-	BSSS & 4 & 36 &  \\ 
-	BSSS & 8 & 35 &  \\ 
-	BSSS & 12 & 36 &  \\ 
-	BSSS & 16 & 36 &  \\ 
-	BSCB1 & 0 & 36 &  \\ 
-	BSCB1 & 4 & 36 &  \\ 
-	BSCB1 & 8 & 35 &  \\ 
-	BSCB1 & 12 & 36 &  \\ 
-	BSCB1 & 16 & 36 &  \\ 
-	 &  &  &  \\ 
-	BSSS Founders Genotyped &  &  &  \\ 
-	Inbred & Background / Pedigree &  &  \\ 
+\caption{Founder lines genotyped. } %Backgrounds of the founder inbreds are from\citet{hagdorn2003molecular}
+	\subcaption*{BSSS Founders Genotyped} 
+	\begin{tabular}{ lllll }
+	Inbred & Background / Pedigree &  & Notes \\ \hline
 	Ind\_Tr9\_1\_1\_6 & Reid Early Dent (Troyer Strain) &  &  \\ 
@@
 	A3G-3-1-3 & BL345BxIAI129 &  &  \\ 
-	Ind\_Fe2\_1073* & Troyer Reid (Early) &  &  \\ 
+	Ind\_Fe2\_1073 & Troyer Reid (Early) &  & Parent of unavailable line F1B1 \\ 
 	Ill\_Hy & IL High Yield &  &  \\ 
@@
 	Ind\_AH83 & Funk 176A &  &  \\ 
-	Ind\_B2* & Troyer Reid (Late Butler), parent of a founder &  &  \\ 
-	LE23 & IL Low Ear &  &  \\ 
-	*Parents of unavailable line F1B1 &  &  &  \\ 
-	 &  &  &  \\ 
-	BSCB1 Founders Genotyped &  &  &  \\ 
-	Inbred & Background / Pedigree &  &  \\ 
+	Ind\_B2 & Troyer Reid (Late Butler) &  & Parent of unavailable line F1B1 \\ 
+	LE23 & IL Low Ear &  &  \\ \hline
+	\end{tabular} 
+	\bigskip
+
+	\subcaption*{BSCB1 Founders Genotyped} 
+	\begin{tabular}{lll}
+	Inbred & Background / Pedigree \\ \hline
 	I205 & Iodent &  &  \\ 
-	Oh51A & [(OH56xWf9)Oh56] (Wooster Clarage x ?) &  &  \\ 
-	A340 & 4-29 x 64 (Silver King x Northwestern Dent) &  &  \\ 
-	Ill\_Hy & IL High Yield &  &  \\ 
-	Oh33 & Clarage &  &  \\ 
-	Oh07 & C.I.540xIII.L &  &  \\ 
-	R4 & Funk Yellow Dent &  &  \\ 
-	Oh40B & eight line LSC composite &  &  \\ 
-	P8 & Palin Reid &  &  \\ 
-	L317 & LSC &  &  \\ 
-	CC5 & Golden Glow (W23) &  &  \\ 
-	 &  &  &  \\ 
-	Founders not analyzed &  &  &  \\ 
-	Inbred & Group & Reason &  \\ 
-	CI.540 & BSSS & heterozygous genotype &  \\ 
-	F1B1 & BSSS & unavailable &  \\ 
-	CI.617 & BSSS & unavailable &  \\ 
-	WD456 & BSSS & unavailable &  \\ 
-	K230 & BSCB1 & source segregates phenotypically &  \\ 
-	 &  &  &  \\ 
-	Derived Lines Used &  &  &  \\ 
-	Inbred & Group & Cycle & Notes \\ 
+	Oh51A & [(OH56xWf9)Oh56] (Wooster Clarage x ?)   \\ 
+	A340 & 4-29 x 64 (Silver King x Northwestern Dent)   \\ 
+	Ill\_Hy & IL High Yield   \\ 
+	Oh33 & Clarage  \\ 
+	Oh07 & C.I.540xIII.L  \\ 
+	R4 & Funk Yellow Dent   \\ 
+	Oh40B & eight line LSC composite  \\ 
+	P8 & Palin Reid   \\ 
+	L317 & LSC   \\ 
+	CC5 & Golden Glow (W23)  \\ 
+	\end{tabular}
+
+	\subcaption*{Founders not analyzed} 
+	\begin{tabular}{llll}
+	Inbred & Group & Reason   \\ \hline
+	CI.540 & BSSS & heterozygous genotype  \\ 
+	F1B1 & BSSS & unavailable   \\ 
+	CI.617 & BSSS & unavailable   \\ 
+	WD456 & BSSS & unavailable   \\ 
+	K230 & BSCB1 & source segregates phenotypically   \\ 
+	\end{tabular}
+\label{tab:founders}
+\end{table}
+
+\clearpage
+\begin{table}
+\caption{Derived lines genotyped. }	\begin{tabular}{llll}
+	Inbred & Group & Cycle & Notes \\ \hline
 	B10 & BSSS & 0 & thrown out; poor data quality \\ 
@@
 	B97 & BSCB1 & 9 &  \\ 
-	    \label{tab:founders}  % caption is needed to make this work
 	\end{tabular}
+	\label{tab:derived}  % caption is needed to make this work
 \end{table}
-\newpage
+
+\clearpage
 
 \begin{table}
-\caption{Evidence of minor contamination between cycles 4 and 8 in BSSS.}
-\begin{tabular}{ | l | l | l | l | }
-\hline
-	\textbf{SNPs polymorphic in cycle} & \textbf{But not cycle} & \textbf{BSSS} & \textbf{BSCB1} \\ \hline
-	4 & 0 & 290 & 225 \\ \hline
-	8 & 0 & 2242 & 319 \\ \hline
-	12 & 0 & 1227 & 181 \\ \hline
-	16 & 0 & 344 & 128 \\ \hline
-	8 & 4 & 4499 & 1081 \\ \hline
-	12 & 4 & 2328 & 669 \\ \hline
-	16 & 4 & 551 & 477 \\ \hline
-	12 & 8 & 1821 & 1405 \\ \hline
-	16 & 8 & 178 & 997 \\ \hline
-	16 & 12 & 542 & 666 \\ \hline
-	0 & Founders & 1202 & 1707 \\ \hline
-	4 & Founders & 822 & 885 \\ \hline
-	8 & Founders & 1201 & 798 \\ \hline
-	12 & Founders & 816 & 550 \\ \hline
-	16 & Founders & 445 & 460 \\ \hline
+\caption{Number of individuals genotyped at each cycle.}
+\centering
+\begin{tabular}{  cccc }
+	Population  & Cycle & \# plants &   \\ \hline
+	BSSS & 0 & 34 &  \\ 
+	BSSS & 4 & 36 &  \\ 
+	BSSS & 8 & 35 &  \\ 
+	BSSS & 12 & 36 &  \\ 
+	BSSS & 16 & 36 &  \\ 
+	BSCB1 & 0 & 36 &  \\ 
+	BSCB1 & 4 & 36 &  \\ 
+	BSCB1 & 8 & 35 &  \\ 
+	BSCB1 & 12 & 36 &  \\ 
+	BSCB1 & 16 & 36 &  \\ 
 \end{tabular}
-		    \label{tab:contamination}  % caption is needed to make this work
+\label{tab:plants}
 \end{table}
-\newpage
-
-\begin{table}
-\caption{Heterozygosity of SNPs in cycle 8 but not cycle 4 of the BSSS. }
-\begin{tabular}{  l | l | l | }
-\multicolumn{1}{l}{} & \multicolumn{1}{l}{BSSS} & \multicolumn{1}{l}{BSCB1}   \\ \cline{2-3}
-	Mean & 0.052 & 0.052   \\ \cline{2-3}
-	Median & 0.028 & 0.028   \\ \cline{2-3}
-\end{tabular}
-\end{table}
+\clearpage
 
@@
 \caption{Switch error rates from computationally phasing ‘hybrids’ simulated from derived inbred lines.}
+\centering
+
 \begin{tabular}{ | l | l | l | l | l | l | }
@@
 \end{table}
+
+\clearpage
+
+\begin{table}
+\caption{Evidence of minor contamination between cycles 4 and 8 in BSSS.}
+\centering
+
+\subcaption*{Polymorphic SNPs}
+\begin{tabular}{ cccc }
+	{SNPs polymorphic in cycle} & {But not cycle} & {BSSS} & {BSCB1} \\ \hline
+	4 & 0 & 290 & 225 \\ 
+	8 & 0 & 2242 & 319 \\
+	12 & 0 & 1227 & 181 \\ 
+	16 & 0 & 344 & 128 \\ 
+	8 & 4 & 4499 & 1081 \\ 
+	12 & 4 & 2328 & 669 \\ 
+	16 & 4 & 551 & 477 \\ 
+	12 & 8 & 1821 & 1405 \\ 
+	16 & 8 & 178 & 997 \\ 
+	16 & 12 & 542 & 666 \\ 
+	0 & Founders & 1202 & 1707 \\ 
+	4 & Founders & 822 & 885 \\ 
+	8 & Founders & 1201 & 798 \\ 
+	12 & Founders & 816 & 550 \\ 
+	16 & Founders & 445 & 460 \\ \hline
+\end{tabular}
+\bigskip
+\subcaption*{Heterozygosity of SNPs in cycle 8 but not cycle 4 of the BSSS. }
+\begin{tabular}{  lcc }
+\multicolumn{1}{l}{} & \multicolumn{1}{l}{BSSS} & \multicolumn{1}{l}{BSCB1}   \\ \cline{2-3}
+	Mean & 0.052 & 0.052   \\ 
+	Median & 0.028 & 0.028   \\ 
+\end{tabular}
+		    \label{tab:contamination}  % caption is needed to make this work
+\end{table}

### 2015-08-27 main.tex
edits to discussion -- jri finalish draft

@@
 Although selective sweeps remaining from domestication are clearly visible, the impact of selection during modern breeding appears comparatively small in terms of its impact on genomic diversity \citep{Hufford2012b,van2012historical}, despite steady, heritable improvement in phenotype \citep{duvick2005contribution}.
-The lack of distinct selection signals \rev{in broad germplasm collections may be due to population-specific selection within germplasm subgroups; in this case selection should be easier to detect in populations maintained in an individual program.} This \rev{possibliity} is supported by the measured success in identifying targets of selection in individual experimental populations \rev{under directional selection for specific phenotypes such as seed size and ear number} \citep{hirsch2014insights, beissinger2014genome}.
+The lack of distinct selection signals \rev{in broad germplasm collections may be due to population-specific selection within germplasm subgroups; in this case selection should be easier to detect in populations maintained in an individual program.}
+This \rev{possibliity} is supported by the measured success in identifying targets of selection in individual experimental populations \rev{under directional selection for specific phenotypes such as seed size and ear number} \citep{hirsch2014insights, beissinger2014genome}.
 
@@
 
-Reciprocal recurrent selection serves as a model for the method of hybrid maize improvement \citep{duvick2004long} and similar patterns of diversity and population structure can be seen broadly across North American maize germplasm \citep{van2012historical}, suggesting , genetic drift has most likely played a large role in the current genetic structure of modern maize.
-Several key inbreds in the ‘stiff-stalk’ heterotic group, B73, B37 and B14, were derived from the BSSS population \citep{darrah1986, troyer1999background}. 
-B37 and B14 were derived from cycle 0, and B73 was derived from a half-sib recurrent selection program also started with the BSSS population. 
-We examined these three inbreds at the pericentromeric regions listed in Table \ref{tab:fix}, and found that in most cases they carry different haplotypes from those that rose to high frequency in the RRS experiment.
-Near isogenic lines that substitute these centromeric haplotypes from key inbreds could be used to determine if the haplotypes at high frequency in cycle 16 of BSSS provide a phenotypic advantage when crossed with BSCB1.
-\jri{removed paragraph on drift at cycle 0.
-shortens discussion, felt out of place and we mention both possibilities already in the results}
-
-%Both our study and previous work \citep{labate1997molecular} identified genetic differentiation between cycle 0 of BSSS and the founder inbreds, \rev{likely reflecting the loss of some rare alleles.
-%If this differentiation reflects drift during population maintenance before initiation of the RRS, it may have impacted the trajectory of the RRS program and its derived inbreds. 
-%However, cycle 0 has also undergone maintenance since the beginning of the RRS experiment, so some} drift may have occurred during this time as well.
-
 Although drift can explain most of the genetic structure genome-wide, phenotypic data provide clear evidence that selection has altered the frequencies of favorable alleles in the BSSS and BSCB1 populations.
 Numerous experiments have shown that the selected populations and the hybrids formed from them \rev{exhibit genetic gain for hybrid yield as well plant architecture and tolerance to high-density planting} \citep{smith1983evaluation, keeratinijakal1993responses, schnicker1993interpopulation, holthaus1995population, brekke2011selection, brekke2011selectionb, edwards2011changes, lauer2012morphological}.
-Since these same trends are observed across wider North American germplasm \citep{tollenaar1989genetic}, identification and characterization of loci conferring phenotypic improvement in BSSS and BSCB1 could reveal some of the genetic mechanisms by which maize yield has steadily improved over the decades.
 We find that heterozygosity falls more than expected across the genome as a whole, and, \rev{though drift imposes limitations on the power to detect selection at individual loci,} genomic regions of extremely low diversity evident at cycle 16 are unlikely to be produced by simulation.
 We further show that an identity-by-descent, haplotype-based approach provides additional power to identify selected regions, as it can distinguish between the fixation of rare and common haplotypes. 
+\rev{These analyses show that most likely targets of selection occur at different loci in the two populations, a result consistent with analyses in commercial breeding programs \citep{feng2006temporal} and which may help explain the lack of selection seen in previous analyses across numerous breeding programs \citep{van2012historical}.}
 
-\citet{van2012historical} assessed variation in and tested for selection across a wide range of North American maize germplasm using the same SNP array. 
-In spite of the large genetic gains made in decades of maize breeding \citep{duvick2005contribution}, \citet{van2012historical} identified  few loci showing  selected loci were identified in that study, the overall effects on genomic diversity and haplotype patterns were relatively small compared with the effects of genetic drift.
-One explanation for the limited overall impact of selection could arise because different haplotypes and loci are selected in different breeding programs. 
-In support of this idea, we find that the most likely targets of selection between the BSSS and BSCB1 populations occur at non-overlapping loci. 
-This non-overlap between putatively selected loci in complementary populations has also been observed in commercial breeding programs \citep{feng2006temporal}.
-
-The observation that the same targets of selection are not observed in the opposing heterotic populations bears implications for the genetic mechanisms responsible for heterosis and the success of maize hybrids.
-Classic overdominance models of heterosis predict that at a single locus, two distinct alleles confer heterozygote advantage when combined, while the dominance model predicts that heterosis is driven by dominance effects and the complementation of linked alleles in low-recombination regions (dominance or pseudo-overdominance).
-In the case of true overdominance,  selection should lead to decreased heterozygosity at a locus in both populations as complementary haplotypes are fixed in each group \rev{\citep[e.g.][]{guo2014maize}}.
+The observation that different targets of selection are  observed in  opposing heterotic populations bears implications for the genetic mechanisms responsible for heterosis and the success of maize hybrids.
+Classic overdominance models of heterosis predict that at a single locus two distinct alleles confer heterozygote advantage when combined.
+Alternatively, the dominance model predicts that heterosis is driven by dominance effects and the complementation of linked alleles in low-recombination regions (dominance or pseudo-overdominance). 
+In the case of true overdominance, selection should thus lead to decreased heterozygosity at the same locus in both populations as complementary haplotypes are fixed in each group \rev{\citep[e.g.][]{guo2014maize}}.
 We find \rev{little evidence to support this genetic phenomenon, finding only two 2 cM windows genome-wide in which both populations show significantly reduced heterozygosity.} 
@@
 \rev{Selection against homozygosity of the fixed haplotype will thus} have little impact on diversity in the second population.
-
-Our data are consistent with such a process having been important for hybrid improvement in the Iowa RSS experiment, and the lack of extreme differentiation seen over time across US Corn Belt germplasm is consistent with a role for such selection in maize hybrid improvement in general \citep{van2012historical}.
 
@@
 In the case of individual breeding programs, genome-wide genotyping such as we conducted here can identify the lines carrying the recombinant haplotypes and introgressions necessary to conduct such an experiment.
+
+Reciprocal recurrent selection serves as a model for the method of hybrid maize improvement \citep{duvick2004long}.
+Similar patterns of diversity and population structure can be seen broadly across North American maize germplasm \citep{van2012historical}, suggesting that genetic drift has most likely played a large role in the current genetic structure of modern maize.
+Nonetheless, breeding has contributed substantially to yield improvements \citep{duvick2005contribution}, and the identification and characterization of loci conferring phenotypic improvement in BSSS and BSCB1 could reveal some of the genetic mechanisms by which maize yield has steadily improved over the decades.
+Several key inbreds in the ‘stiff-stalk’ heterotic group --- B73, B37 and B14 --- were derived from the BSSS population \citep{darrah1986, troyer1999background}. 
+%B37 and B14 were derived from cycle 0, and B73 was derived from a half-sib recurrent selection program also started with the BSSS population. 
+We examined these three inbreds at the pericentromeric regions listed in Table \ref{tab:fix}, and found that in most cases they carry different haplotypes from those that rose to high frequency in the RRS experiment.
+Near isogenic lines that substitute these centromeric haplotypes from key inbreds could be used to determine if the haplotypes at high frequency in cycle 16 of BSSS provide a phenotypic advantage when crossed with BSCB1.
+Our observation of population-specific selection is consistent with a model of selection for complementation in the Iowa RSS experiment, and the lack of extreme differentiation seen over time across US Corn Belt germplasm may suggest a role for such selection in maize hybrid improvement in general \citep{van2012historical}.
 

### 2015-08-27 sup1.tex
edits to discussion -- jri finalish draft

@@
 
-\begin{tabular}{ | l | l | l | l | l | l | }
-
-\hline
-	\textbf{Simulated 'Hybrid'} & \textbf{Derived Lines Used as Priors} & \textbf{Population} & \textbf{Possible Switches} & \textbf{Switch errors} & \textbf{Rate} \\ \hline
-	B11xB67 & all & BSSS & 12195 & 41 & 0.003 \\ \hline
-	B17xB44 & all & BSSS & 12230 & 58 & 0.005 \\ \hline
-	B39xB37 & all & BSSS & 11485 & 129 & 0.011 \\ \hline
-	B43xB69 & all & BSSS & 12266 & 26 & 0.002 \\ \hline
-	B73xB72 & all & BSSS & 12043 & 47 & 0.004 \\ \hline
-	B78xB94 & all & BSSS & 11658 & 51 & 0.004 \\ \hline
-	B89xB84 & all & BSSS & 11517 & 53 & 0.005 \\ \hline
-	B11xB67 & cycle 0 & BSSS & 12195 & 45 & 0.004 \\ \hline
-	B17xB44 & cycle 0 & BSSS & 12230 & 54 & 0.004 \\ \hline
-	B39xB37 & cycle 0 & BSSS & 11485 & 126 & 0.011 \\ \hline
-	BB43xB69 & cycle 0 & BSSS & 12266 & 32 & 0.003 \\ \hline
-	B42xB54 & all & BSCB1 & 11260 & 38 & 0.003 \\ \hline
-	B90xB97 & all & BSCB1 & 6499 & 52 & 0.008 \\ \hline
-	B91xB95 & all & BSCB1 & 7830 & 56 & 0.007 \\ \hline
-	B99xB97 & all & BSCB1 & 6779 & 52 & 0.008 \\ \hline
+\begin{tabular}{ cccccc  }
+	{Simulated 'Hybrid'} & {Derived Lines Used as Priors} & {Population} & {Possible Switches} & {Switch errors} & {Rate} \\ \hline
+	B11xB67 & all & BSSS & 12195 & 41 & 0.003 \\ 
+	B17xB44 & all & BSSS & 12230 & 58 & 0.005 \\ 
+	B39xB37 & all & BSSS & 11485 & 129 & 0.011 \\ 
+	B43xB69 & all & BSSS & 12266 & 26 & 0.002 \\ 
+	B73xB72 & all & BSSS & 12043 & 47 & 0.004 \\ 
+	B78xB94 & all & BSSS & 11658 & 51 & 0.004 \\ 
+	B89xB84 & all & BSSS & 11517 & 53 & 0.005 \\ 
+	B11xB67 & cycle 0 & BSSS & 12195 & 45 & 0.004 \\ 
+	B17xB44 & cycle 0 & BSSS & 12230 & 54 & 0.004 \\ 
+	B39xB37 & cycle 0 & BSSS & 11485 & 126 & 0.011 \\ 
+	BB43xB69 & cycle 0 & BSSS & 12266 & 32 & 0.003 \\ 
+	B42xB54 & all & BSCB1 & 11260 & 38 & 0.003 \\ 
+	B90xB97 & all & BSCB1 & 6499 & 52 & 0.008 \\ 
+	B91xB95 & all & BSCB1 & 7830 & 56 & 0.007 \\ 
+	B99xB97 & all & BSCB1 & 6779 & 52 & 0.008 \\ 
 	BB42xB54 & cycle 0 & BSCB1 & 11260 & 57 & 0.005 \\\hline
@@
 \centering
-
 \subcaption*{Polymorphic SNPs}

### 2015-08-27 main.tex
edits to Intro/Abstract

@@
 \abstract{
-Although maize is naturally an outcrossing organism, modern breeding takes advantage of highly inbred lines that are used in controlled crosses to produce hybrids. 
+Although maize is naturally an outcrossing organism, modern breeding develops of highly inbred lines that are used in controlled crosses to produce hybrids. 
 The United States Department of Agriculture reciprocal recurrent selection experiment between the Iowa Stiff Stalk Synthetic (BSSS) and the Iowa Corn Borer Synthetic No.
-1 (BSCB1) populations represents one of the longest \rev{running experiments to understand the genetic bases} for hybrid performance.
+1 (BSCB1) populations represents one of the longest \rev{running experiments to understand the response to selection} \jpg{changed the wording here, because the experiment's goal was more broad than undesrtanding the genetic basis} for hybrid performance.
 To investigate the genomic impact of this selection program, \rev{we genotyped the progenitor lines and over 600 individuals across multiple cycles of selection using a genome-wide panel of $\sim 40,000$ SNPs.
@@
 \section*{Introduction} 
-Hybrid maize, first developed in the early 20th century \citep{crow1998}, was so successful that within a few decades it had completely replaced mass-selected open pollinated varieties in the United States \citep{crabb1947hybrid}.
-The shift towards development of inbred lines based on their ability to generate good hybrids – referred to as ‘combining ability’ – constituted an abrupt change from the open-pollinated mass selection that breeders practiced for millennia \citep{anderson1944sources, troyer1999background}.
+Hybrid maize, first developed in the early 20th century \citep{crow1998}, rapidly and completely replaced mass-selected open pollinated varieties in the United States \citep{crabb1947hybrid}. 
+The shift towards development of inbred lines based on their ability to generate good hybrids –-- referred to as ‘combining ability’ –-- constituted an abrupt change from the open-pollinated mass selection that breeders practiced for millennia \citep{anderson1944sources, troyer1999background}.
 Maize inbred lines are now partitioned into separate ‘heterotic groups’ that maximize performance and hybrid vigor (heterosis) \rev{for yield} when inbreds from different heterotic groups are crossed with each other \citep{tracy2006historical}.
@@
 These studies have examined a broad spectrum of germplasm at various points in the history of maize to search for the signals of population structure and artificial selection \citep{Hufford2012b, van2012historical,jiao2012genome}.
-Although selective sweeps remaining from domestication are clearly visible, the impact of selection during modern breeding appears comparatively small in terms of its impact on genomic diversity \citep{Hufford2012b,van2012historical}, despite steady, heritable improvement in phenotype \citep{duvick2005contribution}.
+Although selective sweeps from the initial domestication of maize are clearly visible, localized genomic signals of selection during modern breeding remain difficult to observe \citep{Hufford2012b,van2012historical} despite steady, heritable improvement in phenotype \citep{duvick2005contribution}.
 The lack of distinct selection signals \rev{in broad germplasm collections may be due to population-specific selection within germplasm subgroups; in this case selection should be easier to detect in populations maintained in an individual program.}
@@
 \rev{In this study, we examine a different experiment in which the method and target of selection --- reciprocal recurrent selection for hybrid yield --- closely mirror those used in the generation of modern maize hybrids \citep{comstock1949breeding, duvick2004long}.
-Proposed by \citet{comstock1949breeding}, reciprocal recurrent selection simultaneously improves two populations for their ability to produce high yielding hybrids when individuals from the two populations are crossed.
-In this method, the two populations are isolated genetically and lines from each population are selected based on the phenotype of their cross-population hybrids. 
-The superio lines within each population are then intermated to generate lines for the next cycle. 
-In this way, the two populations simultaneously evolve improved combining ability with one another. 
+Reciprocal recurrent selection is a method, initially proposed by \citet{comstock1949breeding}, in which lines from two populations are evaluated based on the phenotype of the hybrids each line produces when crossed with the opposing population.
+The selected lines are then intermated within each population to generate lines for the next cycle of recurrent selection.
+%Reciprocal recurrent selection simultaneously improves two populations for their ability to produce high yielding hybrids when individuals from the two populations are crossed. 
+This procedure results in two closed and genetically isolated populations that simultaneously evolve improved combining ability with one another. 
 The USDA-ARS at Ames, Iowa has conducted a reciprocal recurrent selection program (hereafter referred to as the Iowa RRS) with the Iowa Stiff Stalk Synthetic (BSSS) and Iowa Corn Borer Synthetic No.
-1 (BSCB1) populations for 18 cycles of selection \citep{penny1971twenty}.
+1 (BSCB1) populations for 18 cycles of selection for hybrid yield \citep{penny1971twenty}.
+\jpg{I think we need Jode's reference here as well if we are going to mention it has gone through 18 cycles.}
 This program represents one of the best-documented public experiments on selection for combining ability and hybrid performance. 
@@
 The BSSS population was formed from 16 inbred lines selected for stalk strength in 1933 and 1934 (Sprague, 1946) and the BSCB1 population was formed from a set of 12 inbred lines \citep{hallauer1974registration}.}
-Founder lines (Table \ref{tab:founders}) from each population were randomly mated to create the `cycle 0' base population.
+The founder inbreds (Table \ref{tab:founders}) were randomly mated to create each `cycle 0' base population.
 \rev{Testcross progenies were formed by self-pollinating 100 individuals in each population and simultaneously crossing the same 100 individuals onto 10 plants (as females) in the reciprocal population (e.g., a BSSS plant was self-pollinated and crossed to 10 females in BSCB1).
@@
 Change in population structure throughout the Iowa RRS experiment can be observed visually by a principal component analysis (PCA). 
-A joint analysis of individuals from all the selection cycles (Figure \ref{fig:pca}) clearly separates the BSSS and BSCB1 populations along the first axis of variation, with increasing separation as the experiment progressed.
+Analysis of individuals from all the selection cycles (Figure \ref{fig:pca}) clearly separates the BSSS and BSCB1 populations along the first axis of variation, with increasing separation as the experiment progressed.
 The second axis of variation primarily separates the cycles from one another within each population. 
@@
 
-Our analysis of diversity in the Iowa RRS experiment reveals a steady loss of diversity in the BSSS and BSCB1 populations as they became increasingly differentiated from one another over time.
+Our analysis of the Iowa RRS experiment reveals a steady loss of diversity in the BSSS and BSCB1 populations as they became increasingly differentiated from one another over time.
 Principal component analysis shows that as the effective population size and the rates of inbreeding were altered, the rates of change in population structure were altered as well. 

### 2015-08-27 main.tex
edits to Discussion

@@
 These patterns differ markedly, however, from similar experimental evolution in systems such as \emph{Drosophila} \citep{turner2011population, burke2010genome} and \emph{E.
-coli} \citep{tenaillon2012molecular} in which effective population sizes are maintained at levels much higher than in the present work and the effects of selection on diversity are readily discernible.}
+coli} \citep{tenaillon2012molecular} in which the effects of selection on diversity are readily discernible.
+A key difference between these studies and those in maize are the effective population sizes, which were kept much lower to produce a short-term phenotypic response to selection within available field testing resources.} \jpg{My intent in rewording this is to emphasize that Ne is the important difference between the scenarios.}
 
@@
 Numerous experiments have shown that the selected populations and the hybrids formed from them \rev{exhibit genetic gain for hybrid yield as well plant architecture and tolerance to high-density planting} \citep{smith1983evaluation, keeratinijakal1993responses, schnicker1993interpopulation, holthaus1995population, brekke2011selection, brekke2011selectionb, edwards2011changes, lauer2012morphological}.
-We find that heterozygosity falls more than expected across the genome as a whole, and, \rev{though drift imposes limitations on the power to detect selection at individual loci,} genomic regions of extremely low diversity evident at cycle 16 are unlikely to be produced by simulation.
+We find that heterozygosity falls more than expected across the genome as a whole, and, \rev{though drift imposes limitations on the power to detect selection at individual loci,} genomic regions of extremely low diversity evident at cycle 16 are unlikely to be produced by drift alone.
 We further show that an identity-by-descent, haplotype-based approach provides additional power to identify selected regions, as it can distinguish between the fixation of rare and common haplotypes. 
-\rev{These analyses show that most likely targets of selection occur at different loci in the two populations, a result consistent with analyses in commercial breeding programs \citep{feng2006temporal} and which may help explain the lack of selection seen in previous analyses across numerous breeding programs \citep{van2012historical}.}
+\rev{These analyses show that the most likely targets of selection occur at different loci in the two populations, a result consistent with analyses in commercial breeding programs \citep{feng2006temporal} and which may help explain the lack of selection seen in previous analyses across numerous breeding programs \citep{van2012historical}.}
 
@@
 In the case of true overdominance, selection should thus lead to decreased heterozygosity at the same locus in both populations as complementary haplotypes are fixed in each group \rev{\citep[e.g.][]{guo2014maize}}.
+\jpg{One possible criticism is that overdominance could still be happening when each allele is on multiple founder haplotypes.
+Could say something like "Although we cannot rule out soft sweeps of complementary overdominant alleles, the observed pattern more parsimoniously favors..."}
 We find \rev{little evidence to support this genetic phenomenon, finding only two 2 cM windows genome-wide in which both populations show significantly reduced heterozygosity.} 
@@
 Because deleterious alleles will be rare in both populations, most haplotypes in the second population will have a different suite of deleterious variants and will complement the fixed haplotype reasonably well.
-\rev{Selection against homozygosity of the fixed haplotype will thus} have little impact on diversity in the second population.
+\rev{We expect that selection against homozygosity of the fixed haplotype will thus} have little impact on diversity in the second population.
 
-Our results are consistent with the dominance and pseudo-overdominance models, but we caution that the exact outcome of particular models will depend strongly on the effects of selection, drift, and the frequencies of beneficial alleles.
+Although our results better fit the simpler dominance model, we caution that the exact outcome of particular models will depend strongly on the effects of selection, drift, and the frequencies of beneficial alleles.
 Disentangling these factors remains a major challenge that will require careful simulation of maize breeding using a wider range of parameters. 
@@
 Reciprocal recurrent selection serves as a model for the method of hybrid maize improvement \citep{duvick2004long}.
-Similar patterns of diversity and population structure can be seen broadly across North American maize germplasm \citep{van2012historical}, suggesting that genetic drift has most likely played a large role in the current genetic structure of modern maize.
+\jpg{I'm not sure where this paragraph belongs.}Similar patterns of diversity and population structure can be seen broadly across North American maize germplasm \citep{van2012historical}, suggesting that genetic drift has most likely played a large role in the current genetic structure of modern maize.
 Nonetheless, breeding has contributed substantially to yield improvements \citep{duvick2005contribution}, and the identification and characterization of loci conferring phenotypic improvement in BSSS and BSCB1 could reveal some of the genetic mechanisms by which maize yield has steadily improved over the decades.

### 2015-08-27 response.tex
accept most jpg changes

@@
 
-\response{We have endeavored to streamline the discussion where possible while incorporating additional citations (as above) and discussion of load as mentioend by reviewer 1.}
+\response{We have incorporated additional changes requested by reviewers but overall have streamlined the discussion considerably.}
 

### 2015-08-27 main.tex
accept most jpg changes

@@
 \abstract{
-Although maize is naturally an outcrossing organism, modern breeding develops of highly inbred lines that are used in controlled crosses to produce hybrids. 
+Although maize is naturally an outcrossing organism, modern breeding utilizes highly inbred lines in controlled crosses to produce hybrids. 
 The United States Department of Agriculture reciprocal recurrent selection experiment between the Iowa Stiff Stalk Synthetic (BSSS) and the Iowa Corn Borer Synthetic No.
-1 (BSCB1) populations represents one of the longest \rev{running experiments to understand the response to selection} \jpg{changed the wording here, because the experiment's goal was more broad than undesrtanding the genetic basis} for hybrid performance.
+1 (BSCB1) populations represents one of the longest \rev{running experiments to understand the response to selection} for hybrid performance.
 To investigate the genomic impact of this selection program, \rev{we genotyped the progenitor lines and over 600 individuals across multiple cycles of selection using a genome-wide panel of $\sim 40,000$ SNPs.
@@
 These studies have examined a broad spectrum of germplasm at various points in the history of maize to search for the signals of population structure and artificial selection \citep{Hufford2012b, van2012historical,jiao2012genome}.
-Although selective sweeps from the initial domestication of maize are clearly visible, localized genomic signals of selection during modern breeding remain difficult to observe \citep{Hufford2012b,van2012historical} despite steady, heritable improvement in phenotype \citep{duvick2005contribution}.
+Although selective sweeps from the initial domestication of maize are clearly visible, localized genomic signals of selection during modern breeding are difficult to observe \citep{Hufford2012b,van2012historical} despite steady, heritable improvement in phenotype \citep{duvick2005contribution}.
 The lack of distinct selection signals \rev{in broad germplasm collections may be due to population-specific selection within germplasm subgroups; in this case selection should be easier to detect in populations maintained in an individual program.}
-This \rev{possibliity} is supported by the measured success in identifying targets of selection in individual experimental populations \rev{under directional selection for specific phenotypes such as seed size and ear number} \citep{hirsch2014insights, beissinger2014genome}.
+This possibliity is supported by the measured success in identifying targets of selection in individual experimental populations \rev{under directional selection for specific phenotypes such as seed size and ear number} \citep{hirsch2014insights, beissinger2014genome}.
 
 \rev{In this study, we examine a different experiment in which the method and target of selection --- reciprocal recurrent selection for hybrid yield --- closely mirror those used in the generation of modern maize hybrids \citep{comstock1949breeding, duvick2004long}.
-Reciprocal recurrent selection is a method, initially proposed by \citet{comstock1949breeding}, in which lines from two populations are evaluated based on the phenotype of the hybrids each line produces when crossed with the opposing population.
+Reciprocal recurrent selection is a method, initially proposed by \citet{comstock1949breeding}, in which lines from two populations are evaluated based on the phenotype of the hybrids each line produces when crossed with lines from the opposing population.
 The selected lines are then intermated within each population to generate lines for the next cycle of recurrent selection.
-%Reciprocal recurrent selection simultaneously improves two populations for their ability to produce high yielding hybrids when individuals from the two populations are crossed. 
 This procedure results in two closed and genetically isolated populations that simultaneously evolve improved combining ability with one another. 
 The USDA-ARS at Ames, Iowa has conducted a reciprocal recurrent selection program (hereafter referred to as the Iowa RRS) with the Iowa Stiff Stalk Synthetic (BSSS) and Iowa Corn Borer Synthetic No.
-1 (BSCB1) populations for 18 cycles of selection for hybrid yield \citep{penny1971twenty}.
-\jpg{I think we need Jode's reference here as well if we are going to mention it has gone through 18 cycles.}
+1 (BSCB1) populations for 18 cycles of selection for hybrid yield \citep{penny1971twenty, edwards2011changes}.
 This program represents one of the best-documented public experiments on selection for combining ability and hybrid performance. 
@@
 coli} \citep{tenaillon2012molecular} in which the effects of selection on diversity are readily discernible.
-A key difference between these studies and those in maize are the effective population sizes, which were kept much lower to produce a short-term phenotypic response to selection within available field testing resources.} \jpg{My intent in rewording this is to emphasize that Ne is the important difference between the scenarios.}
+A key difference between these studies and those in maize are the effective population sizes, which were kept much lower to produce a short-term phenotypic response to selection within available field testing resources.}
 
@@
 In the case of true overdominance, selection should thus lead to decreased heterozygosity at the same locus in both populations as complementary haplotypes are fixed in each group \rev{\citep[e.g.][]{guo2014maize}}.
-\jpg{One possible criticism is that overdominance could still be happening when each allele is on multiple founder haplotypes.
-Could say something like "Although we cannot rule out soft sweeps of complementary overdominant alleles, the observed pattern more parsimoniously favors..."}
-We find \rev{little evidence to support this genetic phenomenon, finding only two 2 cM windows genome-wide in which both populations show significantly reduced heterozygosity.} 
-The observed pattern instead favors a dominance model, where fixation of a haplotype in one population simply selects against that same haplotype in the other population. 
+We find \rev{little evidence to support this genetic phenomenon, finding only two 2 cM windows genome-wide in which both populations show significantly reduced heterozygosity. 
+Although we cannot rule out soft sweeps of complementary overdominant alleles, the observed pattern more parsimoniously favors} a dominance model, in which fixation of a haplotype in one population simply selects against that same haplotype in the other population.
 \rev{Although strongly deleterious variants were likely purged during the inbreeding process leading to the founder lines, many weakly deleterious alleles can be found segregating at low frequencies among inbreds \citep{Mezmouk2014}.}
@@
 Reciprocal recurrent selection serves as a model for the method of hybrid maize improvement \citep{duvick2004long}.
-\jpg{I'm not sure where this paragraph belongs.}Similar patterns of diversity and population structure can be seen broadly across North American maize germplasm \citep{van2012historical}, suggesting that genetic drift has most likely played a large role in the current genetic structure of modern maize.
+Similar patterns of diversity and population structure can be seen broadly across North American maize germplasm \citep{van2012historical}, suggesting that genetic drift has most likely played a large role in the current genetic structure of modern maize.
 Nonetheless, breeding has contributed substantially to yield improvements \citep{duvick2005contribution}, and the identification and characterization of loci conferring phenotypic improvement in BSSS and BSCB1 could reveal some of the genetic mechanisms by which maize yield has steadily improved over the decades.

### 2015-08-28 main.tex
paragraph

@@
 In other cases, the haplotypes which eventually fixed were at moderate frequency in the cycle 0 populations and drift to fixation in the majority of simulations.
+Several key inbreds in the ‘stiff-stalk’ heterotic group --- B73, B37 and B14 --- were derived from the BSSS population \citep{darrah1986, troyer1999background}. 
+B37 and B14 were derived from cycle 0, and B73 was derived from a half-sib recurrent selection program also started with the BSSS population. 
+We examined these three inbreds at the pericentromeric regions listed in Table \ref{tab:fix}, and found that in most cases they carry different haplotypes from those that rose to high frequency in the RRS experiment.
 
@@
 \rev{Similar observations have been recently made in other reciprocal recurrent selection programs, even in comparisons of multiple replicated populations \citep{romay2012effect,lamkey2014relative}.
-These patterns differ markedly, however, from similar experimental evolution in systems such as \emph{Drosophila} \citep{turner2011population, burke2010genome} and \emph{E.
+Reciprocal recurrent selection serves as a model for the method of hybrid maize improvement \citep{duvick2004long}.
+Similar patterns of diversity and population structure can be seen broadly across North American maize germplasm \citep{van2012historical}, suggesting that genetic drift has most likely played a large role in the current genetic structure of modern maize.
+These patterns differ markedly, however, from experimental evolution in systems such as \emph{Drosophila} \citep{turner2011population, burke2010genome} and \emph{E.
 coli} \citep{tenaillon2012molecular} in which the effects of selection on diversity are readily discernible.
@@
 
-Reciprocal recurrent selection serves as a model for the method of hybrid maize improvement \citep{duvick2004long}.
-Similar patterns of diversity and population structure can be seen broadly across North American maize germplasm \citep{van2012historical}, suggesting that genetic drift has most likely played a large role in the current genetic structure of modern maize.
-Nonetheless, breeding has contributed substantially to yield improvements \citep{duvick2005contribution}, and the identification and characterization of loci conferring phenotypic improvement in BSSS and BSCB1 could reveal some of the genetic mechanisms by which maize yield has steadily improved over the decades.
-Several key inbreds in the ‘stiff-stalk’ heterotic group --- B73, B37 and B14 --- were derived from the BSSS population \citep{darrah1986, troyer1999background}. 
-%B37 and B14 were derived from cycle 0, and B73 was derived from a half-sib recurrent selection program also started with the BSSS population. 
-We examined these three inbreds at the pericentromeric regions listed in Table \ref{tab:fix}, and found that in most cases they carry different haplotypes from those that rose to high frequency in the RRS experiment.
-Near isogenic lines that substitute these centromeric haplotypes from key inbreds could be used to determine if the haplotypes at high frequency in cycle 16 of BSSS provide a phenotypic advantage when crossed with BSCB1.
-Our observation of population-specific selection is consistent with a model of selection for complementation in the Iowa RSS experiment, and the lack of extreme differentiation seen over time across US Corn Belt germplasm may suggest a role for such selection in maize hybrid improvement in general \citep{van2012historical}.
-
 \begin{acknowledgments}

### 2015-08-28 sup1.tex
paragraph

@@
 	\subcaption*{BSCB1 Founders Genotyped} 
-	\begin{tabular}{lll}
+	\begin{tabular}{ll}
 	Inbred & Background / Pedigree \\ \hline
-	I205 & Iodent &  &  \\ 
+	I205 & Iodent   \\ 
 	Oh51A & [(OH56xWf9)Oh56] (Wooster Clarage x ?)   \\ 

### 2015-08-28 main.tex
final maybe

@@
 
-\setcounter{footnote}{1}
+\setcounter{footnote}{0}
 
@@
 \author{
- \small\sfbf{Justin P. Gerke$^{\ast,1}$\thanks{
+ \small\sfbf{Justin P. Gerke$^{\ast}$\thanks{
 Corresponding author:  DuPont Pioneer, 8305 NW 62ND Ave, P.O. Box 7060
@@
  %  \vspace{-3mm}
+  \end{center}
+
    \caption{\rev{Heterozygosity} in each cycle across chromosome 4 of the BSSS (left) and  BSCB1 (right) plotted on the physical (top) and genetic (bottom) map.
@@
 Each data point is color-coded based on a linear transformation of recombination rate (red = low recombination rate).
-\rev{Shaded regions represent 2 cM windows with heterozygosity values significantly lower than expected by simulation at a given cycle at $P<0.001$.} Plots for other chromosomes are shown in in Figure \ref{fig:others2}}.
-Comparing the physical and genetic maps reveals that the regions of fixation are extremely small genetically, but some contain much of the physical content of the chromosome.
+\rev{Shaded regions represent 2 cM windows with heterozygosity values significantly lower than expected by simulation at a given cycle at $P<0.001$.} Plots for other chromosomes are shown in in Figure \ref{fig:others2}.}
 
-\vspace{-6mm}
+\vspace{-4mm}
     \label{fig:genphys}
-  \end{center}
 \end{figure*}
@@
 \caption{Ancestry of haplotypes fixed in the cycle 16 population.}
+\label{tab:fix}
 \fontsize{10}{12}\sf
@@
 
-\label{tab:fix}
 \begin{tabular}{llllll}
@@
 \begin{figure*}[tb]  
+	\centering 
    \includegraphics[width=0.7\linewidth]{Fig_S1_combined.pdf}
@@
 \rev{Similar observations have been recently made in other reciprocal recurrent selection programs, even in comparisons of multiple replicated populations \citep{romay2012effect,lamkey2014relative}.
-Reciprocal recurrent selection serves as a model for the method of hybrid maize improvement \citep{duvick2004long}.
-Similar patterns of diversity and population structure can be seen broadly across North American maize germplasm \citep{van2012historical}, suggesting that genetic drift has most likely played a large role in the current genetic structure of modern maize.
+Reciprocal recurrent selection serves as a model for the method of hybrid maize improvement \citep{duvick2004long}, and similar patterns of diversity and population structure can be seen broadly across North American maize germplasm \citep{van2012historical}.
+Genetic drift has thus most likely played a large role in the current genetic structure of modern maize.
 These patterns differ markedly, however, from experimental evolution in systems such as \emph{Drosophila} \citep{turner2011population, burke2010genome} and \emph{E.
@@
 \rev{We expect that selection against homozygosity of the fixed haplotype will thus} have little impact on diversity in the second population.
-
-Although our results better fit the simpler dominance model, we caution that the exact outcome of particular models will depend strongly on the effects of selection, drift, and the frequencies of beneficial alleles.
-Disentangling these factors remains a major challenge that will require careful simulation of maize breeding using a wider range of parameters. 
+Although our results better fit the simpler dominance model, the ability to distinguish between models will depend strongly on allele frequencies as well as the effects of selection, and drift. 
 This is especially true because in a model of hybrid complementation, genetic drift in one population can alter the selective value of alleles in the other population. 
-When drift and selection interact in such a manner, tests for selection that attempt to independently partition the effects of each force may not provide the full picture. 
-Furthermore, uncertainties in the fine-scale relationship between genetic and physical distance make it difficult to assign significance in forward-simulation approaches. 
-In the end, the best test for selection of specific genomic regions will ultimately be conducted by phenotypic observation in the field in balanced tests of different haplotypes. 
-In the case of individual breeding programs, genome-wide genotyping such as we conducted here can identify the lines carrying the recombinant haplotypes and introgressions necessary to conduct such an experiment.
+Given these complexities, empirical evaluation of the effects of putatively selected haplotypes will play a key role in distinguishing opposing genetic models.
 

### 2015-08-30 main.tex
final edits
final

@@
  \small\sfbf{Justin P. Gerke$^{\ast}$\thanks{
-Corresponding author:  DuPont Pioneer, 8305 NW 62ND Ave, P.O. Box 7060
-Johnston, IA, 50131   E-mail: \mbox{justin.gerke@gmail.com}}, Jode W. Edwards$^{\dag}$, Katherine E. Guill$^{\ddag}$, Jeffrey Ross-Ibarra$^{\S}$}\thanks{
+Corresponding author: E-mail: \mbox{justin.gerke@gmail.com}}, Jode W. Edwards$^{\dag}$, Katherine E. Guill$^{\ddag}$, Jeffrey Ross-Ibarra$^{\S}$}\thanks{
 Corresponding author:  Department of Plant Sciences, University of California, Davis, California 95616. 
@@
 
-%\date{Revised manuscript for \emph{Genetics}, \today}
+\date{Revised manuscript for \emph{Genetics}, \today}
 
@@
 coli} \citep{tenaillon2012molecular} in which the effects of selection on diversity are readily discernible.
-A key difference between these studies and those in maize are the effective population sizes, which were kept much lower to produce a short-term phenotypic response to selection within available field testing resources.}
+A key difference between these studies and those in maize are the effective population sizes, which were kept much lower in maize to produce a short-term phenotypic response to selection within available field testing resources.}
 
@@
 \begin{acknowledgments}
-We thank Oscar ``Howie'' Smith and members of the Ross-Ibarra lab for comments on earlier versions of the manuscript.
+We thank Oscar ``Howie'' Smith,  members of the Ross-Ibarra lab, and two anonymous reviewers for comments on earlier versions of the manuscript.
+We also thank the editor Stephen Wright for his patience.
 J.P.G received support for this research as a Merck Fellow of the Life Sciences Research Foundation.

### 2015-09-12 main.tex
made Stephen's typo changes

@@
 We therefore took steps to improve the positions of the SNP markers on the genetic and physical maps relative to version 5A.59 of the maize genome assembly. 
-The probable physical position of each SNP based was obtained by comparing SNP context sequences to the genome sequence. 
+The probable physical position of each SNP was obtained by comparing SNP context sequences to the genome sequence. 
 For this purpose, SNP context sequences were defined as the sequence 25 bp upstream of the SNP, the bp representing the SNP itself, followed by 25 bp downstream of the SNP, making a total sequence length of 51bp.
@@
 Selection was executed at random in each simulation, so the patterns observed across simulations represent the expected distribution of effects caused only by recombination and genetic drift. 
-We conducted 10,000 simulations, \rev{modeling recombination using the IBM genetic map \citep{lee2002expanding}, which is based on a cross between between a BSSS-derived inbred line used to construct the reference genome (B73) and an inbred with  with co-ancestry from both the BSSS and BSCB1 founder germplasm (Mo17).}
+We conducted 10,000 simulations, \rev{modeling recombination using the IBM genetic map \citep{lee2002expanding}, which is based on a cross between between a BSSS-derived inbred line used to construct the reference genome (B73) and an inbred with co-ancestry from both the BSSS and BSCB1 founder germplasm (Mo17).}
 

### 2015-09-15 Fig3/Fig3plotting.R
figs 1-5 converted to jpeg

@@ -22,4 +22,6 @@ ggplot(data=both) + aes(x=physpos/1000000, y=H, colour=Population) + facet_grid(
 dev.off()
 
 
-
+jpeg("../fig3.jpg", quality=500,  width=700, height=900)
+ggplot(data=both) + aes(x=physpos/1000000, y=H, colour=Population) + facet_grid(LG~Population) + geom_point(size=1) + scale_colour_manual(values=c("blue", "red2")) + theme_bw() + xlab("Physical Position (MB)")
+dev.off()

### 2015-09-15 Fig4/fig4.R

@@ -108,3 +108,9 @@ save_plot("../fig4_combined.pdf", allfour,
           base_aspect_ratio = 1.3
 )
 
+save_plot("../fig4_combined.jpg", allfour,
+          ncol = 2, # we're saving a grid plot of 2 columns
+          nrow = 2, # and 2 rows
+          # each individual subplot should have an aspect ratio of 1.3
+          base_aspect_ratio = 1.3
+)

### 2015-09-15 FigS1/combineplot.R

@@ -12,3 +12,10 @@ save_plot("../fig_S1_combined.pdf", combined,
           nrow = 1, 
           base_aspect_ratio = 1.1
 )
+
+combined = plot_grid(plot_BSSS, plot_BSCB1, labels=c("A", "B"))
+save_plot("../fig_5.jpg", combined,
+          ncol = 2, 
+          nrow = 1, 
+          base_aspect_ratio = 1.1
+)

### 2015-09-15 Fig1/pcafuncs.R
added code for figures 1 and 2

@@ -1,0 +1,117 @@
+rm(list=ls())
+load("../data/new_full_set.RData")
+source("../rcode/writing_functions.R")
+source("../rcode/building_functions.R")
+source("../rcode/formatting_functions.R")
+
+
+convert_numeric = function(allele){
+  calls = unique(allele)[!is.na(unique(allele))]
+  is.homozygote = sapply(calls, function(x){
+    myalleles = unlist(strsplit(x,''))
+    boolean = myalleles[1] == myalleles[2]
+    return(boolean)
+  })
+  allele[allele == calls[is.homozygote][1]] = 0 
+  allele[allele == calls[is.homozygote][2]] = 2
+  allele[allele == calls[!is.homozygote][1]] = 1
+  return(as.numeric(allele))
+  }
+  
+
+mynumeric = t(apply(new_full_set$calls,1,convert_numeric))
+rownames(mynumeric) = rownames(new_full_set$calls)
+colnames(mynumeric) = colnames(new_full_set$calls)
+
+prune_monos = function(datamat){
+  polymorphics = apply(datamat, 1, function(x){sum(!is.na(unique(x)))})
+  keeps = polymorphics>1
+  newmat = datamat[keeps,]
+  return(newmat)
+}
+
+centermean = function(datamat){
+  mymeans = getmeans(datamat)
+  centered = (datamat - mymeans)
+  centered[is.na(centered)] = 0
+  return(list(centered, mymeans))
+}
+
+getmeans = function(datamat){
+  means = apply(datamat, 1, mean, na.rm=T)
+  return(means)
+}
+
+runPCA = function(datamat){
+  centerfunc = centermean(datamat)
+  means = centerfunc[[2]]
+  centered = centerfunc[[1]]
+  centered = t(centered)
+  centered = centered/(sqrt(ncol(centered)))
+  PCs = prcomp(centered, center=FALSE, retx=T)
+  loadings = t(t(PCs$rotation) *PCs$sdev)
+  projection = centered %*% loadings
+  varexp = (PCs$sdev^2)/sum(PCs$sdev^2)
+  results = list(M=centered, projections=projection, loadings=loadings, eigs = PCs$sdev^2, means = means, nmarks=ncol(centered), varexp = varexp)
+  return(results)
+}
+
+make_projection = function(PCAobj,newdat){
+  newdat = t(newdat)
[…]

### 2015-09-15 main.tex
no blue text
added current address
no rev

@@
 \newcommand{\jpg}[1]{\todo[size=\scriptsize, color=gerkegreen]{#1}}
-\newcommand{\rev}[1]{\textcolor{blue}{\bf #1}}
+\newcommand{\rev}[1]{#1}
 
@@
  \small\sfbf{Justin P. Gerke$^{\ast}$\thanks{
-Corresponding author: E-mail: \mbox{justin.gerke@gmail.com}}, Jode W. Edwards$^{\dag}$, Katherine E. Guill$^{\ddag}$, Jeffrey Ross-Ibarra$^{\S}$}\thanks{
-Corresponding author:  Department of Plant Sciences, University of California, Davis, California 95616. 
+Corresponding author. Current Address: DuPont Pioneer, Johnston, Iowa 50131. E-mail: \mbox{justin.gerke@gmail.com}
+}, Jode W. Edwards$^{\dag}$, Katherine E. Guill$^{\ddag}$, Jeffrey Ross-Ibarra$^{\S}$}\thanks{
+Corresponding author.  Department of Plant Sciences, University of California, Davis, California 95616. 
     E-mail: \mbox{rossibarra@ucdavis.edu}},\\ 

### 2015-09-17 Fig3/Fig3plotting.R
final touches, I think

@@ -22,6 +22,6 @@ ggplot(data=both) + aes(x=physpos/1000000, y=H, colour=Population) + facet_grid(
 dev.off()
 
 
-jpeg("../fig3.jpg", quality=500,  width=700, height=900)
+jpeg("../Fig_S2.jpg", quality=1000,  width=700, height=900)
 ggplot(data=both) + aes(x=physpos/1000000, y=H, colour=Population) + facet_grid(LG~Population) + geom_point(size=1) + scale_colour_manual(values=c("blue", "red2")) + theme_bw() + xlab("Physical Position (MB)")
 dev.off()

### 2015-09-17 Fig4/allchroms.R

@@ -86,7 +86,7 @@ ChrBSCB1 = list(phys=myphysplot, gen=mygenplot)
 
 library(cowplot)
 allfour = plot_grid(ChrBSSS[["phys"]], ChrBSCB1[["phys"]], ChrBSSS[["gen"]], ChrBSCB1[["gen"]], labels=c("A", "B", "C", "D"), ncol=2)
-save_plot(paste("./chrom", Chrom, "_combined.pdf", sep=''), allfour,
+save_plot(paste("./chrom", Chrom, "_combined.jpg", sep=''), allfour,
           ncol = 2, # we're saving a grid plot of 2 columns
           nrow = 2, # and 2 rows
           # each individual subplot should have an aspect ratio of 1.3

### 2015-09-17 FigS2/FigS2plotting.R

@@ -34,8 +34,9 @@ genplot = ggplot(data=alldata) + blocks + BSSS_H + BSCB1_H + facet_grid(LG~.) + 
   ggtitle("Genetic Space")
 #dev.off()
 
+library(cowplot)
 combined = plot_grid(physplot, genplot, labels=c("A", "B"), nrow=2)
-save_plot("../fig_S2_combined.pdf", combined,
+save_plot("../Fig3.jpg", combined,
           base_aspect_ratio = 0.9,
           base_height = 10
 )

### 2015-09-17 make_fileS2.R

@@ -1,0 +1,7 @@
+library(dplyr)
+
+data = read.table("./data/2cM_sim_collated_results.txt", header=T)
+mydata = select(data, Cycle, LG, phystart, phystop, 
+                genstart, genstop, unph_H_BSSS, unph_H_BSCB, 
+                H_BSSS, H_BSCB, BSSS_quantile_0.001, BSCB_quantile_0.001)
+write.table(mydata, file="File_S2.txt", row.names=F, quote=F)
